---
title: "Maternal Brain: analyses of proportion data by behavior (caste*day)"
author: "M.A. Hagadorn"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = TRUE,
	background = "#F7F7F7",
	highlight = TRUE)
```

## Load in Required Packages

```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(plyr)
library(dplyr)
library(Rmisc)
library(ggplot2)
library(stats)
library(nortest)
library(MASS)
library(here)
library(pwr) #power analysis
library(effectsize)
library(rcompanion) #plotNormalDensity
library(lme4) #glmer

installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```


# Individual Behaviors: Caste * DayGrouping
## Load in Data: individual behaviors proportion
```{r loaddata, include=FALSE}
#read in data
#data set only contains 'no manipulation' queens and workers
#NO FOSTER DATA INCLUDED
behavior_prop <- read.csv2("G:/DesktopDocuments/Documents/PHD/Manuscripts/Bee/Bimp_MaternalBrainVolumetric/BehaviorData/BehaviorData_ProportionData_INDxT.csv", sep=",", stringsAsFactors = FALSE)
print(str(behavior_prop))

#assign data columns as numeric
numb_cols <- c(6:18)
behavior_prop[,numb_cols] <- apply(behavior_prop[ , numb_cols], 2, function(x) as.numeric(x))

#Assign factors
behavior_prop$colony <- factor(behavior_prop$colony)
behavior_prop$caste <- factor(behavior_prop$caste)
behavior_prop$DayGrouping <- factor(behavior_prop$DayGrouping)

print(str(behavior_prop))
```


## Larval Feeding
### Summary Statistics
```{r ss_LF}
ss_LF <- summarySE(behavior_prop,
                          measurevar = "LF",
                          groupvars = c("caste","DayGrouping"))
kable(ss_LF)
```


### Generalized Linear Model (BIONOMIAL)
```{r AOD_LF}
#mixed model to check for colony effect
summary(glmer(LF~caste*DayGrouping + (1|colony), data=behavior_prop, family=binomial(link = "logit")))

#no effect of colony

#linear models
AOD_LF.intx <- glm(LF~caste*DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
AOD_LF <- glm(LF~caste+DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
LF <- glm(LF~caste, data=behavior_prop, family=binomial(link = "logit"))

####
#now lets compare models to see what the best fit is.
anova(AOD_LF.intx, AOD_LF, LF)

#model with the lowest ACI should be used
#if difference is less than 2.0, then go with simpler
AIC(AOD_LF.intx, AOD_LF, LF)

#AIC for no interaction is lower
#most appropriate model is no interaction

# summary(AOD_LF)
# anova(AOD_LF, test="Chisq")
summary(LF)
```



## Pollen Feeding
### Summary Statistics
```{r ss_PF}
ss_PF <- summarySE(behavior_prop,
                          measurevar = "PF",
                          groupvars = c("caste","DayGrouping"))
kable(ss_PF)
```


### Generalized Linear Model (BIONOMIAL)
```{r AOD_PF}
#mixed model to check for colony effect
summary(glmer(PF~caste*DayGrouping + (1|colony), data=behavior_prop, family=binomial(link = "logit")))

#no effect of colony

#linear models
AOD_PF.intx <- glm(PF~caste*DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
AOD_PF <- glm(PF~caste+DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
PF <- glm(PF~caste, data=behavior_prop, family=binomial(link = "logit"))

####
#now lets compare models to see what the best fit is.
anova(AOD_PF.intx, AOD_PF, PF)

#model with the lowest ACI should be used
#if difference is less than 2.0, then go with simpler
AIC(AOD_PF.intx, AOD_PF, PF)

#AIC for no interaction is lower
#most appropriate model is no interaction


summary(PF)

```




## Drinking Sugar Water
### Summary Statistics
```{r ss_DS}
ss_DS <- summarySE(behavior_prop,
                          measurevar = "DS",
                          groupvars = c("caste","DayGrouping"))
kable(ss_DS)
```


### Generalized Linear Model (BIONOMIAL)
```{r AOD_DS}
#mixed model to check for colony effect
summary(glmer(DS~caste*DayGrouping + (1|colony), data=behavior_prop, family=binomial(link = "logit")))

#no effect of colony

#linear models
AOD_DS.intx <- glm(DS~caste*DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
AOD_DS <- glm(DS~caste+DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
DS <- glm(DS~caste, data=behavior_prop, family=binomial(link = "logit"))
####
#now lets compare models to see what the best fit is.
anova(AOD_DS.intx, AOD_DS, DS)

#model with the lowest ACI should be used
#if difference is less than 2.0, then go with simpler
AIC(AOD_DS.intx, AOD_DS, DS)

#AIC for no interaction is lower
#most appropriate model is no interaction


# summary(AOD_DS)
summary(DS)

```





## Incubation
### Summary Statistics
```{r ss_ICB}
ss_ICB <- summarySE(behavior_prop,
                          measurevar = "ICB",
                          groupvars = c("caste","DayGrouping"))
kable(ss_ICB)
```


### Generalized Linear Model (BIONOMIAL)
```{r AOD_ICB}
#mixed model to check for colony effect
summary(glmer(ICB~caste*DayGrouping + (1|colony), data=behavior_prop, family=binomial(link = "logit")))

#no effect of colony

#linear models
AOD_ICB.intx <- glm(ICB~caste*DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
AOD_ICB <- glm(ICB~caste+DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
ICB <- glm(ICB~caste, data=behavior_prop, family=binomial(link = "logit"))

####
#now lets compare models to see what the best fit is.
anova(AOD_ICB.intx, AOD_ICB)
anova(AOD_ICB, ICB)

#model with the lowest ACI should be used
#if difference is less than 2.0, then go with simpler
AIC(AOD_ICB.intx, AOD_ICB, ICB)

#within 2.0; going with simplier model
#most appropriate model is no interaction

# 
# summary(AOD_ICB)
summary(ICB)

```





## Inspecting
### Summary Statistics
```{r ss_INS}
ss_INS <- summarySE(behavior_prop,
                          measurevar = "INS",
                          groupvars = c("caste","DayGrouping"))
kable(ss_INS)
```


### Generalized Linear Model (BIONOMIAL)
```{r AOD_INS}
#mixed model to check for colony effect
#summary(glmer(INS ~ caste*DayGrouping + (1|colony), data=behavior_prop, family=binomial(link = "logit")))
###Complete separation because of 0 and 1.0

#no effect of colony

#linear models
AOD_INS.intx <- glm(INS~caste*DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
AOD_INS <- glm(INS~caste+DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
INS <- glm(INS~caste, data=behavior_prop, family=binomial(link = "logit"))

####
#now lets compare models to see what the best fit is.
anova(AOD_INS.intx, AOD_INS)

#model with the lowest ACI should be used
#if difference is less than 2.0, then go with simpler
AIC(AOD_INS.intx, AOD_INS, INS)

#AIC is lower for the no interactoin model
#most appropriate model is no interaction


# summary(AOD_INS)
summary(INS)

```




## Manipulating Wax
### Summary Statistics
```{r ss_MW}
ss_MW <- summarySE(behavior_prop,
                          measurevar = "MW",
                          groupvars = c("caste","DayGrouping"))
kable(ss_MW)
```


### Generalized Linear Model (BIONOMIAL)
```{r AOD_MW}
#mixed model to check for colony effect
summary(glmer(MW ~ caste*DayGrouping + (1|colony), data=behavior_prop, family=binomial(link = "logit")))


#no effect of colony

#linear models
AOD_MW.intx <- glm(MW~caste*DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
AOD_MW <- glm(MW~caste+DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
MW <- glm(MW~caste, data=behavior_prop, family=binomial(link = "logit"))

####
#now lets compare models to see what the best fit is.
anova(AOD_MW.intx, AOD_MW)

#model with the lowest ACI should be used
#if difference is less than 2.0, then go with simpler
AIC(AOD_MW.intx, AOD_MW, MW)

#AIC is lower for the no interaction model
#most appropriate model is no interaction


# summary(AOD_MW)
summary(MW)

```





## Perching
### Summary Statistics
```{r ss_PER}
ss_PER <- summarySE(behavior_prop,
                          measurevar = "PER",
                          groupvars = c("caste","DayGrouping"))
kable(ss_PER)
```


### Generalized Linear Model (BIONOMIAL)
```{r AOD_PER}
#mixed model to check for colony effect
summary(glmer(PER ~ caste*DayGrouping + (1|colony), data=behavior_prop, family=binomial(link = "logit")))


#no effect of colony

#linear models
AOD_PER.intx <- glm(PER~caste*DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
AOD_PER <- glm(PER~caste+DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
PER <- glm(PER~caste, data=behavior_prop, family=binomial(link = "logit"))

####
#now lets compare models to see what the best fit is.
anova(AOD_PER.intx, AOD_PER)

#model with the lowest ACI should be used
#if difference is less than 2.0, then go with simpler
AIC(AOD_PER.intx, AOD_PER, PER)

#AIC is lower for the no interaction model
#most appropriate model is no interaction


#summary(AOD_PER)
summary(PER)
```




## Buzzing
### Summary Statistics
```{r ss_BZ}
ss_BZ <- summarySE(behavior_prop,
                          measurevar = "BZ",
                          groupvars = c("caste","DayGrouping"))
kable(ss_BZ)
```

No occurrences, no analysis.



## Patrolling
### Summary Statistics
```{r ss_PAT}
ss_PAT <- summarySE(behavior_prop,
                          measurevar = "PAT",
                          groupvars = c("caste","DayGrouping"))
kable(ss_PAT)
```

### Generalized Linear Model (BIONOMIAL)
```{r AOD_PAT}
#mixed model to check for colony effect
summary(glmer(PAT ~ caste*DayGrouping + (1|colony), data=behavior_prop, family=binomial(link = "logit")))


#no effect of colony

#linear models
AOD_PAT.intx <- glm(PAT~caste*DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
AOD_PAT <- glm(PAT~caste+DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
PAT <- glm(PAT~caste, data=behavior_prop, family=binomial(link = "logit"))

####
#now lets compare models to see what the best fit is.
anova(AOD_PAT.intx, AOD_PAT,  test="Chisq")

#model with the lowest ACI should be used
#if difference is less than 2.0, then go with simpler
AIC(AOD_PAT.intx, AOD_PAT, PAT)

#AIC is lower for the no interaction model
#most appropriate model is no interaction


summary(PAT)

```




## Fanning
### Summary Statistics
```{r ss_FAN}
ss_FAN <- summarySE(behavior_prop,
                          measurevar = "FAN",
                          groupvars = c("caste","DayGrouping"))
kable(ss_FAN)
```

No occurrences, no analysis.




## Egg-laying
### Summary Statistics
```{r ss_EL}
ss_EL <- summarySE(behavior_prop,
                          measurevar = "EL",
                          groupvars = c("caste","DayGrouping"))
kable(ss_EL)
```

### Generalized Linear Model (BIONOMIAL)
```{r AOD_EL}
#mixed model to check for colony effect
summary(glmer(EL ~ caste*DayGrouping + (1|colony), data=behavior_prop, family=binomial(link = "logit")))


#no effect of colony

#linear models
AOD_EL.intx <- glm(EL~caste*DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
AOD_EL <- glm(EL~caste+DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
EL <- glm(EL~caste, data=behavior_prop, family=binomial(link = "logit"))

####
#now lets compare models to see what the best fit is.
anova(AOD_EL.intx, AOD_EL,  test="Chisq")

#model with the lowest ACI should be used
#if difference is less than 2.0, then go with simpler
AIC(AOD_EL.intx, AOD_EL, EL)

#AIC is lower for the no interaction model
#most appropriate model is no interaction


summary(EL)

```



## Inactivity
### Summary Statistics
```{r ss_IA}
ss_IA <- summarySE(behavior_prop,
                          measurevar = "IA",
                          groupvars = c("caste","DayGrouping"))
kable(ss_IA)
```

No occurrences, no analysis.




## Other
### Summary Statistics
```{r ss_OTH}
ss_OTH <- summarySE(behavior_prop,
                          measurevar = "OTH",
                          groupvars = c("caste","DayGrouping"))
kable(ss_OTH)
```

### Generalized Linear Model (BIONOMIAL)
```{r AOD_OTH}
#mixed model to check for colony effect
summary(glmer(OTH ~ caste*DayGrouping + (1|colony), data=behavior_prop, family=binomial(link = "logit")))


#no effect of colony

#linear models
AOD_OTH.intx <- glm(OTH~caste*DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
AOD_OTH <- glm(OTH~caste+DayGrouping, data=behavior_prop, family=binomial(link = "logit"))
OTH <- glm(OTH~caste, data=behavior_prop, family=binomial(link = "logit"))

####
#now lets compare models to see what the best fit is.
anova(AOD_OTH.intx, AOD_OTH,  test="F")

#model with the lowest ACI should be used
#if difference is less than 2.0, then go with simpler
AIC(AOD_OTH.intx, AOD_OTH, OTH)

#AIC is lower for the no interaction model
#most appropriate model is no interaction


summary(OTH)

```








# On Brood Proportion: Caste * DayGrouping
## Load in Data: on brood proportion
```{r loaddata.onbroodprop, include=FALSE}
#read in data
#data set only contains 'no manipulation' queens and workers
#NO FOSTER DATA INCLUDED
behavior_prop_onbrood <- read.csv2("G:/DesktopDocuments/Documents/PHD/Manuscripts/Bee/Bimp_MaternalBrainVolumetric/BehaviorData/BehaviorData_ProportionData_OnbroodOccurences.csv", sep=",", stringsAsFactors = FALSE)
print(str(behavior_prop_onbrood))

#assign data columns as numeric
numb_cols <- c(6:7)
behavior_prop_onbrood[,numb_cols] <- apply(behavior_prop_onbrood[ , numb_cols], 2, function(x) as.numeric(x))

#Assign factors
behavior_prop_onbrood$colony <- factor(behavior_prop_onbrood$colony)
behavior_prop_onbrood$caste <- factor(behavior_prop_onbrood$caste)
behavior_prop_onbrood$DayGrouping <- factor(behavior_prop_onbrood$DayGrouping)

print(str(behavior_prop_onbrood))
```

## On brood proportion analysis
### Summary Statistics
```{r ss_Onbrood}
ss_LF <- summarySE(behavior_prop_onbrood,
                          measurevar = "onbrood",
                          groupvars = c("caste","DayGrouping"))
kable(ss_LF)
```


### Generalized Linear Model (BIONOMIAL)
```{r AOD_Onbrood}
#mixed model to check for colony effect
summary(glmer(onbrood~caste*DayGrouping + (1|colony), data=behavior_prop_onbrood, family=binomial(link = "logit")))

#no effect of colony

#linear models
AOD_onbrood.intx <- glm(onbrood~caste*DayGrouping, data=behavior_prop_onbrood, family=binomial(link = "logit"))
AOD_onbrood <- glm(onbrood~caste+DayGrouping, data=behavior_prop_onbrood, family=binomial(link = "logit"))
onbrood <- glm(onbrood~caste, data=behavior_prop_onbrood, family=binomial(link = "logit"))

####
#now lets compare models to see what the best fit is.
anova(AOD_onbrood.intx, AOD_onbrood,  test="Chisq")

#model with the lowest ACI should be used
#if difference is less than 2.0, then go with simpler
AIC(AOD_onbrood.intx, AOD_onbrood, onbrood)

#AIC for no interaction is lower
#most appropriate model is no interaction


summary(onbrood)

```
