---
title: "Maternal Brain Neuroplasticity: volumetric analyses with queen and worker tx groups together in one ANOVA analysis "
author: "M.A. Hagadorn"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = TRUE,
	background = "#F7F7F7",
	highlight = TRUE)
```


***
## Load in Required Packages

```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(plyr)
library(dplyr)
library(Rmisc)
library(ggplot2)
library(stats)
library(nortest)
library(MASS)
library(here)
library(effsize)
library(lme4)
library(effsize)
library(multcomp)
library(pwr) #power analysis
library(effectsize)
library(rcompanion) #plotNormalDensity
library(car)  #qqplot
library(s20x) ##lavenes test

installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```





## Load in Data
```{r loaddata, include=FALSE}
#read in data
MB_voldata <- read.csv2("../../DataSheets_VolumesinVaryingForm/Bimp_MaternalBrain2018_VolumetricData_proportionData.csv", sep=",", stringsAsFactors = FALSE)

#not knitted
#MB_voldata <- read.csv2("../DataSheets_VolumesinVaryingForm/Bimp_MaternalBrain2018_VolumetricData_proportionData.csv", sep=",", stringsAsFactors = FALSE)

MBdata_all <- MB_voldata[,-c(1, 13:19)]
colnames(MBdata_all)[6] <- "caste"
colnames(MBdata_all)[5] <- "studycode"
print(str(MBdata_all))

##Subset to remove the NE treatment group not needed in this analysis
MBdata <- MBdata_all[MBdata_all$studycode != "NE",]  #only keep samples that do not equal general_TX==NE

#assign data columns as numeric
numb_cols <- c(11:19)
MBdata[,numb_cols] <- apply(MBdata[ , numb_cols], 2, function(x) as.numeric(x))

#add column for naming for manuscript
MBdata <- cbind(MBdata[,1:5], general_tx=rep(NA, length(MBdata$id)), MBdata[,6:19])

MBdata$general_tx <- with(MBdata, ifelse(MBdata$studycode=="RS", "foundress", 
                                          ifelse(MBdata$studycode=="SR", "queenlike",
                                                 ifelse(MBdata$studycode=="NM", "maternalbrain", " "))))

#update the tx names to match the manuscript names
MBdata$tx[MBdata$tx == "RS_W"] <- "foundresslikeW"
MBdata$tx[MBdata$tx == "SR_W"] <- "queenlikeW"
MBdata$tx[MBdata$tx == "NM_W"] <- "maternallikeW"
MBdata$tx[MBdata$tx == "RS_Q"] <- "foundresslikeQ"
MBdata$tx[MBdata$tx == "SR_Q"] <- "queenlikeQ"
MBdata$tx[MBdata$tx == "NM_Q"] <- "maternallikeQ"




#Assign tx, reproductive status and colony as factors
MBdata$tx <- factor(MBdata$tx, levels=c("foundresslikeW","maternallikeW", "queenlikeW", "foundresslikeQ",  "maternallikeQ", "queenlikeQ"))
#MBdata$general_tx <- factor(MBdata$general_tx, levels=c("RS", "SR", "NM"))
MBdata$general_tx <- factor(MBdata$general_tx, levels=c("foundress", "maternalbrain", "queenlike"))
MBdata$caste <- factor(MBdata$caste, levels = c("queen","worker"))
MBdata$colony <- factor(MBdata$colony)
print(str(MBdata))

write.csv(MBdata, "MBvolumetricData_analysisready.csv", row.names = FALSE)

```

## Sample Sizes and Treatment Group Definitions
The following includes samples sizes and definitions for each TX group

```{r TXdef, echo=FALSE}
#library(knitr)
caste <- c(rep("queen", 3), rep("worker", 3))

TX <- c("Foundress-like", "Maternal-like", "Queen-like", "Foundress-like", "Maternal-like", "Queen-like")

Factor_Code <- c("RS_Q", "NM_Q", "SR_Q", "RS_W", "NM_W", "SR_W")

Sample_Size <- c( "8", "8","7","8", "8", "8")

Definition <- c("Foundress-like: - Reproduction/- Brood care (no experience)", "Maternal-like: + Reproduction/+ Brood care (early colony)", "Queen-like: + Reproduction/- Brood care (late colony)", "Foundress-like: - Reproduction/- Brood care (no experience)", "Maternal-like: + Reproduction/+ Brood care (early colony)", "Queen-like: + Reproduction/- Brood care (late colony)")

studyinformation <- cbind(caste, TX, Factor_Code, Sample_Size, Definition)
colnms <- c("Caste", "Treatment Group", "Analysis Code", "Sample Size", "Definition")
kable(studyinformation, col.names = colnms, align=c('l', 'c','c','c', 'l'), caption = "Study specific information, including treatment group definitions and samples sizes.")
```






For these analyses, we want to make comparisons across *all* treatment groups, regardless of what caste the samples were taken from. As such, we have coded NE, RS, SR, and NM with *_Q* or *_W* for this analysis. We will do a one-way analysis of variance with the fixed factor as **tx**. Prior to doing those analyses, we will investigate the effects of colony as a random factor to make sure it doesn't need to be included in the model.
```{r functions for analyses}

#FUNCTION: Calculate percent increase between means
p.inc <- function(mean1,mean2){
 return((mean1-mean2)/(mean2) * 100)
}

```



***
# One Way ANOVA-6 levels
Prior to conducting these analyses, we assessed a mixed effects model including colony of origin as a random factor. Assessment of those models indicated that the effect of colony was close to zero. As such, in subsequent models it was removed. So, we moved forward with a 2-way ANOVA assessing caste and treatment group with interactions.

## Lip
### Summary Statistics
```{r ss_lips, echo=FALSE}
ss_lip <- summarySE(MBdata,
                          measurevar = "lip_CB",
                          groupvars = c("tx"))
kable(ss_lip)
```

### Box Plot and 2-way ANOVA
```{r bp_lip, echo=FALSE, fig.height=5.5, fig.width=6.5}
legendsize = 7
x_location <- c(.67, .91, 1.17, 1.68, 1.91, 2.17)
#lip specific
sig.code_lip <- c("ab","ab","b","ab","ab","a")
height_lip <- c(0.06856684  +0.0028,  0.07036425 +.0046,0.07474857 +0.0018, 0.06736246+.0045,  0.06823144 +.0043, 0.06398787 + 0.0024)


#Boxplot
ggplot(MBdata, aes(x=caste, y=lip_CB, fill=general_tx)) +
  geom_boxplot() +
  scale_fill_manual(values = c( "gray", "cyan3","darkgoldenrod1"), labels = c("Foundress (-/-)", "Maternal Brain (+/+)","Queenlike (+/-)")) +
  ggtitle("Lips") +
  annotate("text", x=x_location, label= sig.code_lip, y = height_lip, vjust = -0.5) +
  scale_y_continuous(name = "Volume Relative to Central Brain", limits = c(0.05,.08)) +
  scale_x_discrete(name = "Caste") +
  theme(text = element_text(color="black",size = 15),
        axis.title = element_text(color="black"),
        axis.text.x = element_text(color="black", size = 15, margin = margin(t = 5, r =0, b = 20, l = 0)),
        axis.text.y = element_text(color="black", size = 15, margin = margin(t = 0, r =5, b = 0, l = 20)),
        axis.ticks.length = unit (.5,"cm"),
        legend.title = element_blank(),
        legend.position = c(.89,.95),
        legend.margin = margin(l=.6, r=.6, t=.5, b=.5),
        legend.spacing.y = unit(0, "pt"),
        legend.spacing.x = unit(0.3, "pt"),
        legend.background = element_rect(fill="white"), #, size=0.3, linetype="solid", colour ="black"),
        legend.key = element_rect(fill="white"),
        legend.text = element_text(size=legendsize),
        legend.key.height = unit(17, "pt"),
        legend.key.width = unit(.4, "cm"),
        panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(size = 1, linetype = "solid",colour = "black"),
        plot.title = element_text(hjust=0.5))
```




```{r lm_lips}
#Negligable effect of colony, random factor removed

lm_lip <- lm(lip_CB ~ tx, data=MBdata)
summary(lm_lip)

#do some post-hocs to see specifically where the differences are
#multiple comparisons
lip_ph <- glht(lm_lip, mcp(tx="Tukey"))
print(summary(lip_ph))
##Get letters for multiple comparisons
cld(lip_ph, decreasing=FALSE)

```




### Model Verifying Assumptions
```{r assumptions_lip, echo=FALSE}

res.lmlip <- residuals(lm_lip)
par(mfrow=c(1,2))
plotNormalDensity(res.lmlip)
qqPlot(MBdata$lip_CB, "norm")

#Anderson Darling test for normality
#Ho: The residuals are normally distributed
#Ha: The residuals are not normally distributed
#aka if the p-value is <0.05 then you can conclude that the data are not normally distributed
print(ad.test(res.lmlip))

#Homogeneity of variance
#Ho: The distributions are the same.
#Ha: The distributions are not the same.
#aka if the p-value is <0.05 then you can conclude a lack of homogenetity of variance.
liplevene <- levene.test(lm_lip)
liplevene$p.value
```

Normality: residuals are normally distributed.
<br />
Variance: variance are NOT significantly different.

### Percent Increase for significant differences
```{r pinc_lip}
#Q_QL vs W_QL

#mean Q_QL
ss_lip$lip_CB[6]

#mean W_QL
ss_lip$lip_CB[3]

p.inc(ss_lip$lip_CB[6], ss_lip$lip_CB[3])
```




***
## Collar
### Summary Statistics
```{r ss_cols, echo=FALSE}
ss_col <- summarySE(MBdata,
                          measurevar = "col_CB",
                          groupvars = c("tx"))
kable(ss_col)
```

### Box Plot and 2-way ANOVA
```{r bp_col, echo=FALSE, fig.height=5.5, fig.width=6.5}
sig.code_col <- c("cd","bcd","d","a","ac","ab")
x_location <- c(.66, .92, 1.17, 1.67, 1.91, 2.18)
height_col <- c( 0.08430262 +0.0044, 0.08361708+.0015, 0.08956015+0.00183, 0.07311177+.0022, 0.07684370 +.0033, 0.07585705 + 0.0015)


#Boxplot
ggplot(MBdata, aes(x=caste, y=col_CB, fill=general_tx)) +
  geom_boxplot() +
  scale_fill_manual(values = c("gray", "cyan3","darkgoldenrod1"), labels = c("Foundress (-/-)", "Maternal Brain (+/+)","Queenlike (+/-)")) +
  ggtitle("Collars") +
  annotate("text", x=x_location, label= sig.code_col, y = height_col, vjust = -0.5) +
  scale_y_continuous(name = "Volume Relative to Central Brain", limits = c(0.06,.1)) +
  scale_x_discrete(name = "Caste") +
  theme(text = element_text(color="black",size = 15),
        axis.title = element_text(color="black"),
        axis.text.x = element_text(color="black", size = 15, margin = margin(t = 5, r =0, b = 20, l = 0)),
        axis.text.y = element_text(color="black", size = 15, margin = margin(t = 0, r =5, b = 0, l = 20)),
        axis.ticks.length = unit (.5,"cm"),
        legend.title = element_blank(),
        legend.position = c(.89,.95),
        legend.margin = margin(l=.6, r=.6, t=.5, b=.5),
        legend.spacing.y = unit(0, "pt"),
        legend.spacing.x = unit(0.3, "pt"),
        legend.background = element_rect(fill="white"), #, size=0.3, linetype="solid", colour ="black"),
        legend.key = element_rect(fill="white"),
        legend.text = element_text(size=legendsize),
        legend.key.height = unit(17, "pt"),
        legend.key.width = unit(.4, "cm"),
        panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(size = 1, linetype = "solid",colour = "black"),
        plot.title = element_text(hjust=0.5))
```




```{r lm_cols}
#Negligable effect of colony, random factor removed

lm_col <- lm(col_CB ~ tx, data=MBdata)
summary(lm_col)

#do some post-hocs to see specifically where the differences are
#multiple comparisons
col_ph <- glht(lm_col, mcp(tx="Tukey"))
print(summary(col_ph))
##Get letters for multiple comparisons
cld(col_ph, decreasing=FALSE)

```



### Model Verifying Assumptions
```{r assumptions_col, echo=FALSE}
res.lmcol <- residuals(lm_col)
par(mfrow=c(1,2))
plotNormalDensity(res.lmcol)
qqPlot(MBdata$col_CB, "norm")

#Anderson Darling test for normality: testing residuals
print(ad.test(res.lmcol))

#Homogeneity of variance
collevene <- levene.test(lm_col)
collevene$p.value
```

Normality: residuals are normally distributed.
<br />
Variance: variance are NOT significantly different.

### Percent Increase for significant differences
```{r pinc_col}
#Q_QL vs W_QL

#sig dif one:
#mean Q_F
ss_col$col_CB[4]
#mean W_F
ss_col$col_CB[1]
p.inc(ss_col$col_CB[4], ss_col$col_CB[1])

#sig dif two:
#mean Q_F
ss_col$col_CB[4]
#mean W_QL
ss_col$col_CB[3]
p.inc(ss_col$col_CB[4], ss_col$col_CB[3])


#sig dif 3:
#mean Q_QL
ss_col$col_CB[6]
#mean W_F
ss_col$col_CB[1]
p.inc(ss_col$col_CB[6], ss_col$col_CB[1])

#sig dif four:
#mean Q_QL
ss_col$col_CB[6]
#mean W_ML
ss_col$col_CB[2]
p.inc(ss_col$col_CB[6], ss_col$col_CB[2])


#sig dif five:
#mean Q_QL
ss_col$col_CB[6]
#mean W_QL
ss_col$col_CB[3]
p.inc(ss_col$col_CB[6], ss_col$col_CB[3])

#sig dif six:
#mean Q_ML
ss_col$col_CB[5]
#mean W_F
ss_col$col_CB[1]
p.inc(ss_col$col_CB[5], ss_col$col_CB[1])
```





***
## Calyces
### Summary Statistics
```{r ss_calycess, echo=FALSE}
ss_calyces <- summarySE(MBdata,
                          measurevar = "calyces_CB",
                          groupvars = c("tx"))
kable(ss_calyces)
```

### Box Plot and 2-way ANOVA
```{r bp_calyces, echo=FALSE, fig.height=5.5, fig.width=6.5}
x_location <- c(.66, .91, 1.16, 1.67, 1.91, 2.17)
sig.code_cal <- c("ab","ab","b", "a","a","a")
height_cal <- c(0.1528695 +0.006, 0.1539813 +.0055, 0.1643087 + 0.0035, 0.1404742 +.008, 0.1450751 +.009, 0.1398449 + 0.0055)

#Boxplot
ggplot(MBdata, aes(x=caste, y=calyces_CB, fill=general_tx)) +
  geom_boxplot() +
  scale_fill_manual(values = c("gray", "cyan3","darkgoldenrod1"), labels = c("Foundress (-/-)", "Maternal Brain (+/+)","Queenlike (+/-)")) +
  ggtitle("Calyces") +
  annotate("text", x=x_location, label= sig.code_cal, y = height_cal, vjust = -0.5) +
  scale_y_continuous(name = "Volume Relative to Central Brain", limits = c(0.11,.18)) +
  scale_x_discrete(name = "Caste") +
  theme(text = element_text(color="black",size = 15),
        axis.title = element_text(color="black"),
        axis.text.x = element_text(color="black", size = 15, margin = margin(t = 5, r =0, b = 20, l = 0)),
        axis.text.y = element_text(color="black", size = 15, margin = margin(t = 0, r =5, b = 0, l = 20)),
        axis.ticks.length = unit (.5,"cm"),
        legend.title = element_blank(),
        legend.position = c(.95,.95),
        legend.margin = margin(l=.6, r=.6, t=.5, b=.5),
        legend.spacing.y = unit(0, "pt"),
        legend.spacing.x = unit(0.3, "pt"),
        legend.background = element_rect(fill="white"), #, size=0.3, linetype="solid", colour ="black"),
        legend.key = element_rect(fill="white"),
        legend.text = element_text(size=legendsize),
        legend.key.height = unit(17, "pt"),
        legend.key.width = unit(.4, "cm"),
        panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(size = 1, linetype = "solid",colour = "black"),
        plot.title = element_text(hjust=0.5))
```




```{r lm_calyces}
#Negligable effect of colony, random factor removed

lm_cal <- lm(calyces_CB ~ tx, data=MBdata)
summary(lm_cal)


#do some post-hocs to see specifically where the differences are
#multiple comparisons
cal_ph <- glht(lm_cal, mcp(tx="Tukey"))
print(summary(cal_ph))
##Get letters for multiple comparisons
cld(cal_ph, decreasing=FALSE)

```



### Model Verifying Assumptions
```{r assumptions_calyces, echo=FALSE}
res.lmcalyces <- residuals(lm_cal)
par(mfrow=c(1,2))
plotNormalDensity(res.lmcalyces)
qqPlot(MBdata$calyces_CB, "norm")

#need to box cox

#Anderson Darling test for normality: testing residuals
print(ad.test(res.lmcalyces))

#Homogeneity of variance
calyceslevene <- levene.test(lm_cal)
calyceslevene$p.value
```

Normality: residuals are normally distributed.
<br />
Variance: variance are not significantly different.
### Percent Increase for significant differences
```{r pinc_calyces}
#Q_QL vs W_QL
#mean Q_QL
ss_calyces$calyces_CB[6]
#mean W_QL
ss_calyces$calyces_CB[1]
p.inc(ss_calyces$calyces_CB[6], ss_calyces$calyces_CB[1])


#Q_QL vs W_QL
#mean Q_QL
ss_calyces$calyces_CB[6]
#mean W_QL
ss_calyces$calyces_CB[2]
p.inc(ss_calyces$calyces_CB[6], ss_calyces$calyces_CB[2])


#Q_QL vs W_QL
#mean Q_QL
ss_calyces$calyces_CB[6]
#mean W_QL
ss_calyces$calyces_CB[3]
p.inc(ss_calyces$calyces_CB[6], ss_calyces$calyces_CB[3])
```




***
## Kenyon Cells
### Summary Statistics
```{r ss_kcs, echo=FALSE}
ss_kcs <- summarySE(MBdata,
                          measurevar = "kcs_CB",
                          groupvars = c("tx"))
kable(ss_kcs)
```

### Box Plot and 2-way ANOVA
```{r bp_kcs, echo=FALSE, fig.height=5.5, fig.width=6.5}
x_location <- c(.67, .91, 1.17, 1.67, 1.91, 2.16)
sig.code_kcs <- c("ab","a","a", "ab","ab","b")
height_kcs <- c(0.09890870 +0.0012, 0.09471199 +.0059, 0.09366606 + 0.0029, 0.09983231 +.0042, 0.09919472 +.0028, 0.10333983 + 0.0017)



#Boxplot
ggplot(MBdata, aes(x=caste, y=kcs_CB, fill=general_tx)) +
  geom_boxplot() +
  scale_fill_manual(values = c("gray", "cyan3","darkgoldenrod1"), labels = c("Foundress (-/-)", "Maternal Brain (+/+)","Queenlike (+/-)")) +
  ggtitle("Kenyon Cells") +
  annotate("text", x=x_location, label= sig.code_kcs, y = height_kcs, vjust = -0.5) +
  scale_y_continuous(name = "Mean volume relative to Central Brain", limits = c(0.08,.125)) +
  scale_x_discrete(name = "Caste") +
  theme(text = element_text(color="black",size = 15),
        axis.title = element_text(color="black"),
        axis.text.x = element_text(color="black", size = 15, margin = margin(t = 5, r =0, b = 20, l = 0)),
        axis.text.y = element_text(color="black", size = 15, margin = margin(t = 0, r =5, b = 0, l = 20)),
        axis.ticks.length = unit (.5,"cm"),
        legend.title = element_blank(),
        legend.position = c(.89,.95),
        legend.margin = margin(l=.6, r=.6, t=.5, b=.5),
        legend.spacing.y = unit(0, "pt"),
        legend.spacing.x = unit(0.3, "pt"),
        legend.background = element_rect(fill="white"), #, size=0.3, linetype="solid", colour ="black"),
        legend.key = element_rect(fill="white"),
        legend.text = element_text(size=legendsize),
        legend.key.height = unit(17, "pt"),
        legend.key.width = unit(.4, "cm"),
        panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(size = 1, linetype = "solid",colour = "black"),
        plot.title = element_text(hjust=0.5))
```




```{r lm_kcs}
#Negligable effect of colony, random factor removed

lm_kcs <- lm(kcs_CB ~ tx, data=MBdata)
summary(lm_kcs)


#do some post-hocs to see specifikcsly where the differences are
#multiple comparisons
kcs_ph <- glht(lm_kcs, mcp(tx="Tukey"))
print(summary(kcs_ph))
##Get letters for multiple comparisons
cld(kcs_ph, decreasing=FALSE)

```





### Model Verifying Assumptions
```{r assumptions_kcs, echo=FALSE}
res.lmkcs <- residuals(lm_kcs)
par(mfrow=c(1,2))
plotNormalDensity(res.lmkcs)
qqPlot(MBdata$kcs_CB, "norm")

#Anderson Darling test for normality: testing residuals
print(ad.test(res.lmkcs))

#Homogeneity of variance
kcslevene <- levene.test(lm_kcs)
kcslevene$p.value
```

Normality: residuals are normally distributed.
<br />
Variance: variance are NOT significantly different.


### Percent Increase for significant differences
```{r pinc_kcs}
#Q_QL vs W_QL

#mean Q_QL
ss_kcs$kcs_CB[6]
#mean W_QL
ss_kcs$kcs_CB[3]
p.inc(ss_kcs$kcs_CB[6], ss_kcs$kcs_CB[3])

#Q_ML vs W_QL

#mean Q_ML
ss_kcs$kcs_CB[5]
#mean W_QL
ss_kcs$kcs_CB[3]
p.inc(ss_kcs$kcs_CB[5], ss_kcs$kcs_CB[3])


```





***
## Mushroom Body Lobes
### Summary Statistics
```{r ss_mblobes, echo=FALSE}
ss_mblobes <- summarySE(MBdata,
                          measurevar = "mblobes_CB",
                          groupvars = c("tx"))
kable(ss_mblobes)
```

### Box Plot and 2-way ANOVA
```{r bp_mblobes, echo=FALSE, fig.height=5.5, fig.width=6.5}
x_location <- c(.66, .91, 1.16, 1.66, 1.91, 2.16)
sig.code_kcs <- c("a","a","a","a","a","a")
height_kcs <- c(0.11294111 +0.005, 0.11209956 +.0031, 0.11294111 + 0.0019, 0.11001868 +.003, 0.10114805 +.0145, 0.10977559 + 0.002)


#Boxplot
ggplot(MBdata, aes(x=caste, y=mblobes_CB, fill=general_tx)) +
  geom_boxplot() +
  scale_fill_manual(values = c("gray", "cyan3","darkgoldenrod1"), labels = c("Foundress (-/-)", "Maternal Brain (+/+)","Queenlike (+/-)")) +
  ggtitle("Mushroom Body Lobes") +
  annotate("text", x=x_location, label= sig.code_kcs, y = height_kcs, vjust = -0.5) +
  scale_y_continuous(name = "Volume Relative to Central Brain", limits = c(0.09,.13)) +
  scale_x_discrete(name = "Caste") +
  theme(text = element_text(color="black",size = 15),
        axis.title = element_text(color="black"),
        axis.text.x = element_text(color="black", size = 15, margin = margin(t = 5, r =0, b = 20, l = 0)),
        axis.text.y = element_text(color="black", size = 15, margin = margin(t = 0, r =5, b = 0, l = 20)),
        axis.ticks.length = unit (.5,"cm"),
        legend.title = element_blank(),
        legend.position = c(.89,.95),
        legend.margin = margin(l=.6, r=.6, t=.5, b=.5),
        legend.spacing.y = unit(0, "pt"),
        legend.spacing.x = unit(0.3, "pt"),
        legend.background = element_rect(fill="white"), #, size=0.3, linetype="solid", colour ="black"),
        legend.key = element_rect(fill="white"),
        legend.text = element_text(size=legendsize),
        legend.key.height = unit(17, "pt"),
        legend.key.width = unit(.4, "cm"),
        panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(size = 1, linetype = "solid",colour = "black"),
        plot.title = element_text(hjust=0.5))
```




```{r lm_mblobes}
#Negligable effect of colony, random factor removed

lm_mblobes <- lm(mblobes_CB ~ tx, data=MBdata)
summary(lm_mblobes)

#No significant difference; no post-hoc
```



### Model Verifying Assumptions
```{r assumptions_mblobes, echo=FALSE}

res.lmmblobes <- residuals(lm_mblobes)
par(mfrow=c(1,2))
plotNormalDensity(res.lmmblobes)
qqPlot(MBdata$mblobes_CB, "norm")

#Anderson Darling test for normality: testing module residuals
print(ad.test(res.lmmblobes))


#Homogeneity of variance
mblobeslevene <- levene.test(lm_mblobes)
mblobeslevene$p.value
```

Normality: residuals are normally distributed.
<br />
Variance: variance are NOT significantly different.



***
## Neuropil
### Summary Statistics
```{r ss_neuropil, echo=FALSE}
ss_neuropil <- summarySE(MBdata,
                          measurevar = "neuropil_CB",
                          groupvars = c("tx"))
kable(ss_neuropil)
```

### Box Plot and 2-way ANOVA
```{r bp_neuropil, echo=FALSE, fig.height=5.5, fig.width=6.5}
x_location <- c(.66, .91, 1.17, 1.675, 1.91, 2.175)
sig.code_kcs <- c("bc","bc","c", "a","ab","a")
height_kcs <- c(0.2675858 +0.0045, 0.2660809 +.0055, 0.2772498 - 0.0005, 0.2504929 +.004, 0.2554510 +.008,0.2496205 + 0.0025)

#Boxplot
ggplot(MBdata, aes(x=caste, y=neuropil_CB, fill=general_tx)) +
  geom_boxplot() +
   scale_fill_manual(values = c("gray", "cyan3","darkgoldenrod1"), labels = c("Foundress (-/-)", "Maternal Brain (+/+)","Queenlike (+/-)")) +
  ggtitle("Total Neuropil") +
  annotate("text", x=x_location, label= sig.code_kcs, y = height_kcs, vjust = -0.5) +
  scale_y_continuous(name = "Volume Relative to Central Brain", limits = c(0.22,.30)) +
  scale_x_discrete(name = "Caste") +
  theme(text = element_text(color="black",size = 15),
        axis.title = element_text(color="black"),
        axis.text.x = element_text(color="black", size = 15, margin = margin(t = 5, r =0, b = 20, l = 0)),
        axis.text.y = element_text(color="black", size = 15, margin = margin(t = 0, r =5, b = 0, l = 20)),
        axis.ticks.length = unit (.5,"cm"),
        legend.title = element_blank(),
        legend.position = c(.89,.95),
        legend.margin = margin(l=.6, r=.6, t=.5, b=.5),
        legend.spacing.y = unit(0, "pt"),
        legend.spacing.x = unit(0.3, "pt"),
        legend.background = element_rect(fill="white"), #, size=0.3, linetype="solid", colour ="black"),
        legend.key = element_rect(fill="white"),
        legend.text = element_text(size=legendsize),
        legend.key.height = unit(17, "pt"),
        legend.key.width = unit(.4, "cm"),
        panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(size = 1, linetype = "solid",colour = "black"),
        plot.title = element_text(hjust=0.5))
```




```{r lm_neuropil}
#Negligable effect of colony, random factor removed

lm_neuropil <- lm(neuropil_CB ~ tx, data=MBdata)
summary(lm_neuropil)


#do some post-hocs to see specifikcsly where the differences are
#multiple comparisons
neuropil_ph <- glht(lm_neuropil, mcp(tx="Tukey"))
print(summary(neuropil_ph))
##Get letters for multiple comparisons
cld(neuropil_ph, decreasing=FALSE)
```



### Model Verifying Assumptions
```{r assumptions_neuropil, echo=FALSE}

res.lmneuropil <- residuals(lm_neuropil)
par(mfrow=c(1,2))
plotNormalDensity(res.lmneuropil)
qqPlot(MBdata$neuropil_CB, "norm")

#Anderson Darling test for normality: testing module residuals
print(ad.test(res.lmneuropil))

#Homogeneity of variance
neuropillevene <- levene.test(lm_neuropil)
neuropillevene$p.value
```

Normality: residuals are normally distributed.
<br />
Variance: variance are NOT significantly different.

### Percent Increase for significant differences
```{r pinc_neuropil}

#Q_QL vs W_F
#mean Q_QL
ss_neuropil$neuropil_CB[6]
#mean W_F
ss_neuropil$neuropil_CB[1]
p.inc(ss_neuropil$neuropil_CB[6], ss_neuropil$neuropil_CB[1])

#Q_QL vs W_ML
#mean Q_QL
ss_neuropil$neuropil_CB[6]
#mean W_ML
ss_neuropil$neuropil_CB[2]
p.inc(ss_neuropil$neuropil_CB[6], ss_neuropil$neuropil_CB[2])

#Q_QL vs W_QL
#mean Q_QL
ss_neuropil$neuropil_CB[6]
#mean W_QL
ss_neuropil$neuropil_CB[3]
p.inc(ss_neuropil$neuropil_CB[6], ss_neuropil$neuropil_CB[3])



#Q_F vs W_F
#mean Q_F
ss_neuropil$neuropil_CB[4]
#mean W_F
ss_neuropil$neuropil_CB[1]
p.inc(ss_neuropil$neuropil_CB[4], ss_neuropil$neuropil_CB[1])

#Q_F vs W_QL
#mean Q_F
ss_neuropil$neuropil_CB[4]
#mean W_QL
ss_neuropil$neuropil_CB[3]
p.inc(ss_neuropil$neuropil_CB[4], ss_neuropil$neuropil_CB[3])

#Q_ML vs W_F
#mean Q_ML
ss_neuropil$neuropil_CB[5]
#mean W_F
ss_neuropil$neuropil_CB[1]
p.inc(ss_neuropil$neuropil_CB[5], ss_neuropil$neuropil_CB[1])

#Q_ML vs W_QL
#mean Q_ML
ss_neuropil$neuropil_CB[5]
#mean W_QL
ss_neuropil$neuropil_CB[3]
p.inc(ss_neuropil$neuropil_CB[5], ss_neuropil$neuropil_CB[3])

```


***
## Neuropil:Kenyon cell
### Summary Statistics
```{r ss_nk, echo=FALSE}
ss_nk <- summarySE(MBdata,
                          measurevar = "nk",
                          groupvars = c("tx"))
kable(ss_nk)
```

### Box Plot and 2-way ANOVA
```{r bp_nk, echo=FALSE, fig.height=5.5, fig.width=6.5}
x_location <- c(.67, .91, 1.17, 1.67, 1.91, 2.17)
sig.code_nk <- c("bc","cd","d","ab","ac","a")
height_nk <- c(2.707984 +0.1, 2.819165 +.13, 2.965873 + 0.035, 2.515303 +.065, 2.582801 +.12,2.419149 + 0.018)



#Boxplot
ggplot(MBdata, aes(x=caste, y=nk, fill=general_tx)) +
  geom_boxplot() +
   scale_fill_manual(values = c("gray", "cyan3","darkgoldenrod1"), labels = c("Foundress (-/-)", "Maternal Brain (+/+)","Queenlike (+/-)")) +
  ggtitle("Neuropil:Kenyon cell") +
  annotate("text", x=x_location, label= sig.code_nk, y = height_nk, vjust = -0.5) +
  scale_y_continuous(name = "Volume Relative to Central Brain", limits = c(2.0,3.5)) +
  scale_x_discrete(name = "Caste") +
  theme(text = element_text(color="black",size = 15),
        axis.title = element_text(color="black"),
        axis.text.x = element_text(color="black", size = 15, margin = margin(t = 5, r =0, b = 20, l = 0)),
        axis.text.y = element_text(color="black", size = 15, margin = margin(t = 0, r =5, b = 0, l = 20)),
        axis.ticks.length = unit (.5,"cm"),
        legend.title = element_blank(),
        legend.position = c(.89,.95),
        legend.margin = margin(l=.6, r=.6, t=.5, b=.5),
        legend.spacing.y = unit(0, "pt"),
        legend.spacing.x = unit(0.3, "pt"),
        legend.background = element_rect(fill="white"), #, size=0.3, linetype="solid", colour ="black"),
        legend.key = element_rect(fill="white"),
        legend.text = element_text(size=legendsize),
        legend.key.height = unit(17, "pt"),
        legend.key.width = unit(.4, "cm"),
        panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(size = 1, linetype = "solid",colour = "black"),
        plot.title = element_text(hjust=0.5))
```




```{r lm_nk}
#no effect of colony

lm_nk <- lm(nk ~ tx, data=MBdata)
summary(lm_nk)


#do some post-hocs to see specifikcsly where the differences are
#multiple comparisons
nk_ph <- glht(lm_nk, mcp(tx="Tukey"))
print(summary(nk_ph))
##Get letters for multiple comparisons
cld(nk_ph, decreasing=FALSE)

```



### Model Verifying Assumptions
```{r assumptions_nk, echo=FALSE}

res.lmnk <- residuals(lm_nk)
par(mfrow=c(1,2))
plotNormalDensity(res.lmnk)
qqPlot(MBdata$nk, "norm") #close

#Anderson Darling test for normality: testing module residuals
print(ad.test(res.lmnk)) # close

#Homogeneity of variance
nklevene <- levene.test(lm_nk)
nklevene$p.value
```

Normality: residuals are normally distributed.
<br />
Variance: variance are NOT significantly different.


### Percent Increase for significant differences
```{r pinc_nk}

#Q_QL vs Q_F
#mean Q_QL
ss_nk$nk[6]
#mean Q_F
ss_nk$nk[4]
p.inc(ss_nk$nk[6], ss_nk$nk[4])

#Q_QL vs W_F
#mean Q_QL
ss_nk$nk[6]
#mean W_F
ss_nk$nk[1]
p.inc(ss_nk$nk[6], ss_nk$nk[1])

#Q_QL vs W_ML
#mean Q_QL
ss_nk$nk[6]
#mean W_ML
ss_nk$nk[2]
p.inc(ss_nk$nk[6], ss_nk$nk[2])


#Q_QL vs W_QL
#mean Q_QL
ss_nk$nk[6]
#mean W_QL
ss_nk$nk[3]
p.inc(ss_nk$nk[6], ss_nk$nk[3])


#Q_F vs W_QL
#mean Q_F
ss_nk$nk[4]
#mean W_QL
ss_nk$nk[3]
p.inc(ss_nk$nk[4], ss_nk$nk[3])

#Q_ML vs W_F
#mean Q_ML
ss_nk$nk[5]
#mean W_F
ss_nk$nk[1]
p.inc(ss_nk$nk[5], ss_nk$nk[1])

#Q_ML vs W_QL
#mean Q_ML
ss_nk$nk[5]
#mean W_QL
ss_nk$nk[3]
p.inc(ss_nk$nk[5], ss_nk$nk[3])

```




***
## Antennal Lobes
### Summary Statistics
```{r ss_als, echo=FALSE}
ss_als <- summarySE(MBdata,
                          measurevar = "als_CB",
                          groupvars = c("tx"))
kable(ss_als)
```

### Box Plot and 2-way ANOVA
```{r bp_als, echo=FALSE, fig.height=5.5, fig.width=6.5}
x_location <- c(.66, .9, 1.16, 1.67, 1.91, 2.17)
sig.code_als <- c("ab","ab","b", "a","a","a")
height_als <- c(0.05689339 +0.0013, 0.05677000 +.0016, 0.06007301 + 0.00135, 0.05337774 +.0025, 0.05261469 +.0006, 0.05419108 + 0.0015)


#Boxplot
ggplot(MBdata, aes(x=caste, y=als_CB, fill=general_tx)) +
  geom_boxplot() +
  scale_fill_manual(values = c("gray", "cyan3","darkgoldenrod1"), labels = c("Foundress (-/-)", "Maternal Brain (+/+)","Queenlike (+/-)")) +
  ggtitle("Antennal Lobes") +
  annotate("text", x=x_location, label= sig.code_kcs, y = height_kcs, vjust = -0.5) +
  scale_y_continuous(name = "Volume Relative to Central Brain", limits = c(0.04,.07)) +
  scale_x_discrete(name = "Caste") +
  theme(text = element_text(color="black",size = 15),
        axis.title = element_text(color="black"),
        axis.text.x = element_text(color="black", size = 15, margin = margin(t = 5, r =0, b = 20, l = 0)),
        axis.text.y = element_text(color="black", size = 15, margin = margin(t = 0, r =5, b = 0, l = 20)),
        axis.ticks.length = unit (.5,"cm"),
        legend.title = element_blank(),
        legend.position = c(.89,.95),
        legend.margin = margin(l=.6, r=.6, t=.5, b=.5),
        legend.spacing.y = unit(0, "pt"),
        legend.spacing.x = unit(0.3, "pt"),
        legend.background = element_rect(fill="white"), #, size=0.3, linetype="solid", colour ="black"),
        legend.key = element_rect(fill="white"),
        legend.text = element_text(size=legendsize),
        legend.key.height = unit(17, "pt"),
        legend.key.width = unit(.4, "cm"),
        panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(size = 1, linetype = "solid",colour = "black"),
        plot.title = element_text(hjust=0.5))
```




```{r lm_als}
#no effect of colony

lm_als <- lm(als_CB ~ tx, data=MBdata)
summary(lm_als)


#do some post-hocs to see specifikcsly where the differences are
#multiple comparisons
als_ph <- glht(lm_als, mcp(tx="Tukey"))
print(summary(als_ph))
##Get letters for multiple comparisons
cld(als_ph, decreasing=FALSE)

```



### Model Verifying Assumptions
```{r assumptions_als, echo=FALSE}

res.lmals <- residuals(lm_als)
par(mfrow=c(1,2))
plotNormalDensity(res.lmals)
qqPlot(MBdata$als_CB, "norm")

#Anderson Darling test for normality: testing module residuals
print(ad.test(res.lmals))

#Homogeneity of variance
alslevene <- levene.test(lm_als)
alslevene$p.value
```

Normality: residuals are normally distributed.
<br />
Variance: variance are NOT significantly different.

```{r pinc_als}

#Q_QL vs Q_F
#mean Q_QL
ss_als$als_CB[6]
#mean W_F
ss_als$als_CB[1]
p.inc(ss_als$als_CB[6], ss_als$als_CB[1])

#Q_QL vs W_ML
#mean Q_QL
ss_als$als_CB[6]
#mean W_ML
ss_als$als_CB[2]
p.inc(ss_als$als_CB[6], ss_als$als_CB[2])

#Q_QL vs W_QL
#mean Q_QL
ss_als$als_CB[6]
#mean W_QL
ss_als$als_CB[3]
p.inc(ss_als$als_CB[6], ss_als$als_CB[3])




```

