---
title: "B. impatiens Maternal Brain: Gene Set Enrichment Analyses"
author: "M.A. Hagadorn"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#school="D:
#setwd("D:/DesktopDocuments/Documents/PHD/Studies/Bombus_impatiens/Bimpatiens_MaternalBrainGeneExpression/RNASeqAnalysis/BimpMB_RNAanalysis/GSEA")

#home="H:
setwd("H:/DesktopDocuments/Documents/PHD/Studies/Bombus_impatiens/Bimpatiens_MaternalBrainGeneExpression/RNASeqAnalysis/BimpMB_RNAanalysis/GSEA")
```

```{r loadpackages, warning=FALSE, echo=FALSE, message=FALSE, include=TRUE}
#Library Necessary Packages
library(tidyverse)
library(lattice)
library(lintr)
library(AnnotationDbi)
library(topGO)
library(rrvgo)
library(org.Dm.eg.db)
library(writexl)
library(openxlsx)
library(edgeR)
library(dplyr)
library(tidyr)
library(tibble)
options(stringsAsFactors = FALSE)
```


```{r sessioninformation, echo=FALSE}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```

# Data Preperation for GO Enrichment Analysis
In 2022, I used the following resource to work through the GO analysis:
https://bioconductor.org/packages/release/bioc/vignettes/topGO/inst/doc/topGO.pdf. I am building this analysis using the same resource and work through.

## Load in Gene Data
```{r LoadinGeneData, echo=FALSE}
#pulling from the document of DEG results. This contains all filtered genes and the genes are the same across the DEG analyses, just with different pairwise results.
filteredGeneList <- read.csv2(file = "../DEG/DEGFullResults/DEGananalysisResults_FLWMLW.csv", header=TRUE, sep=",")
rownames(filteredGeneList) <- filteredGeneList$X
filteredGeneList <- filteredGeneList[,-1]

```


## Data preparation
### Read in Mapping File
```{r readinannotation}
#read in the GO annotation file for BIMP
#GAF file was downloaded from NCBI 2/13/2024 (BIMP 2.2) and modified to include the go ids, terms, and definitions
#https://www.ncbi.nlm.nih.gov/datasets/genome/?taxon=132113    Selected BIMP_2.2
gene2mapping <- readMappings("Gene2GoMap_unique.txt", sep = " ", IDsep=",")
```


### 'Universe' Gene List
```{r allgeneslist}
#List of all genes
geneUniverse <- rownames(filteredGeneList) #give list of all gene names
#This was the filtered Set used in the DEG analysis

length(geneUniverse)
```

### Build GoData object
Using the following resource as a guide: https://ucdavis-bioinformatics-training.github.io/2018-June-RNA-Seq-Workshop/friday/enrichment.html
```{r geneSel.fun}
#function that returns genes < 0.05 adj pvalue
topDiffGenes <- function(allScore){
  return(allScore < 0.05)
}
```


# GSEA by Pairwise Comparisions 

## Across Castes
### FLW v FLQ
#### 'Gene' list for DEGs
```{r degsgeneslistFLWFLQ}
DEG.data.FLWFLQ <- read.csv2(file = "../DEG/DEGFullResults/DEGananalysisResults_FLWFLQ.csv", header=TRUE, sep=",")
rownames(DEG.data.FLWFLQ ) <- DEG.data.FLWFLQ$X
DEG.data.FLWFLQ  <- DEG.data.FLWFLQ [,-1]
```


I discovered that having NAs in the padj is a problem when running the KS test via: https://www.biostars.org/p/91474/. Removed the NAs to move forward.
```{r removeNAs_FLWFLQ}
table(is.na(DEG.data.FLWFLQ$padj))
NAlist_FLWFLQ <- which(is.na(DEG.data.FLWFLQ$padj))
rownames(DEG.data.FLWFLQ)[NAlist_FLWFLQ]

#gene list for topgo input
DEG.genelist.FLWFLQ  <- as.numeric(DEG.data.FLWFLQ$padj)
names(DEG.genelist.FLWFLQ) <- geneUniverse


##NEED TO REMOVE NAs IN PADJ TO PROCEED
#VERIFY SAME POSITIONS ACROSS ALL PAIRWISE
FLWFLQ_napos <- which(is.na(DEG.genelist.FLWFLQ), arr.ind = TRUE)

#List of NA genes removed
FLWFLQ_napos


#Create a new data frame to use going forward W/O those padj NAs
#This is needed for the KS tests
DEG.genelist.FLWFLQ <- DEG.genelist.FLWFLQ[-FLWFLQ_napos]
length(DEG.genelist.FLWFLQ)

#get an idea of the top diff genes
table(topDiffGenes(DEG.genelist.FLWFLQ))
table(DEG.genelist.FLWFLQ < 0.05)

```


#### BP
```{r BP_DEG}
DEgenes.FLWFLQ.GO.bp <- new("topGOdata",
                description = paste("DEG Go", "BP", "FLWFLQ", sep = "_"), 
                ontology = "BP",
                allGenes = DEG.genelist.FLWFLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLWFLQ.bp_result_fisher <- runTest(DEgenes.FLWFLQ.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.FLWFLQ.bp_result_fisher

DEG_BP.FLWFLQ.table <- GenTable(DEgenes.FLWFLQ.GO.bp, 
                                raw.p.value = DEG.FLWFLQ.bp_result_fisher, 
                                topNodes = 50, 
                                numChar = 120)

write.csv(DEG_BP.FLWFLQ.table, "Results/ByOntology/FLWFLQ_BPTerms.csv")

```

#### MF
```{r mf_DEG}
DEgenes.FLWFLQ.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "FLWFLQ", sep = "_"), 
                ontology = "MF",
                allGenes = DEG.genelist.FLWFLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLWFLQ.mf_result_fisher <- runTest(DEgenes.FLWFLQ.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.FLWFLQ.mf_result_fisher

DEG_MF.FLWFLQ.table <- GenTable(DEgenes.FLWFLQ.GO.mf, 
                                raw.p.value = DEG.FLWFLQ.mf_result_fisher, 
                                topNodes = 50, numChar = 120)

write.csv(DEG_MF.FLWFLQ.table, "Results/ByOntology/FLWFLQ_MFTerms.csv")
```



#### CC
```{r cc_DEG}
DEgenes.FLWFLQ.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "FLWFLQ", sep = "_"), 
                ontology = "CC",
                allGenes = DEG.genelist.FLWFLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLWFLQ.cc_result_fisher <- runTest(DEgenes.FLWFLQ.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.FLWFLQ.cc_result_fisher

DEG_CC.FLWFLQ.table <- GenTable(DEgenes.FLWFLQ.GO.cc, 
                                raw.p.value = DEG.FLWFLQ.cc_result_fisher, 
                                topNodes = 50,
                                numChar = 120)

write.csv(DEG_CC.FLWFLQ.table, "Results/ByOntology/FLWFLQ_CCTerms.csv")
```


```{r largetable.DEG_Node=1}
#ALL
DEG_BP.FLWFLQ.table <- cbind(DEG_BP.FLWFLQ.table, ontology=rep("BP", length(DEG_BP.FLWFLQ.table$GO.ID)))
DEG_MF.FLWFLQ.table <- cbind(DEG_MF.FLWFLQ.table, ontology=rep("MF", length(DEG_MF.FLWFLQ.table$GO.ID)))
DEG_CC.FLWFLQ.table <- cbind(DEG_CC.FLWFLQ.table, ontology=rep("CC", length(DEG_CC.FLWFLQ.table$GO.ID)))

DEG_GOallontology.FLWFLQ <-rbind(DEG_BP.FLWFLQ.table, DEG_MF.FLWFLQ.table, DEG_CC.FLWFLQ.table)
write.csv(DEG_GOallontology.FLWFLQ, "Results/AllGosOntologyCombined/DEGs_GOallontology_FLWFLQ.csv")


#Less than .01
DEG_BP.FLWFLQ.table.01 <- DEG_BP.FLWFLQ.table[DEG_BP.FLWFLQ.table$raw.p.value < 0.01,]
DEG_MF.FLWFLQ.table.01 <- DEG_MF.FLWFLQ.table[DEG_MF.FLWFLQ.table$raw.p.value < 0.01,]
DEG_CC.FLWFLQ.table.01 <- DEG_CC.FLWFLQ.table[DEG_CC.FLWFLQ.table$raw.p.value < 0.01,]

DEG_GOallontology.FLWFLQ.01 <-rbind(DEG_BP.FLWFLQ.table.01, DEG_MF.FLWFLQ.table.01, DEG_CC.FLWFLQ.table.01)
write.csv(DEG_GOallontology.FLWFLQ.01, "Results/DEGs_GOallontology_FLWFLQ.01.csv")
```


#### Pull out genes for Significant Go terms




### MLW v MLQ
#### 'Gene' list for DEGs
```{r degsgeneslistMLWMLQ}
DEG.data.MLWMLQ <- read.csv2(file = "../DEG/DEGFullResults/DEGananalysisResults_MLWMLQ.csv", header=TRUE, sep=",")
rownames(DEG.data.MLWMLQ) <- DEG.data.MLWMLQ$X
DEG.data.MLWMLQ <- DEG.data.MLWMLQ[,-1]
```


I discovered that having NAs in the padj is a problem when running the KS test via: https://www.biostars.org/p/91474/. Removed the NAs to move forward.
```{r removeNAs_MLWMLQ}
table(is.na(DEG.data.MLWMLQ$padj))
NAlist_MLWMLQ <- which(is.na(DEG.data.MLWMLQ$padj))
rownames(DEG.data.MLWMLQ)[NAlist_MLWMLQ]

#gene list for topgo input
DEG.genelist.MLWMLQ  <- as.numeric(DEG.data.MLWMLQ$padj)
names(DEG.genelist.MLWMLQ) <- geneUniverse


##NEED TO REMOVE NAs IN PADJ TO PROCEED
#VERIFY SAME POSITIONS ACROSS ALL PAIRWISE
MLWMLQ_napos <- which(is.na(DEG.genelist.MLWMLQ), arr.ind = TRUE)



#Create a new data frame to use going forward W/O those padj NAs
#This is needed for the KS tests
DEG.genelist.MLWMLQ <- DEG.genelist.MLWMLQ[-MLWMLQ_napos]

#get an idea of the top diff genes
table(topDiffGenes(DEG.genelist.MLWMLQ))
table(DEG.genelist.MLWMLQ < 0.05)
```


#### BP
```{r BP_DEG_MLWMLQ}
DEgenes.MLWMLQ.GO.bp <- new("topGOdata",
                description = paste("DEG Go", "BP", "MLWMLQ", sep = "_"),
                ontology = "BP",
                allGenes = DEG.genelist.MLWMLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.MLWMLQ.bp_result_fisher <- runTest(DEgenes.MLWMLQ.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.MLWMLQ.bp_result_fisher

DEG_BP.MLWMLQ.table <- GenTable(DEgenes.MLWMLQ.GO.bp, 
                                raw.p.value = DEG.MLWMLQ.bp_result_fisher, 
                                topNodes = 50, 
                                numChar = 120)

write.csv(DEG_BP.MLWMLQ.table, "Results/ByOntology/MLWMLQ_BPTerms.csv")
```

#### MF
```{r mf_DEG_MLWMLQ}
DEgenes.MLWMLQ.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "MLWMLQ", sep = "_"),
                ontology = "MF",
                allGenes = DEG.genelist.MLWMLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.MLWMLQ.mf_result_fisher <- runTest(DEgenes.MLWMLQ.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.MLWMLQ.mf_result_fisher

DEG_MF.MLWMLQ.table <- GenTable(DEgenes.MLWMLQ.GO.mf, 
                                raw.p.value = DEG.MLWMLQ.mf_result_fisher, 
                                topNodes = 50, 
                                numChar = 120)

write.csv(DEG_MF.MLWMLQ.table, "Results/ByOntology/MLWMLQ_MFTerms.csv")
```



#### CC
```{r cc_DEG_MLWMLQ}
DEgenes.MLWMLQ.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "MLWMLQ", sep = "_"),
                ontology = "CC",
                allGenes = DEG.genelist.MLWMLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.MLWMLQ.cc_result_fisher <- runTest(DEgenes.MLWMLQ.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.MLWMLQ.cc_result_fisher

DEG_CC.MLWMLQ.table <- GenTable(DEgenes.MLWMLQ.GO.cc, 
                                raw.p.value = DEG.MLWMLQ.cc_result_fisher, 
                                topNodes = 50, 
                                numChar = 120)

write.csv(DEG_CC.MLWMLQ.table, "Results/ByOntology/MLWMLQ_CCTerms.csv")
```


```{r MLWMLQlargetable.DEG_Node=1}
#ALL
DEG_BP.MLWMLQ.table <- cbind(DEG_BP.MLWMLQ.table, ontology=rep("BP", length(DEG_BP.MLWMLQ.table$GO.ID)))
DEG_MF.MLWMLQ.table <- cbind(DEG_MF.MLWMLQ.table, ontology=rep("MF", length(DEG_MF.MLWMLQ.table$GO.ID)))
DEG_CC.MLWMLQ.table <- cbind(DEG_CC.MLWMLQ.table, ontology=rep("CC", length(DEG_CC.MLWMLQ.table$GO.ID)))

DEG_GOallontology.MLWMLQ <-rbind(DEG_BP.MLWMLQ.table, DEG_MF.MLWMLQ.table, DEG_CC.MLWMLQ.table)
write.csv(DEG_GOallontology.MLWMLQ, "Results/AllGosOntologyCombined/DEGs_GOallontology_MLWMLQ.csv")


#Less than .01
DEG_BP.MLWMLQ.table.01 <- DEG_BP.MLWMLQ.table[DEG_BP.MLWMLQ.table$raw.p.value < 0.01,]
DEG_MF.MLWMLQ.table.01 <- DEG_MF.MLWMLQ.table[DEG_MF.MLWMLQ.table$raw.p.value < 0.01,]
DEG_CC.MLWMLQ.table.01 <- DEG_CC.MLWMLQ.table[DEG_CC.MLWMLQ.table$raw.p.value < 0.01,]

DEG_GOallontology.MLWMLQ.01 <-rbind(DEG_BP.MLWMLQ.table.01, DEG_MF.MLWMLQ.table.01, DEG_CC.MLWMLQ.table.01)
write.csv(DEG_GOallontology.MLWMLQ.01, "Results/DEGs_GOallontology_MLWMLQ.01.csv")

```



### QLW v QLQ
#### 'Gene' list for DEGs
```{r degsgeneslistQLWQLQ}
DEG.data.QLWQLQ <- read.csv2(file = "../DEG/DEGFullResults/DEGananalysisResults_QLWQLQ.csv", header=TRUE, sep=",")
rownames(DEG.data.QLWQLQ) <- DEG.data.QLWQLQ$X
DEG.data.QLWQLQ <- DEG.data.QLWQLQ[,-1]
```


I discovered that having NAs in the padj is a problem when running the _fisher test via: https://www.biostars.org/p/91474/. Removed the NAs to move forward.
```{r removeNAs_QLWQLQ}
table(is.na(DEG.data.QLWQLQ$padj))
NAlist_QLWQLQ <- which(is.na(DEG.data.QLWQLQ$padj))
rownames(DEG.data.QLWQLQ)[NAlist_QLWQLQ]

#gene list for topgo input
DEG.genelist.QLWQLQ  <- as.numeric(DEG.data.QLWQLQ$padj)
names(DEG.genelist.QLWQLQ) <- geneUniverse


##NEED TO REMOVE NAs IN PADJ TO PROCEED
#VERIFY SAME POSITIONS ACROSS ALL PAIRWISE
QLWQLQ_napos <- which(is.na(DEG.genelist.QLWQLQ), arr.ind = TRUE)



#Create a new data frame to use going forward W/O those padj NAs
#This is needed for the _fisher tests
DEG.genelist.QLWQLQ <- DEG.genelist.QLWQLQ[-QLWQLQ_napos]

table(topDiffGenes(DEG.genelist.QLWQLQ))
table(DEG.genelist.QLWQLQ < 0.05)
```


#### BP
```{r BP_DEG_QLWQLQ}
DEgenes.QLWQLQ.GO.bp <- new("topGOdata",
                description = paste("DEG Go", "BP", "QLWQLQ", sep = "_"),
                ontology = "BP",
                allGenes = DEG.genelist.QLWQLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.QLWQLQ.bp_result_fisher <- runTest(DEgenes.QLWQLQ.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.QLWQLQ.bp_result_fisher

DEG_BP.QLWQLQ.table <- GenTable(DEgenes.QLWQLQ.GO.bp, 
                                raw.p.value = DEG.QLWQLQ.bp_result_fisher, 
                                topNodes = 50, 
                                numChar = 120)

write.csv(DEG_BP.QLWQLQ.table, "Results/ByOntology/QLWQLQ_BPTerms.csv")
```

#### MF
```{r mf_DEG_QLWQLQ}
DEgenes.QLWQLQ.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "QLWQLQ", sep = "_"),
                ontology = "MF",
                allGenes = DEG.genelist.QLWQLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.QLWQLQ.mf_result_fisher <- runTest(DEgenes.QLWQLQ.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.QLWQLQ.mf_result_fisher

DEG_MF.QLWQLQ.table <- GenTable(DEgenes.QLWQLQ.GO.mf, 
                                raw.p.value = DEG.QLWQLQ.mf_result_fisher, 
                                topNodes = 50,
                                numChar = 120)

write.csv(DEG_MF.QLWQLQ.table, "Results/ByOntology/QLWQLQ_MFTerms.csv")
```



#### CC
```{r cc_DEG_QLWQLQ}
DEgenes.QLWQLQ.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "QLWQLQ", sep = "_"),
                ontology = "CC",
                allGenes = DEG.genelist.QLWQLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.QLWQLQ.cc_result_fisher <- runTest(DEgenes.QLWQLQ.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.QLWQLQ.cc_result_fisher

DEG_CC.QLWQLQ.table <- GenTable(DEgenes.QLWQLQ.GO.cc, 
                                raw.p.value = DEG.QLWQLQ.cc_result_fisher,
                                topNodes = 50, 
                                numChar = 120)

write.csv(DEG_CC.QLWQLQ.table, "Results/ByOntology/QLWQLQ_CCTerms.csv")
```


```{r QLWQLQlargetable.DEG_Node=1}
#ALL
DEG_BP.QLWQLQ.table <- cbind(DEG_BP.QLWQLQ.table, ontology=rep("BP", length(DEG_BP.QLWQLQ.table$GO.ID)))
DEG_MF.QLWQLQ.table <- cbind(DEG_MF.QLWQLQ.table, ontology=rep("MF", length(DEG_MF.QLWQLQ.table$GO.ID)))
DEG_CC.QLWQLQ.table <- cbind(DEG_CC.QLWQLQ.table, ontology=rep("CC", length(DEG_CC.QLWQLQ.table$GO.ID)))

DEG_GOallontology.QLWQLQ <-rbind(DEG_BP.QLWQLQ.table, DEG_MF.QLWQLQ.table, DEG_CC.QLWQLQ.table)
write.csv(DEG_GOallontology.QLWQLQ, "Results/AllGosOntologyCombined/DEGs_GOallontology_QLWQLQ.csv")


#Less than .01
DEG_BP.QLWQLQ.table.01 <- DEG_BP.QLWQLQ.table[DEG_BP.QLWQLQ.table$raw.p.value < 0.01,]
DEG_MF.QLWQLQ.table.01 <- DEG_MF.QLWQLQ.table[DEG_MF.QLWQLQ.table$raw.p.value < 0.01,]
DEG_CC.QLWQLQ.table.01 <- DEG_CC.QLWQLQ.table[DEG_CC.QLWQLQ.table$raw.p.value < 0.01,]

DEG_GOallontology.QLWQLQ.01 <-rbind(DEG_BP.QLWQLQ.table.01, DEG_MF.QLWQLQ.table.01, DEG_CC.QLWQLQ.table.01)
write.csv(DEG_GOallontology.QLWQLQ.01, "Results/DEGs_GOallontology_QLWQLQ.01.csv")
```




## Within Castes--WORKERS
### FLW v MLW
#### 'Gene' list for DEGs
```{r degsgeneslistFLWMLW}
DEG.data.FLWMLW <- read.csv2(file = "../DEG/DEGFullResults/DEGananalysisResults_FLWMLW.csv", header=TRUE, sep=",")
rownames(DEG.data.FLWMLW) <- DEG.data.FLWMLW$X
DEG.data.FLWMLW <- DEG.data.FLWMLW[,-1]
```


I discovered that having NAs in the padj is a problem when running the _fisher test via: https://www.biostars.org/p/91474/. Removed the NAs to move forward.
```{r removeNAs_FLWMLW}
table(is.na(DEG.data.FLWMLW$padj))
NAlist_FLWMLW <- which(is.na(DEG.data.FLWMLW$padj))
rownames(DEG.data.FLWMLW)[NAlist_FLWMLW]

#gene list for topgo input
DEG.genelist.FLWMLW <- as.numeric(DEG.data.FLWMLW$padj)
names(DEG.genelist.FLWMLW) <- geneUniverse


##NEED TO REMOVE NAs IN PADJ TO PROCEED
#VERIFY SAME POSITIONS ACROSS ALL PAIRWISE
FLWMLW_napos <- which(is.na(DEG.genelist.FLWMLW), arr.ind = TRUE)



#Create a new data frame to use going forward W/O those padj NAs
#This is needed for the _fisher tests
DEG.genelist.FLWMLW <- DEG.genelist.FLWMLW[-FLWMLW_napos]

table(topDiffGenes(DEG.genelist.FLWMLW))
table(DEG.genelist.FLWMLW < 0.05)

```


#### BP
```{r BP_DEG_FLWMLW}
DEgenes.FLWMLW.GO.bp <- new("topGOdata",
                description = paste("DEG Go", "BP", "FLWMLW", sep = "_"),
                ontology = "BP",
                allGenes = DEG.genelist.FLWMLW, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLWMLW.bp_result_fisher <- runTest(DEgenes.FLWMLW.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.FLWMLW.bp_result_fisher

DEG_BP.FLWMLW.table <- GenTable(DEgenes.FLWMLW.GO.bp, 
                                raw.p.value = DEG.FLWMLW.bp_result_fisher, 
                                topNodes = 50, 
                                numChar = 120)

write.csv(DEG_BP.FLWMLW.table, "Results/ByOntology/FLWMLW_BPTerms.csv")
```

#### MF
```{r mf_DEG_FLWMLW}
DEgenes.FLWMLW.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "FLWMLW", sep = "_"),
                ontology = "MF",
                allGenes = DEG.genelist.FLWMLW, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLWMLW.mf_result_fisher <- runTest(DEgenes.FLWMLW.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.FLWMLW.mf_result_fisher

DEG_MF.FLWMLW.table <- GenTable(DEgenes.FLWMLW.GO.mf,
                                raw.p.value = DEG.FLWMLW.mf_result_fisher,
                                topNodes = 50, 
                                numChar = 120)

write.csv(DEG_MF.FLWMLW.table, "Results/ByOntology/FLWMLW_MFTerms.csv")
```



#### CC
```{r cc_DEG_FLWMLW}
DEgenes.FLWMLW.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "FLWMLW", sep = "_"),
                ontology = "CC",
                allGenes = DEG.genelist.FLWMLW, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLWMLW.cc_result_fisher <- runTest(DEgenes.FLWMLW.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.FLWMLW.cc_result_fisher

DEG_CC.FLWMLW.table <- GenTable(DEgenes.FLWMLW.GO.cc,
                                raw.p.value = DEG.FLWMLW.cc_result_fisher, 
                                topNodes = 50,
                                numChar = 120)

write.csv(DEG_CC.FLWMLW.table, "Results/ByOntology/FLWMLW_CCTerms.csv")
```

```{r FLWMLWlargetable.DEG_Node=1}
#ALL
DEG_BP.FLWMLW.table <- cbind(DEG_BP.FLWMLW.table, ontology=rep("BP", length(DEG_BP.FLWMLW.table$GO.ID)))
DEG_MF.FLWMLW.table <- cbind(DEG_MF.FLWMLW.table, ontology=rep("MF", length(DEG_MF.FLWMLW.table$GO.ID)))
DEG_CC.FLWMLW.table <- cbind(DEG_CC.FLWMLW.table, ontology=rep("CC", length(DEG_CC.FLWMLW.table$GO.ID)))

DEG_GOallontology.FLWMLW <-rbind(DEG_BP.FLWMLW.table, DEG_MF.FLWMLW.table, DEG_CC.FLWMLW.table)
write.csv(DEG_GOallontology.FLWMLW, "Results/AllGosOntologyCombined/DEGs_GOallontology_FLWMLW.csv")


#Less than .01
DEG_BP.FLWMLW.table.01 <- DEG_BP.FLWMLW.table[DEG_BP.FLWMLW.table$raw.p.value < 0.01,]
DEG_MF.FLWMLW.table.01 <- DEG_MF.FLWMLW.table[DEG_MF.FLWMLW.table$raw.p.value < 0.01,]
DEG_CC.FLWMLW.table.01 <- DEG_CC.FLWMLW.table[DEG_CC.FLWMLW.table$raw.p.value < 0.01,]

DEG_GOallontology.FLWMLW.01 <-rbind(DEG_BP.FLWMLW.table.01, DEG_MF.FLWMLW.table.01, DEG_CC.FLWMLW.table.01)
write.csv(DEG_GOallontology.FLWMLW.01, "Results/DEGs_GOallontology_FLWMLW.01.csv")
```



### FLW v QLW--NO DEGs
No analysis conducted because no DEGs.

### MLW v QLW--NO DEGs
No analysis conducted because no DEGs.






## Within Castes--QUEENS
### FLQ v MLQ
#### 'Gene' list for DEGs
```{r degsgeneslistFLQMLQ}
DEG.data.FLQMLQ <- read.csv2(file = "../DEG/DEGFullResults/DEGananalysisResults_FLQMLQ.csv", header=TRUE, sep=",")
rownames(DEG.data.FLQMLQ) <- DEG.data.FLQMLQ$X
DEG.data.FLQMLQ <- DEG.data.FLQMLQ[,-1]
```


I discovered that having NAs in the padj is a problem when running the KS test via: https://www.biostars.org/p/91474/. Removed the NAs to move forward.
```{r removeNAs_FLQMLQ}
table(is.na(DEG.data.FLQMLQ$padj))
NAlist_FLQMLQ <- which(is.na(DEG.data.FLQMLQ$padj))
rownames(DEG.data.FLQMLQ)[NAlist_FLQMLQ]

#gene list for topgo input
DEG.genelist.FLQMLQ <- as.numeric(DEG.data.FLQMLQ$padj)
names(DEG.genelist.FLQMLQ) <- geneUniverse


##NEED TO REMOVE NAs IN PADJ TO PROCEED
#VERIFY SAME POSITIONS ACROSS ALL PAIRWISE
FLQMLQ_napos <- which(is.na(DEG.genelist.FLQMLQ), arr.ind = TRUE)



#Create a new data frame to use going forward W/O those padj NAs
#This is needed for the _fisher tests
DEG.genelist.FLQMLQ <- DEG.genelist.FLQMLQ[-FLQMLQ_napos]


table(topDiffGenes(DEG.genelist.FLQMLQ))
table(DEG.genelist.FLQMLQ < 0.05)

```


#### BP
```{r BP_DEG_FLQMLQ}
DEgenes.FLQMLQ.GO.bp <- new("topGOdata",
                description = paste("DEG Go", "BP", "FLQMLQ", sep = "_"),
                ontology = "BP",
                allGenes = DEG.genelist.FLQMLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLQMLQ.bp_result_fisher <- runTest(DEgenes.FLQMLQ.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.FLQMLQ.bp_result_fisher

DEG_BP.FLQMLQ.table <- GenTable(DEgenes.FLQMLQ.GO.bp, 
                                raw.p.value = DEG.FLQMLQ.bp_result_fisher,
                                topNodes = 50, 
                                numChar = 120)

write.csv(DEG_BP.FLQMLQ.table, "Results/ByOntology/FLQMLQ_BPTerms.csv")
```

#### MF
```{r mf_DEG_FLQMLQ}
DEgenes.FLQMLQ.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "FLQMLQ", sep = "_"),
                ontology = "MF",
                allGenes = DEG.genelist.FLQMLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLQMLQ.mf_result_fisher <- runTest(DEgenes.FLQMLQ.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.FLQMLQ.mf_result_fisher

DEG_MF.FLQMLQ.table <- GenTable(DEgenes.FLQMLQ.GO.mf, 
                                raw.p.value = DEG.FLQMLQ.mf_result_fisher,
                                topNodes = 50,
                                numChar = 120)

write.csv(DEG_MF.FLQMLQ.table, "Results/ByOntology/FLQMLQ_MFTerms.csv")
```



#### CC
```{r cc_DEG_FLQMLQ}
DEgenes.FLQMLQ.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "FLQMLQ", sep = "_"),
                ontology = "CC",
                allGenes = DEG.genelist.FLQMLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLQMLQ.cc_result_fisher <- runTest(DEgenes.FLQMLQ.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.FLQMLQ.cc_result_fisher

DEG_CC.FLQMLQ.table <- GenTable(DEgenes.FLQMLQ.GO.cc, 
                                raw.p.value = DEG.FLQMLQ.cc_result_fisher,
                                topNodes = 50,
                                numChar = 120)

write.csv(DEG_CC.FLQMLQ.table, "Results/ByOntology/FLQMLQ_CCTerms.csv")
```

```{r FLQMLQlargetable.DEG_Node=1}
#ALL
DEG_BP.FLQMLQ.table <- cbind(DEG_BP.FLQMLQ.table, ontology=rep("BP", length(DEG_BP.FLQMLQ.table$GO.ID)))
DEG_MF.FLQMLQ.table <- cbind(DEG_MF.FLQMLQ.table, ontology=rep("MF", length(DEG_MF.FLQMLQ.table$GO.ID)))
DEG_CC.FLQMLQ.table <- cbind(DEG_CC.FLQMLQ.table, ontology=rep("CC", length(DEG_CC.FLQMLQ.table$GO.ID)))

DEG_GOallontology.FLQMLQ <-rbind(DEG_BP.FLQMLQ.table, DEG_MF.FLQMLQ.table, DEG_CC.FLQMLQ.table)
write.csv(DEG_GOallontology.FLQMLQ, "Results/AllGosOntologyCombined/DEGs_GOallontology_FLQMLQ.csv")


#Less than .01
DEG_BP.FLQMLQ.table.01 <- DEG_BP.FLQMLQ.table[DEG_BP.FLQMLQ.table$raw.p.value < 0.01,]
DEG_MF.FLQMLQ.table.01 <- DEG_MF.FLQMLQ.table[DEG_MF.FLQMLQ.table$raw.p.value < 0.01,]
DEG_CC.FLQMLQ.table.01 <- DEG_CC.FLQMLQ.table[DEG_CC.FLQMLQ.table$raw.p.value < 0.01,]

DEG_GOallontology.FLQMLQ.01 <-rbind(DEG_BP.FLQMLQ.table.01, DEG_MF.FLQMLQ.table.01, DEG_CC.FLQMLQ.table.01)
write.csv(DEG_GOallontology.FLQMLQ.01, "Results/DEGs_GOallontology_FLQMLQ.01.csv")
```



### FLQ v QLQ
#### 'Gene' list for DEGs
```{r degsgeneslistFLQQLQ}
DEG.data.FLQQLQ <- read.csv2(file = "../DEG/DEGFullResults/DEGananalysisResults_FLQQLQ.csv", header=TRUE, sep=",")
rownames(DEG.data.FLQQLQ) <- DEG.data.FLQQLQ$X
DEG.data.FLQQLQ <- DEG.data.FLQQLQ[,-1]
```


I discovered that having NAs in the padj is a problem when running the _fisher test via: https://www.biostars.org/p/91474/. Removed the NAs to move forward.
```{r removeNAs_FLQQLQ}
table(is.na(DEG.data.FLQQLQ$padj))
NAlist_FLQQLQ <- which(is.na(DEG.data.FLQQLQ$padj))
rownames(DEG.data.FLQQLQ)[NAlist_FLQQLQ]

#gene list for topgo input
DEG.genelist.FLQQLQ <- as.numeric(DEG.data.FLQQLQ$padj)
names(DEG.genelist.FLQQLQ) <- geneUniverse


##NEED TO REMOVE NAs IN PADJ TO PROCEED
#VERIFY SAME POSITIONS ACROSS ALL PAIRWISE
FLQQLQ_napos <- which(is.na(DEG.genelist.FLQQLQ), arr.ind = TRUE)



#Create a new data frame to use going forward W/O those padj NAs
#This is needed for the _fisher tests
DEG.genelist.FLQQLQ <- DEG.genelist.FLQQLQ[-FLQQLQ_napos]


table(topDiffGenes(DEG.genelist.FLQQLQ))
table(DEG.genelist.FLQQLQ < 0.05)
```


#### BP
```{r BP_DEG_FLQQLQ}
DEgenes.FLQQLQ.GO.bp <- new("topGOdata",
                description = paste("DEG Go", "BP", "FLQQLQ", sep = "_"),
                ontology = "BP",
                allGenes = DEG.genelist.FLQQLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLQQLQ.bp_result_fisher <- runTest(DEgenes.FLQQLQ.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.FLQQLQ.bp_result_fisher

DEG_BP.FLQQLQ.table <- GenTable(DEgenes.FLQQLQ.GO.bp,
                                raw.p.value = DEG.FLQQLQ.bp_result_fisher,
                                topNodes = 50,
                                numChar = 120)

write.csv(DEG_BP.FLQQLQ.table, "Results/ByOntology/FLQQLQ_BPTerms.csv")
```

#### MF
```{r mf_DEG_FLQQLQ}
DEgenes.FLQQLQ.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "FLQQLQ", sep = "_"),
                ontology = "MF",
                allGenes = DEG.genelist.FLQQLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLQQLQ.mf_result_fisher <- runTest(DEgenes.FLQQLQ.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.FLQQLQ.mf_result_fisher

DEG_MF.FLQQLQ.table <- GenTable(DEgenes.FLQQLQ.GO.mf, 
                                raw.p.value = DEG.FLQQLQ.mf_result_fisher,
                                topNodes = 50, 
                                numChar = 120)

write.csv(DEG_MF.FLQQLQ.table, "Results/ByOntology/FLQQLQ_MFTerms.csv")
```



#### CC
```{r cc_DEG_FLQQLQ}
DEgenes.FLQQLQ.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "FLQQLQ", sep = "_"),
                ontology = "CC",
                allGenes = DEG.genelist.FLQQLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLQQLQ.cc_result_fisher <- runTest(DEgenes.FLQQLQ.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.FLQQLQ.cc_result_fisher

DEG_CC.FLQQLQ.table <- GenTable(DEgenes.FLQQLQ.GO.cc,
                                raw.p.value = DEG.FLQQLQ.cc_result_fisher,
                                topNodes = 50,
                                numChar = 120)

write.csv(DEG_CC.FLQQLQ.table, "Results/ByOntology/FLQQLQ_CCTerms.csv")
```

```{r FLQQLQlargetable.DEG_Node=1}
#ALL
DEG_BP.FLQQLQ.table <- cbind(DEG_BP.FLQQLQ.table, ontology=rep("BP", length(DEG_BP.FLQQLQ.table$GO.ID)))
DEG_MF.FLQQLQ.table <- cbind(DEG_MF.FLQQLQ.table, ontology=rep("MF", length(DEG_MF.FLQQLQ.table$GO.ID)))
DEG_CC.FLQQLQ.table <- cbind(DEG_CC.FLQQLQ.table, ontology=rep("CC", length(DEG_CC.FLQQLQ.table$GO.ID)))

DEG_GOallontology.FLQQLQ <-rbind(DEG_BP.FLQQLQ.table, DEG_MF.FLQQLQ.table, DEG_CC.FLQQLQ.table)
write.csv(DEG_GOallontology.FLQQLQ, "Results/AllGosOntologyCombined/DEGs_GOallontology_FLQQLQ.csv")


#Less than .01
DEG_BP.FLQQLQ.table.01 <- DEG_BP.FLQQLQ.table[DEG_BP.FLQQLQ.table$raw.p.value < 0.01,]
DEG_MF.FLQQLQ.table.01 <- DEG_MF.FLQQLQ.table[DEG_MF.FLQQLQ.table$raw.p.value < 0.01,]
DEG_CC.FLQQLQ.table.01 <- DEG_CC.FLQQLQ.table[DEG_CC.FLQQLQ.table$raw.p.value < 0.01,]

DEG_GOallontology.FLQQLQ.01 <-rbind(DEG_BP.FLQQLQ.table.01, DEG_MF.FLQQLQ.table.01, DEG_CC.FLQQLQ.table.01)
write.csv(DEG_GOallontology.FLQQLQ.01, "Results/DEGs_GOallontology_FLQQLQ.01.csv")
```




### MLQ v QLQ
#### 'Gene' list for DEGs
```{r degsgeneslistMLQQLQ}
DEG.data.MLQQLQ <- read.csv2(file = "../DEG/DEGFullResults/DEGananalysisResults_MLQQLQ.csv", header=TRUE, sep=",")
rownames(DEG.data.MLQQLQ) <- DEG.data.MLQQLQ$X
DEG.data.MLQQLQ <- DEG.data.MLQQLQ[,-1]
```


I discovered that having NAs in the padj is a problem when running the _fisher test via: https://www.biostars.org/p/91474/. Removed the NAs to move forward.
```{r removeNAs_MLQQLQ}
table(is.na(DEG.data.MLQQLQ$padj))
NAlist_MLQQLQ <- which(is.na(DEG.data.MLQQLQ$padj))
rownames(DEG.data.MLQQLQ)[NAlist_MLQQLQ]

#gene list for topgo input
DEG.genelist.MLQQLQ <- as.numeric(DEG.data.MLQQLQ$padj)
names(DEG.genelist.MLQQLQ) <- geneUniverse


##NEED TO REMOVE NAs IN PADJ TO PROCEED
#VERIFY SAME POSITIONS ACROSS ALL PAIRWISE
MLQQLQ_napos <- which(is.na(DEG.genelist.MLQQLQ), arr.ind = TRUE)



#Create a new data frame to use going forward W/O those padj NAs
#This is needed for the _fisher tests
DEG.genelist.MLQQLQ <- DEG.genelist.MLQQLQ[-MLQQLQ_napos]


table(topDiffGenes(DEG.genelist.MLQQLQ))
table(DEG.genelist.MLQQLQ < 0.05)
```


#### BP
```{r BP_DEG_MLQQLQ}
DEgenes.MLQQLQ.GO.bp <- new("topGOdata",
                description = paste("DEG Go", "BP", "MLQQLQ", sep = "_"),
                ontology = "BP",
                allGenes = DEG.genelist.MLQQLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.MLQQLQ.bp_result_fisher <- runTest(DEgenes.MLQQLQ.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.MLQQLQ.bp_result_fisher

DEG_BP.MLQQLQ.table <- GenTable(DEgenes.MLQQLQ.GO.bp, 
                                raw.p.value = DEG.MLQQLQ.bp_result_fisher,
                                topNodes = 50,
                                numChar = 120)

write.csv(DEG_BP.MLQQLQ.table, "Results/ByOntology/MLQQLQ_BPTerms.csv")
```

#### MF
```{r mf_DEG_MLQQLQ}
DEgenes.MLQQLQ.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "MLQQLQ", sep = "_"),
                ontology = "MF",
                allGenes = DEG.genelist.MLQQLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.MLQQLQ.mf_result_fisher <- runTest(DEgenes.MLQQLQ.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.MLQQLQ.mf_result_fisher

DEG_MF.MLQQLQ.table <- GenTable(DEgenes.MLQQLQ.GO.mf,
                                raw.p.value = DEG.MLQQLQ.mf_result_fisher,
                                topNodes =50,
                                numChar = 120)

write.csv(DEG_MF.MLQQLQ.table, "Results/ByOntology/MLQQLQ_MFTerms.csv")
```



#### CC
```{r cc_DEG_MLQQLQ}
DEgenes.MLQQLQ.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "MLQQLQ", sep = "_"),
                ontology = "CC",
                allGenes = DEG.genelist.MLQQLQ, # all genes in the "universe" aka our experiment
                geneSel = topDiffGenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.MLQQLQ.cc_result_fisher <- runTest(DEgenes.MLQQLQ.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.MLQQLQ.cc_result_fisher

DEG_CC.MLQQLQ.table <- GenTable(DEgenes.MLQQLQ.GO.cc,
                                raw.p.value = DEG.MLQQLQ.cc_result_fisher,
                                topNodes = 50,
                                numChar = 120)

write.csv(DEG_CC.MLQQLQ.table, "Results/ByOntology/MLQQLQ_CCTerms.csv")
```

```{r MLQQLQlargetable.DEG_Node=1}
#ALL
DEG_BP.MLQQLQ.table <- cbind(DEG_BP.MLQQLQ.table, ontology=rep("BP", length(DEG_BP.MLQQLQ.table$GO.ID)))
DEG_MF.MLQQLQ.table <- cbind(DEG_MF.MLQQLQ.table, ontology=rep("MF", length(DEG_MF.MLQQLQ.table$GO.ID)))
DEG_CC.MLQQLQ.table <- cbind(DEG_CC.MLQQLQ.table, ontology=rep("CC", length(DEG_CC.MLQQLQ.table$GO.ID)))

DEG_GOallontology.MLQQLQ <-rbind(DEG_BP.MLQQLQ.table, DEG_MF.MLQQLQ.table, DEG_CC.MLQQLQ.table)
write.csv(DEG_GOallontology.MLQQLQ, "Results/AllGosOntologyCombined/DEGs_GOallontology_MLQQLQ.csv")


#Less than .01
DEG_BP.MLQQLQ.table.01 <- DEG_BP.MLQQLQ.table[DEG_BP.MLQQLQ.table$raw.p.value < 0.01,]
DEG_MF.MLQQLQ.table.01 <- DEG_MF.MLQQLQ.table[DEG_MF.MLQQLQ.table$raw.p.value < 0.01,]
DEG_CC.MLQQLQ.table.01 <- DEG_CC.MLQQLQ.table[DEG_CC.MLQQLQ.table$raw.p.value < 0.01,]

DEG_GOallontology.MLQQLQ.01 <-rbind(DEG_BP.MLQQLQ.table.01, DEG_MF.MLQQLQ.table.01, DEG_CC.MLQQLQ.table.01)
write.csv(DEG_GOallontology.MLQQLQ.01, "Results/DEGs_GOallontology_MLQQLQ.01.csv")
```




```{r alluniqueGO}
##BP##
# Gather all GO terms across all analyses
all_GOs_BP <- c(DEG_BP.FLWFLQ.table.01$GO.ID, DEG_BP.MLWMLQ.table.01$GO.ID, DEG_BP.QLWQLQ.table.01$GO.ID, DEG_BP.FLWMLW.table.01$GO.ID, DEG_BP.FLQMLQ.table.01$GO.ID, DEG_BP.FLQQLQ.table.01$GO.ID, DEG_BP.MLQQLQ.table.01$GO.ID)

# Save only the unique GO.IDs
# We will pull from this to build df
unique_GO_BP <- unique(all_GOs_BP)

# Make df to store the results using the unique GO IDs.
#All columns will be this length because it is just T/F later
GO_df_BP <- data.frame(GOID = unique_GO_BP)

# List of data frame to check aka the results
#doing this just for P < 0.01
GOlists <- list(
  FLWFLQ = DEG_BP.FLWFLQ.table.01, 
  MLWMLQ = DEG_BP.MLWMLQ.table.01, 
  QLWQLQ = DEG_BP.QLWQLQ.table.01, 
  FLWMLW = DEG_BP.FLWMLW.table.01, 
  FLQMLQ = DEG_BP.FLQMLQ.table.01, 
  FLQQLQ = DEG_BP.FLQQLQ.table.01, 
  MLQQLQ = DEG_BP.MLQQLQ.table.01
)

# Check for each GO.ID in individual dfs from the list of dfs
#return TRUE/FALSE
check_for_GOID <- function(df, id) {
  return(id %in% df$GO.ID)  # Check if GO.ID is present in the current data frame
}

# Loop through each data frame in GOlists and check for each GO.ID
for (df_name in names(GOlists)) {
  # Get the current data frame
  current_df <- GOlists[[df_name]]
  
  # Check if each GO.ID is present in the current data frame and store the result
  GO_df_BP[[df_name]] <- sapply(GO_df_BP$GOID, function(id) check_for_GOID(current_df, id))
}

#Modify the results so that only TRUE is shown
GO_df_BP<- GO_df_BP %>%
  mutate(across(everything(), ~ ifelse(. == FALSE, "", .)))

print(GO_df_BP)


write.csv(GO_df_BP, "Results/AllResults_fisherp01BP.csv")



##MF##
# Gather all GO terms across all analyses
all_GOs_MF <- c(DEG_MF.FLWFLQ.table.01$GO.ID, DEG_MF.MLWMLQ.table.01$GO.ID, DEG_MF.QLWQLQ.table.01$GO.ID, DEG_MF.FLWMLW.table.01$GO.ID, DEG_MF.FLQMLQ.table.01$GO.ID, DEG_MF.FLQQLQ.table.01$GO.ID, DEG_MF.MLQQLQ.table.01$GO.ID)

# Save only the unique GO.IDs
# We will pull from this to build df
unique_GO_MF <- unique(all_GOs_MF)

# Make df to store the results using the unique GO IDs.
#All columns will be this length because it is just T/F later
GO_df_MF <- data.frame(GOID = unique_GO_MF)

# List of data frame to check aka the results
#doing this just for P < 0.01
GOlists_MF <- list(
  FLWFLQ = DEG_MF.FLWFLQ.table.01, 
  MLWMLQ = DEG_MF.MLWMLQ.table.01, 
  QLWQLQ = DEG_MF.QLWQLQ.table.01, 
  FLWMLW = DEG_MF.FLWMLW.table.01, 
  FLQMLQ = DEG_MF.FLQMLQ.table.01, 
  FLQQLQ = DEG_MF.FLQQLQ.table.01, 
  MLQQLQ = DEG_MF.MLQQLQ.table.01
)


# Loop through each data frame in GOlists and check for each GO.ID
for (df_name in names(GOlists_MF)) {
  # Get the current data frame
  current_df <- GOlists_MF[[df_name]]
  
  # Check if each GO.ID is present in the current data frame and store the result
  GO_df_MF[[df_name]] <- sapply(GO_df_MF$GOID, function(id) check_for_GOID(current_df, id))
}

#Modify the results so that only TRUE is shown
GO_df_MF <- GO_df_MF %>%
  mutate(across(everything(), ~ ifelse(. == FALSE, "", .)))

print(GO_df_MF)


write.csv(GO_df_MF, "Results/AllResults_fisherp01MF.csv")




##CC##
# Gather all GO terms across all analyses
all_GOs_CC <- c(DEG_CC.FLWFLQ.table.01$GO.ID, DEG_CC.MLWMLQ.table.01$GO.ID, DEG_CC.QLWQLQ.table.01$GO.ID, DEG_CC.FLWMLW.table.01$GO.ID, DEG_CC.FLQMLQ.table.01$GO.ID, DEG_CC.FLQQLQ.table.01$GO.ID, DEG_CC.MLQQLQ.table.01$GO.ID)

# Save only the unique GO.IDs
# We will pull from this to build df
unique_GO_CC <- unique(all_GOs_CC)

# Make df to store the results using the unique GO IDs.
#All columns will be this length because it is just T/F later
GO_df_CC <- data.frame(GOID = unique_GO_CC)

# List of data frame to check aka the results
#doing this just for P < 0.01
GOlists_CC <- list(
  FLWFLQ = DEG_CC.FLWFLQ.table.01, 
  MLWMLQ = DEG_CC.MLWMLQ.table.01, 
  QLWQLQ = DEG_CC.QLWQLQ.table.01, 
  FLWMLW = DEG_CC.FLWMLW.table.01, 
  FLQMLQ = DEG_CC.FLQMLQ.table.01, 
  FLQQLQ = DEG_CC.FLQQLQ.table.01, 
  MLQQLQ = DEG_CC.MLQQLQ.table.01
)


# Loop through each data frame in GOlists and check for each GO.ID
for (df_name in names(GOlists_CC)) {
  # Get the current data frame
  current_df <- GOlists_CC[[df_name]]
  
  # Check if each GO.ID is present in the current data frame and store the result
  GO_df_CC[[df_name]] <- sapply(GO_df_CC$GOID, function(id) check_for_GOID(current_df, id))
}

#Modify the results so that only TRUE is shown
GO_df_CC <- GO_df_CC %>%
  mutate(across(everything(), ~ ifelse(. == FALSE, "", .)))

print(GO_df_CC)


write.csv(GO_df_CC, "Results/AllResults_fisherp01CC.csv")
write_xlsx(list(BP=GO_df_BP, MF=GO_df_MF, CC=GO_df_CC), path = "Results/AllAnalyses_truesetup.xlsx", col_names = TRUE)


# library(xlsx)
# if(Sys.getenv("JAVA_HOME")!=""){
#     Sys.setenv(JAVA_HOME="")
# }
# library(rJava)
# write.xlsx(DEG_BP.FLWFLQ.table.01, file="Results/AllAnalyses.csv", sheetName="sheet1", row.names=FALSE)
# write.xlsx(DEG_BP.MLWMLQ.table.01, file="Results/AllAnalyses.csv", sheetName="sheet2", append=TRUE, row.names=FALSE)
# write.xlsx(DEG_BP.QLWQLQ.table.01, file="Results/AllAnalyses.csv", sheetName="sheet3", append=TRUE, row.names=FALSE)
# write.xlsx(DEG_BP.FLWMLW.table.01, file="Results/AllAnalyses.csv", sheetName="sheet4", append=TRUE, row.names=FALSE)
# write.xlsx(DEG_BP.FLQMLQ.table.01, file="Results/AllAnalyses.csv", sheetName="sheet5", append=TRUE, row.names=FALSE)
# write.xlsx(DEG_BP.FLQQLQ.table.01, file="Results/AllAnalyses.csv", sheetName="sheet6", append=TRUE, row.names=FALSE)
# write.xlsx(DEG_BP.MLQQLQ.table.01, file="Results/AllAnalyses.csv", sheetName="sheet7", append=TRUE, row.names=FALSE)

library(writexl)

GOlists_AllOntology <- list(
  FLWFLQ = DEG_GOallontology.FLWFLQ.01, 
  MLWMLQ = DEG_GOallontology.MLWMLQ.01, 
  QLWQLQ = DEG_GOallontology.QLWQLQ.01, 
  FLWMLW = DEG_GOallontology.FLWMLW.01, 
  FLQMLQ = DEG_GOallontology.FLQMLQ.01, 
  FLQQLQ = DEG_GOallontology.FLQQLQ.01, 
  MLQQLQ = DEG_GOallontology.MLQQLQ.01
)

write_xlsx(GOlists_AllOntology, path = "Results/AllAnalyses.xlsx", col_names = TRUE)

```


```{r adddefs}
FULLtermsandDefs <- read.csv2(file = "../GSEA/combineGO.csv", header=TRUE, sep=",")

term_def <- FULLtermsandDefs[,c("GO", "term", "definition")]

term_def2 <- unique(term_def, by = "GO")
colnames(term_def2)[1] <- c("GO.ID")



DEG_GOallontology.FLWFLQ.01_mergetermdef <- merge(DEG_GOallontology.FLWFLQ.01, term_def2, by = "GO.ID", all.x = TRUE)
DEG_GOallontology.MLWMLQ.01_mergetermdef <- merge(DEG_GOallontology.MLWMLQ.01, term_def2, by = "GO.ID", all.x = TRUE)
DEG_GOallontology.QLWQLQ.01_mergetermdef <- merge(DEG_GOallontology.QLWQLQ.01, term_def2, by = "GO.ID", all.x = TRUE)
DEG_GOallontology.FLWMLW.01_mergetermdef <- merge(DEG_GOallontology.FLWMLW.01, term_def2, by = "GO.ID", all.x = TRUE)
DEG_GOallontology.FLQMLQ.01_mergetermdef <- merge(DEG_GOallontology.FLQMLQ.01, term_def2, by = "GO.ID", all.x = TRUE)
DEG_GOallontology.FLQQLQ.01_mergetermdef <- merge(DEG_GOallontology.FLQQLQ.01, term_def2, by = "GO.ID", all.x = TRUE)
DEG_GOallontology.MLQQLQ.01_mergetermdef <- merge(DEG_GOallontology.MLQQLQ.01, term_def2, by = "GO.ID", all.x = TRUE)

#verifying terms match
table(DEG_GOallontology.FLWFLQ.01_mergetermdef$Term==DEG_GOallontology.FLWFLQ.01_mergetermdef$term)
table(DEG_GOallontology.MLWMLQ.01_mergetermdef$Term==DEG_GOallontology.MLWMLQ.01_mergetermdef$term)
table(DEG_GOallontology.QLWQLQ.01_mergetermdef$Term==DEG_GOallontology.QLWQLQ.01_mergetermdef$term)
table(DEG_GOallontology.FLWMLW.01_mergetermdef$Term==DEG_GOallontology.FLWMLW.01_mergetermdef$term)
table(DEG_GOallontology.FLQMLQ.01_mergetermdef$Term==DEG_GOallontology.FLQMLQ.01_mergetermdef$term)
table(DEG_GOallontology.FLQQLQ.01_mergetermdef$Term==DEG_GOallontology.FLQQLQ.01_mergetermdef$term)
#one false
#False is 31; MAH verified and it is okay
table(DEG_GOallontology.MLQQLQ.01_mergetermdef$Term==DEG_GOallontology.MLQQLQ.01_mergetermdef$term)


#All Terms=terms; removing "terms" column
DEG_GOallontology.FLWFLQ.01_mergetermdef <- DEG_GOallontology.FLWFLQ.01_mergetermdef[,-8]
DEG_GOallontology.MLWMLQ.01_mergetermdef <- DEG_GOallontology.MLWMLQ.01_mergetermdef[,-8]
DEG_GOallontology.QLWQLQ.01_mergetermdef <- DEG_GOallontology.QLWQLQ.01_mergetermdef[,-8]
DEG_GOallontology.FLWMLW.01_mergetermdef <- DEG_GOallontology.FLWMLW.01_mergetermdef[,-8]
DEG_GOallontology.FLQMLQ.01_mergetermdef <- DEG_GOallontology.FLQMLQ.01_mergetermdef[,-8]
DEG_GOallontology.FLQQLQ.01_mergetermdef <- DEG_GOallontology.FLQQLQ.01_mergetermdef[,-8]
DEG_GOallontology.MLQQLQ.01_mergetermdef <- DEG_GOallontology.MLQQLQ.01_mergetermdef[,-8]

#Gsub the < symbols from these when present so I can save as numeric and sort
DEG_GOallontology.FLWFLQ.01_mergetermdef$raw.p.value <- gsub("< ", "", DEG_GOallontology.FLWFLQ.01_mergetermdef$raw.p.value)
DEG_GOallontology.FLQMLQ.01_mergetermdef$raw.p.value <- gsub("< ", "", DEG_GOallontology.FLQMLQ.01_mergetermdef$raw.p.value)
DEG_GOallontology.FLQMLQ.01_mergetermdef$raw.p.value <- gsub("<", "", DEG_GOallontology.FLQMLQ.01_mergetermdef$raw.p.value)



#make sure P is numeric
#All Terms=terms; removing "terms" column

as.numeric(DEG_GOallontology.FLWFLQ.01_mergetermdef$raw.p.value)
as.numeric(DEG_GOallontology.MLWMLQ.01_mergetermdef$raw.p.value)
as.numeric(DEG_GOallontology.QLWQLQ.01_mergetermdef$raw.p.value)
as.numeric(DEG_GOallontology.FLWMLW.01_mergetermdef$raw.p.value)
as.numeric(DEG_GOallontology.FLQMLQ.01_mergetermdef$raw.p.value)
as.numeric(DEG_GOallontology.FLQQLQ.01_mergetermdef$raw.p.value)
as.numeric(DEG_GOallontology.MLQQLQ.01_mergetermdef$raw.p.value)


#order by group and then ascending pvalue
# those in sci not will be at the bottom
DEG_GOallontology.FLWFLQ.01_mergetermdef <- DEG_GOallontology.FLWFLQ.01_mergetermdef %>%
  arrange(ontology, raw.p.value) 
DEG_GOallontology.MLWMLQ.01_mergetermdef <- DEG_GOallontology.MLWMLQ.01_mergetermdef %>%
  arrange(ontology, raw.p.value)
DEG_GOallontology.QLWQLQ.01_mergetermdef <- DEG_GOallontology.QLWQLQ.01_mergetermdef %>%
  arrange(ontology, raw.p.value) 
DEG_GOallontology.FLWMLW.01_mergetermdef <- DEG_GOallontology.FLWMLW.01_mergetermdef %>%
  arrange(ontology, raw.p.value)
DEG_GOallontology.FLQMLQ.01_mergetermdef <- DEG_GOallontology.FLQMLQ.01_mergetermdef %>%
  arrange(ontology, raw.p.value) 
DEG_GOallontology.FLQQLQ.01_mergetermdef <- DEG_GOallontology.FLQQLQ.01_mergetermdef %>%
  arrange(ontology, raw.p.value) 
DEG_GOallontology.MLQQLQ.01_mergetermdef <- DEG_GOallontology.MLQQLQ.01_mergetermdef %>%
  arrange(ontology, raw.p.value)


#write them all to CSV
write.csv(DEG_GOallontology.FLWFLQ.01_mergetermdef, "Results/TermsDefs/FLWFLQ_fisher1_TD.csv")
write.csv(DEG_GOallontology.MLWMLQ.01_mergetermdef, "Results/TermsDefs/FLWFLQ_fisher1_TD.csv")
write.csv(DEG_GOallontology.QLWQLQ.01_mergetermdef, "Results/TermsDefs/FLWFLQ_fisher1_TD.csv")
write.csv(DEG_GOallontology.FLWMLW.01_mergetermdef, "Results/TermsDefs/FLWFLQ_fisher1_TD.csv")
write.csv(DEG_GOallontology.FLQMLQ.01_mergetermdef, "Results/TermsDefs/FLWFLQ_fisher1_TD.csv")
write.csv(DEG_GOallontology.FLQQLQ.01_mergetermdef, "Results/TermsDefs/FLWFLQ_fisher1_TD.csv")
write.csv(DEG_GOallontology.MLQQLQ.01_mergetermdef, "Results/TermsDefs/FLWFLQ_fisher1_TD.csv")


GOlists_AllOntology_TD <- list(
  FLWFLQ = DEG_GOallontology.FLWFLQ.01_mergetermdef, 
  MLWMLQ = DEG_GOallontology.MLWMLQ.01_mergetermdef, 
  QLWQLQ = DEG_GOallontology.QLWQLQ.01_mergetermdef, 
  FLWMLW = DEG_GOallontology.FLWMLW.01_mergetermdef, 
  FLQMLQ = DEG_GOallontology.FLQMLQ.01_mergetermdef, 
  FLQQLQ = DEG_GOallontology.FLQQLQ.01_mergetermdef, 
  MLQQLQ = DEG_GOallontology.MLQQLQ.01_mergetermdef
)


write_xlsx(GOlists_AllOntology_TD, path = "Results/AllAnalyses_TDincluded.xlsx", col_names = TRUE)

```


```{r saveTOPgodata}
save(DEgenes.FLWFLQ.GO.bp, DEgenes.FLWFLQ.GO.mf, DEgenes.FLWFLQ.GO.cc, DEgenes.MLWMLQ.GO.bp, DEgenes.MLWMLQ.GO.mf, DEgenes.MLWMLQ.GO.cc, DEgenes.QLWQLQ.GO.bp, DEgenes.QLWQLQ.GO.mf, DEgenes.QLWQLQ.GO.cc, DEgenes.FLWMLW.GO.bp, DEgenes.FLWMLW.GO.mf, DEgenes.FLWMLW.GO.cc, DEgenes.FLQMLQ.GO.bp, DEgenes.FLQMLQ.GO.mf, DEgenes.FLQMLQ.GO.cc, DEgenes.FLQQLQ.GO.bp, DEgenes.FLQQLQ.GO.mf, DEgenes.FLQQLQ.GO.cc, DEgenes.MLQQLQ.GO.bp, DEgenes.MLQQLQ.GO.mf, DEgenes.MLQQLQ.GO.cc, file = "Results/GeneTables/TopGOobjects.Rdata")


```








# Rerun using upregulated and downregulate DEGs
In this analysis, we want to separate out DEGs into upregulated and downregulated groups prior to running the geneset enrichment analyses.

## Log Fold Change Scores
```{r upregulatedsplit}
##no DEGs up for FLWMLW, FLWQLW, MLQQLW, MLQQLQ


#FLWFLQ
sig.FLWFLQ <- subset(DEG.data.FLWFLQ, DEG.data.FLWFLQ$significant=="Significant")
str(sig.FLWFLQ)
  ##Make numeric
  sig.FLWFLQ[ , 1:6] <- lapply(sig.FLWFLQ[ , 1:6], as.numeric)
DEgenes_UP.FLWFLQ <- rownames(sig.FLWFLQ[sig.FLWFLQ$log2FoldChange > 1.5,])


######DEgenes_LFC.FLWFLQ <- rownames(sig.FLWFLQ[sig.FLWFLQ$log2FoldChange > 1.5,])



#these will be the same moving forward; so making numeric

#MLWMLQ
sig.MLWMLQ <- subset(DEG.data.MLWMLQ, DEG.data.MLWMLQ$significant=="Significant")
  ##Make numeric
  sig.MLWMLQ[ , 1:6] <- lapply(sig.MLWMLQ[ , 1:6], as.numeric)
DEgenes_UP.MLWMLQ <- rownames(sig.MLWMLQ[sig.MLWMLQ$log2FoldChange > 1.5,])

#QLWQLQ
sig.QLWQLQ <- subset(DEG.data.QLWQLQ, DEG.data.QLWQLQ$significant=="Significant")
  ##Make numeric
  sig.QLWQLQ[ , 1:6] <- lapply(sig.QLWQLQ[ , 1:6], as.numeric)
DEgenes_UP.QLWQLQ <- rownames(sig.QLWQLQ[sig.QLWQLQ$log2FoldChange > 1.5,])


#FLQMLQ
sig.FLQMLQ <- subset(DEG.data.FLQMLQ, DEG.data.FLQMLQ$significant=="Significant")
  ##Make numeric
  sig.FLQMLQ[ , 1:6] <- lapply(sig.FLQMLQ[ , 1:6], as.numeric)
DEgenes_UP.FLQMLQ <- rownames(sig.FLQMLQ[sig.FLQMLQ$log2FoldChange > 1.5,])



#FLQQLQ
sig.FLQQLQ <- subset(DEG.data.FLQQLQ, DEG.data.FLQQLQ$significant=="Significant")
  ##Make numeric
  sig.FLQQLQ[ , 1:6] <- lapply(sig.FLQQLQ[ , 1:6], as.numeric)
DEgenes_UP.FLQQLQ <- rownames(sig.FLQQLQ[sig.FLQQLQ$log2FoldChange > 1.5,])

```


## Downregulated genes only
```{r downregulatedsplit}
##no DEGs up for FLWQLW, MLQQLW, MLQQLQ  
#MLQQLQ only has 1 downregulated gene, so not running a go on that.


#FLWFLQ
DEgenes_DOWN.FLWFLQ <- rownames(sig.FLWFLQ[sig.FLWFLQ$log2FoldChange < -1.5,])

#MLWMLQ
DEgenes_DOWN.MLWMLQ <- rownames(sig.MLWMLQ[sig.MLWMLQ$log2FoldChange < -1.5,])

#QLWQLQ
DEgenes_DOWN.QLWQLQ <- rownames(sig.QLWQLQ[sig.QLWQLQ$log2FoldChange < -1.5,])

#FLQMLQ
DEgenes_DOWN.FLQMLQ <- rownames(sig.FLQMLQ[sig.FLQMLQ$log2FoldChange < -1.5,])

#FLQQLQ
DEgenes_DOWN.FLQQLQ <- rownames(sig.FLQQLQ[sig.FLQQLQ$log2FoldChange < -1.5,])


#FLWMLW
sig.FLWMLW <- subset(DEG.data.FLWMLW, DEG.data.FLWMLW$significant=="Significant")
  ##Make numeric
  sig.FLWMLW[ , 1:6] <- lapply(sig.FLWMLW[ , 1:6], as.numeric)
DEgenes_DOWN.FLWMLW <- rownames(sig.FLWMLW[sig.FLWMLW$log2FoldChange < -1.5,])

```


### FLWFLQ
#### Upregulated
#### BP
```{r BP_DEG_FLWFLQ_up}
GeneList_FLWFLQ_up <- as.integer(geneUniverse %in% DEgenes_UP.FLWFLQ)
names(GeneList_FLWFLQ_up) <- geneUniverse
LFCgenes <- function(x) x == 1

#verifying number of upreg genes with LFC < 1.5
table(GeneList_FLWFLQ_up)
#verifying the function
table(LFCgenes(GeneList_FLWFLQ_up))


DEgenes.FLWFLQ_UP.GO.bp <- new("topGOdata",
                description = paste("DEGUpGo", "BP", "FLWFLQ", sep = "_"),
                ontology = "BP",
                # all upregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLWFLQ_up, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Run enrichment test using Fisher's exact test
DEG.FLWFLQ_UP.bp_result_Fisher <- runTest(DEgenes.FLWFLQ_UP.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.FLWFLQ_UP.bp_result_Fisher

DEG.UP_BP.FLWFLQ.table <- GenTable(DEgenes.FLWFLQ_UP.GO.bp, 
                                   raw.p.value = DEG.FLWFLQ_UP.bp_result_Fisher, 
                                   topNodes = 50, 
                                   numChar = 120)

write.csv(DEG.UP_BP.FLWFLQ.table, "Results/UpDownSplit/UP/byontology/FLWFLQ_UP_BPTerms.csv")
```




#### MF
```{r mf_DEG_FLWFLQ_up}
DEgenes.FLWFLQ_UP.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "FLWFLQ", sep = "_"),
                ontology = "MF",
                # all upregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLWFLQ_up, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLWFLQ_UP.mf_result_fisher <- runTest(DEgenes.FLWFLQ_UP.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.FLWFLQ_UP.mf_result_fisher

DEG.UP_MF.FLWFLQ.table <- GenTable(DEgenes.FLWFLQ_UP.GO.mf, 
                                   raw.p.value = DEG.FLWFLQ_UP.mf_result_fisher, 
                                   topNodes = 50, 
                                   numChar = 120)

write.csv(DEG.UP_MF.FLWFLQ.table, "Results/UpDownSplit/UP/byontology/FLWFLQ_UP_MFTerms.csv")
```



#### CC
```{r cc_DEG_FLWFLQ_up}
DEgenes.FLWFLQ_UP.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "FLWFLQ", sep = "_"),
                ontology = "CC",
                # all upregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLWFLQ_up, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLWFLQ_UP.cc_result_fisher <- runTest(DEgenes.FLWFLQ_UP.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.FLWFLQ_UP.cc_result_fisher

DEG.UP_CC.FLWFLQ.table <- GenTable(DEgenes.FLWFLQ_UP.GO.cc, 
                                   raw.p.value = DEG.FLWFLQ_UP.cc_result_fisher, 
                                   topNodes = 50, 
                                   numChar = 120)

write.csv(DEG.UP_CC.FLWFLQ.table, "Results/UpDownSplit/UP/byontology/FLWFLQ_UP_CCTerms.csv")
```

```{r FLWFLQlargetable.UP}
#ALL
DEG.UP_BP.FLWFLQ.table <- cbind(DEG.UP_BP.FLWFLQ.table, ontology=rep("BP", length(DEG.UP_BP.FLWFLQ.table$GO.ID)))
DEG.UP_MF.FLWFLQ.table <- cbind(DEG.UP_MF.FLWFLQ.table, ontology=rep("MF", length(DEG.UP_MF.FLWFLQ.table$GO.ID)))
DEG.UP_CC.FLWFLQ.table <- cbind(DEG.UP_CC.FLWFLQ.table, ontology=rep("CC", length(DEG.UP_CC.FLWFLQ.table$GO.ID)))

DEG_GOallontology.FLWFLQ.UPregulated <-rbind(DEG.UP_BP.FLWFLQ.table, DEG.UP_MF.FLWFLQ.table, DEG.UP_CC.FLWFLQ.table)
write.csv(DEG_GOallontology.FLWFLQ.UPregulated, "Results/UpDownSplit/UP/ALL/DEGs_GOallontology_FLWFLQ_upregulated.csv")


#Less than .01
DEG.UP_BP.FLWFLQ.table.01 <- DEG.UP_BP.FLWFLQ.table[DEG.UP_BP.FLWFLQ.table$raw.p.value < 0.01,]
DEG.UP_MF.FLWFLQ.table.01 <- DEG.UP_MF.FLWFLQ.table[DEG.UP_MF.FLWFLQ.table$raw.p.value < 0.01,]
DEG.UP_CC.FLWFLQ.table.01 <- DEG.UP_CC.FLWFLQ.table[DEG.UP_CC.FLWFLQ.table$raw.p.value < 0.01,]

DEG_GOallontology.FLWFLQ.UPregulated.01 <-rbind(DEG.UP_BP.FLWFLQ.table.01, DEG.UP_MF.FLWFLQ.table.01, DEG.UP_CC.FLWFLQ.table.01)
write.csv(DEG_GOallontology.FLWFLQ.UPregulated.01, "Results/UpDownSplit/point1percent/DEGs_GOallontology_FLWFLQ_upregulated.01.csv")
```


#### Downregulated
#### BP
```{r BP_DEG_FLWFLQ_down}
GeneList_FLWFLQ_down <- as.integer(geneUniverse %in% DEgenes_DOWN.FLWFLQ)
names(GeneList_FLWFLQ_down) <- geneUniverse

#verifying number of upreg genes with LFC < -1.5
table(GeneList_FLWFLQ_down)
#verifying the function
table(LFCgenes(GeneList_FLWFLQ_down))

DEgenes.FLWFLQ_DOWN.GO.bp <- new("topGOdata",
                description = paste("DEGDownGo", "BP", "FLWFLQ", sep = "_"),
                ontology = "BP",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLWFLQ_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLWFLQ_DOWN.bp_result_fisher <- runTest(DEgenes.FLWFLQ_DOWN.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.FLWFLQ_DOWN.bp_result_fisher

DEG.DOWN_BP.FLWFLQ.table <- GenTable(DEgenes.FLWFLQ_DOWN.GO.bp, 
                                     raw.p.value = DEG.FLWFLQ_DOWN.bp_result_fisher, 
                                     topNodes = 50,
                                     numChar = 120)

write.csv(DEG.DOWN_BP.FLWFLQ.table, "Results/UpDownSplit/DOWN/byontology/FLWFLQ_DOWN_BPTerms.csv")
```


#### MF
```{r mf_DEG_FLWFLQ_down}
DEgenes.FLWFLQ_DOWN.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "FLWFLQ", sep = "_"),
                ontology = "MF",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLWFLQ_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLWFLQ_DOWN.mf_result_fisher <- runTest(DEgenes.FLWFLQ_DOWN.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.FLWFLQ_DOWN.mf_result_fisher

DEG.DOWN_MF.FLWFLQ.table <- GenTable(DEgenes.FLWFLQ_DOWN.GO.mf, 
                                     raw.p.value = DEG.FLWFLQ_DOWN.mf_result_fisher, 
                                     topNodes = 50, 
                                     numChar = 120)

write.csv(DEG.DOWN_MF.FLWFLQ.table, "Results/UpDownSplit/DOWN/byontology/FLWFLQ_DOWN_MFTerms.csv")
```



#### CC
```{r cc_DEG_FLWFLQ_down}
DEgenes.FLWFLQ_DOWN.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "FLWFLQ", sep = "_"),
                ontology = "CC",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLWFLQ_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLWFLQ_DOWN.cc_result_fisher <- runTest(DEgenes.FLWFLQ_DOWN.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.FLWFLQ_DOWN.cc_result_fisher

DEG.DOWN_CC.FLWFLQ.table <- GenTable(DEgenes.FLWFLQ_DOWN.GO.cc, 
                                     raw.p.value = DEG.FLWFLQ_DOWN.cc_result_fisher, 
                                     topNodes = 50,
                                     numChar = 120)

write.csv(DEG.DOWN_CC.FLWFLQ.table, "Results/UpDownSplit/DOWN/byontology/FLWFLQ_DOWN_CCTerms.csv")
```

```{r FLWFLQlargetable.DOWN}
#ALL
DEG.DOWN_BP.FLWFLQ.table <- cbind(DEG.DOWN_BP.FLWFLQ.table, ontology=rep("BP", length(DEG.DOWN_BP.FLWFLQ.table$GO.ID)))
DEG.DOWN_MF.FLWFLQ.table <- cbind(DEG.DOWN_MF.FLWFLQ.table, ontology=rep("MF", length(DEG.DOWN_MF.FLWFLQ.table$GO.ID)))
DEG.DOWN_CC.FLWFLQ.table <- cbind(DEG.DOWN_CC.FLWFLQ.table, ontology=rep("CC", length(DEG.DOWN_CC.FLWFLQ.table$GO.ID)))

DEG_GOallontology.FLWFLQ.DOWNregulated <-rbind(DEG.DOWN_BP.FLWFLQ.table, DEG.DOWN_MF.FLWFLQ.table, DEG.DOWN_CC.FLWFLQ.table)
write.csv(DEG_GOallontology.FLWFLQ.DOWNregulated, "Results/UpDownSplit/DOWN/ALL/DEGs_GOallontology_FLWFLQ_downregulated.csv")


#Less than .01
DEG.DOWN_BP.FLWFLQ.table.01 <- DEG.DOWN_BP.FLWFLQ.table[DEG.DOWN_BP.FLWFLQ.table$raw.p.value < 0.01,]
DEG.DOWN_MF.FLWFLQ.table.01 <- DEG.DOWN_MF.FLWFLQ.table[DEG.DOWN_MF.FLWFLQ.table$raw.p.value < 0.01,]
DEG.DOWN_CC.FLWFLQ.table.01 <- DEG.DOWN_CC.FLWFLQ.table[DEG.DOWN_CC.FLWFLQ.table$raw.p.value < 0.01,]

DEG_GOallontology.FLWFLQ.DOWNregulated.01 <-rbind(DEG.DOWN_BP.FLWFLQ.table.01, DEG.DOWN_MF.FLWFLQ.table.01, DEG.DOWN_CC.FLWFLQ.table.01)
write.csv(DEG_GOallontology.FLWFLQ.DOWNregulated.01, "Results/UpDownSplit/point1percent/DEGs_GOallontology_FLWFLQ_downregulated.01.csv")
```



```{r extractgenesforGOs.FLWFLQ}
#get results table
results.FLWFLQ <- list(BP.ALL = DEG_BP.FLWFLQ.table,
  BP.UP = DEG.UP_BP.FLWFLQ.table, 
  BP.DOWN = DEG.DOWN_BP.FLWFLQ.table,
  MF.ALL = DEG_MF.FLWFLQ.table,
  MF.UP = DEG.UP_MF.FLWFLQ.table, 
  MF.DOWN = DEG.DOWN_MF.FLWFLQ.table,
  CC.ALL = DEG_CC.FLWFLQ.table,
  CC.UP = DEG.UP_CC.FLWFLQ.table, 
  CC.DOWN = DEG.DOWN_CC.FLWFLQ.table
)
  
#get a list of the godata objects
GOdata_list.FLWFLQ <- list(
  BP.ALL = DEgenes.FLWFLQ.GO.bp,
  BP.UP = DEgenes.FLWFLQ_UP.GO.bp, 
  BP.DOWN = DEgenes.FLWFLQ_DOWN.GO.bp,
  MF.ALL = DEgenes.FLWFLQ.GO.mf,
  MF.UP = DEgenes.FLWFLQ_UP.GO.mf, 
  MF.DOWN = DEgenes.FLWFLQ_DOWN.GO.mf,
  CC.ALL = DEgenes.FLWFLQ.GO.cc,
  CC.UP = DEgenes.FLWFLQ_UP.GO.cc, 
  CC.DOWN = DEgenes.FLWFLQ_DOWN.GO.cc
)

#geneID Extraction

# Step 1: Filter significant GO terms (p < 0.01), handling scientific notation
sig.terms_list.FLWFLQ <- lapply(results.FLWFLQ, function(tbl) {
    # Ensure p-values are numeric (handles cases where they are character strings like "1.2e-5")
    pvals <- as.numeric(tbl$raw.p.value)
    
    # Return GO.IDs with p < 0.01
    tbl$GO.ID[which(pvals < 0.01)]
})

# Step 2: Extract significant genes per GO term from each GOdata object
significant_genes_per_GO.FLWFLQ <- mapply(function(GOdata, go_ids) {
    # Get genes associated with each GO term
    genes_per_GO <- genesInTerm(object = GOdata, whichGO = go_ids)
    # Get significant genes from the GOdata object
    sigGenesSet <- sigGenes(GOdata)
    # Intersect genes in each GO term with the significant genes
    lapply(genes_per_GO, function(genes) intersect(genes, sigGenesSet))
  },
GOdata_list.FLWFLQ,
sig.terms_list.FLWFLQ,
SIMPLIFY = FALSE)


# #flatten that B
# # Flatten the nested list into a tidy data frame
# df_significant_genes_per_GO <- map2_df(
#   significant_genes_per_GO,
#   names(significant_genes_per_GO),  # Corresponds to BP.ALL, BP.UP, BP.DOWN
#   function(go_list, group_name) {
#     tibble(
#       GO.ID = names(go_list),
#       significant_genes = go_list,
#       group = group_name
#     ) %>%
#     unnest_longer(significant_genes)  # Expands each gene to its own row
#   }
# )
# 
# 
# 
# write_xlsx(df_significant_genes_per_GO, "Results/GenesPerGO/significant_genes_per_GO_FLWFLQ.BP.xlsx")




#plus terms
# Make a named vector of GO terms (Term names), per group
go_term_names_list.FLWFLQ <- lapply(results.FLWFLQ, function(tbl) {
  setNames(tbl$Term, tbl$GO.ID)
})

# tidy_gene_tables <- map2(
#   significant_genes_per_GO,
#   names(significant_genes_per_GO),
#   function(go_list, group_name) {
#     # Create base table with GO.ID and genes
#     df <- tibble(
#       GO.ID = names(go_list),
#       significant_genes = go_list
#     ) %>%
#     unnest_longer(significant_genes)
# 
#     # Add GO term names using the appropriate lookup table
#     df$Term <- go_term_names_list[[group_name]][df$GO.ID]
# 
#     # Add group label
#     df$Group <- group_name
# 
#     # Reorder columns using base R or dplyr
#     df <- df[, c("Group", "GO.ID", "Term", "significant_genes")]  # <- fixed here
# 
#     return(df)
#   }
# )





####This is the function that will be used across PW comparisons

write_sig_genes_per_GO_to_excel <- function(sig_gene_list, go_term_name_list, output_file) {
  
  # Build a list of tidy data frames (one per group/sheet)
  tidy_gene_tables <- purrr::map2(
    sig_gene_list,
    names(sig_gene_list),
    function(go_list, group_name) {
      # Create base table with GO.ID and genes
      df <- tibble(
        GO.ID = names(go_list),
        significant_genes = go_list
      ) %>%
      unnest_longer(significant_genes)
      
      # Add GO term names from lookup
      df$Term <- go_term_name_list[[group_name]][df$GO.ID]
      
      # Add group and ontology type
      df$Group <- group_name
      df$Ontology <- toupper(stringr::str_extract(group_name, "^[A-Za-z]+"))  # Extract BP, MF, CC

      # Reorder columns
      df <- df[, c("Ontology", "Group", "GO.ID", "Term", "significant_genes")]
      
      return(df)
    }
  )
  
  # Create Excel workbook
  wb <- createWorkbook()
  
  # Add a sheet for each group
  purrr::walk2(tidy_gene_tables, names(tidy_gene_tables), function(df, sheet_name) {
    addWorksheet(wb, sheet_name)
    writeData(wb, sheet = sheet_name, df)
  })
  
  # Save workbook
  saveWorkbook(wb, output_file, overwrite = TRUE)
}


write_sig_genes_per_GO_to_excel(
  sig_gene_list = significant_genes_per_GO.FLWFLQ,
  go_term_name_list = go_term_names_list.FLWFLQ,
  output_file = "Results/GenesPerGO/FLWFLQ_significant_genes_per_GO.xlsx"
)



#Verify the overlap of genes between all, up, down
#BP
genesFLWFLQ_BP_ALL <- unique(unlist(significant_genes_per_GO.FLWFLQ$BP.ALL))
genesFLWFLQ_BP_UP <- unique(unlist(significant_genes_per_GO.FLWFLQ$BP.UP))
genesFLWFLQ_BP_DOWN <- unique(unlist(significant_genes_per_GO.FLWFLQ$BP.DOWN))

  #OVERLAP
  intersect(genesFLWFLQ_BP_ALL, genesFLWFLQ_BP_UP)
  table(genesFLWFLQ_BP_UP %in% genesFLWFLQ_BP_ALL)
  table(genesFLWFLQ_BP_UP %in% names(DEG.genelist.FLWFLQ))
  
  intersect(genesFLWFLQ_BP_ALL, genesFLWFLQ_BP_DOWN)
  table(genesFLWFLQ_BP_DOWN %in% genesFLWFLQ_BP_ALL)
  table(genesFLWFLQ_BP_DOWN %in% names(DEG.genelist.FLWFLQ))



#MF
genesFLWFLQ_MF_ALL <- unique(unlist(significant_genes_per_GO.FLWFLQ$MF.ALL))
genesFLWFLQ_MF_UP <- unique(unlist(significant_genes_per_GO.FLWFLQ$MF.UP)) #no ups
genesFLWFLQ_MF_DOWN <- unique(unlist(significant_genes_per_GO.FLWFLQ$MF.DOWN))

  #OVERLAP
  intersect(genesFLWFLQ_MF_ALL, genesFLWFLQ_MF_UP)
  table(genesFLWFLQ_MF_UP %in% genesFLWFLQ_MF_ALL)
  table(genesFLWFLQ_MF_UP %in% names(DEG.genelist.FLWFLQ))
  
  intersect(genesFLWFLQ_MF_ALL, genesFLWFLQ_MF_DOWN)
  table(genesFLWFLQ_MF_DOWN %in% genesFLWFLQ_MF_ALL)
  table(genesFLWFLQ_MF_DOWN %in% names(DEG.genelist.FLWFLQ))



#CC
genesFLWFLQ_CC_ALL <- unique(unlist(significant_genes_per_GO.FLWFLQ$CC.ALL))
genesFLWFLQ_CC_UP <- unique(unlist(significant_genes_per_GO.FLWFLQ$CC.UP))
genesFLWFLQ_CC_DOWN <- unique(unlist(significant_genes_per_GO.FLWFLQ$CC.DOWN))


  #OVERLAP
  intersect(genesFLWFLQ_CC_ALL, genesFLWFLQ_CC_UP)
  table(genesFLWFLQ_CC_UP %in% genesFLWFLQ_CC_ALL)
  table(genesFLWFLQ_CC_UP %in% names(DEG.genelist.FLWFLQ))
  
  intersect(genesFLWFLQ_CC_ALL, genesFLWFLQ_CC_DOWN)
  table(genesFLWFLQ_CC_DOWN %in% genesFLWFLQ_CC_ALL)
  table(genesFLWFLQ_CC_DOWN %in% names(DEG.genelist.FLWFLQ))


```



### MLWMLQ
#### Upregulated

#### BP
```{r BP_DEG_MLWMLQ_up}
GeneList_MLWMLQ_up <- as.integer(geneUniverse %in% DEgenes_UP.MLWMLQ)
names(GeneList_MLWMLQ_up) <- geneUniverse
LFCgenes <- function(x) x == 1

#verifying number of upreg genes with LFC < 1.5
table(GeneList_MLWMLQ_up)
#verifying the function
table(LFCgenes(GeneList_MLWMLQ_up))


DEgenes.MLWMLQ_UP.GO.bp <- new("topGOdata",
                description = paste("DEGUpGo", "BP", "MLWMLQ", sep = "_"),
                ontology = "BP",
                # all upregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_MLWMLQ_up, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Run enrichment test using Fisher's exact test
DEG.MLWMLQ_UP.bp_result_Fisher <- runTest(DEgenes.MLWMLQ_UP.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.MLWMLQ_UP.bp_result_Fisher

DEG.UP_BP.MLWMLQ.table <- GenTable(DEgenes.MLWMLQ_UP.GO.bp, 
                                   raw.p.value = DEG.MLWMLQ_UP.bp_result_Fisher, 
                                   topNodes = 50, 
                                   numChar = 120)

write.csv(DEG.UP_BP.MLWMLQ.table, "Results/UpDownSplit/UP/byontology/MLWMLQ_UP_BPTerms.csv")
```




#### MF
```{r mf_DEG_MLWMLQ_up}
DEgenes.MLWMLQ_UP.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "MLWMLQ", sep = "_"),
                ontology = "MF",
                # all upregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_MLWMLQ_up, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.MLWMLQ_UP.mf_result_fisher <- runTest(DEgenes.MLWMLQ_UP.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.MLWMLQ_UP.mf_result_fisher

DEG.UP_MF.MLWMLQ.table <- GenTable(DEgenes.MLWMLQ_UP.GO.mf, 
                                   raw.p.value = DEG.MLWMLQ_UP.mf_result_fisher, 
                                   topNodes = 50, 
                                   numChar = 120)

write.csv(DEG.UP_MF.MLWMLQ.table, "Results/UpDownSplit/UP/byontology/MLWMLQ_UP_MFTerms.csv")
```



#### CC
```{r cc_DEG_MLWMLQ_up}
DEgenes.MLWMLQ_UP.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "MLWMLQ", sep = "_"),
                ontology = "CC",
                # all upregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_MLWMLQ_up, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.MLWMLQ_UP.cc_result_fisher <- runTest(DEgenes.MLWMLQ_UP.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.MLWMLQ_UP.cc_result_fisher

DEG.UP_CC.MLWMLQ.table <- GenTable(DEgenes.MLWMLQ_UP.GO.cc, 
                                   raw.p.value = DEG.MLWMLQ_UP.cc_result_fisher, 
                                   topNodes = 50, 
                                   numChar = 120)

write.csv(DEG.UP_CC.MLWMLQ.table, "Results/UpDownSplit/UP/byontology/MLWMLQ_UP_CCTerms.csv")
```

```{r MLWMLQlargetable.UP}
#ALL
DEG.UP_BP.MLWMLQ.table <- cbind(DEG.UP_BP.MLWMLQ.table, ontology=rep("BP", length(DEG.UP_BP.MLWMLQ.table$GO.ID)))
DEG.UP_MF.MLWMLQ.table <- cbind(DEG.UP_MF.MLWMLQ.table, ontology=rep("MF", length(DEG.UP_MF.MLWMLQ.table$GO.ID)))
DEG.UP_CC.MLWMLQ.table <- cbind(DEG.UP_CC.MLWMLQ.table, ontology=rep("CC", length(DEG.UP_CC.MLWMLQ.table$GO.ID)))

DEG_GOallontology.MLWMLQ.UPregulated <-rbind(DEG.UP_BP.MLWMLQ.table, DEG.UP_MF.MLWMLQ.table, DEG.UP_CC.MLWMLQ.table)
write.csv(DEG_GOallontology.MLWMLQ.UPregulated, "Results/UpDownSplit/UP/ALL/DEGs_GOallontology_MLWMLQ_upregulated.csv")


#Less than .01
DEG.UP_BP.MLWMLQ.table.01 <- DEG.UP_BP.MLWMLQ.table[DEG.UP_BP.MLWMLQ.table$raw.p.value < 0.01,]
DEG.UP_MF.MLWMLQ.table.01 <- DEG.UP_MF.MLWMLQ.table[DEG.UP_MF.MLWMLQ.table$raw.p.value < 0.01,]
DEG.UP_CC.MLWMLQ.table.01 <- DEG.UP_CC.MLWMLQ.table[DEG.UP_CC.MLWMLQ.table$raw.p.value < 0.01,]

DEG_GOallontology.MLWMLQ.UPregulated.01 <-rbind(DEG.UP_BP.MLWMLQ.table.01, DEG.UP_MF.MLWMLQ.table.01, DEG.UP_CC.MLWMLQ.table.01)
write.csv(DEG_GOallontology.MLWMLQ.UPregulated.01, "Results/UpDownSplit/point1percent/DEGs_GOallontology_MLWMLQ_upregulated.01.csv")
```


#### Downregulated
#### BP
```{r BP_DEG_MLWMLQ_down}
GeneList_MLWMLQ_down <- as.integer(geneUniverse %in% DEgenes_DOWN.MLWMLQ)
names(GeneList_MLWMLQ_down) <- geneUniverse

#verifying number of upreg genes with LFC < -1.5
table(GeneList_MLWMLQ_down)
#verifying the function
table(LFCgenes(GeneList_MLWMLQ_down))

DEgenes.MLWMLQ_DOWN.GO.bp <- new("topGOdata",
                description = paste("DEGDownGo", "BP", "MLWMLQ", sep = "_"),
                ontology = "BP",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_MLWMLQ_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.MLWMLQ_DOWN.bp_result_fisher <- runTest(DEgenes.MLWMLQ_DOWN.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.MLWMLQ_DOWN.bp_result_fisher

DEG.DOWN_BP.MLWMLQ.table <- GenTable(DEgenes.MLWMLQ_DOWN.GO.bp, 
                                     raw.p.value = DEG.MLWMLQ_DOWN.bp_result_fisher, 
                                     topNodes = 50,
                                     numChar = 120)

write.csv(DEG.DOWN_BP.MLWMLQ.table, "Results/UpDownSplit/DOWN/byontology/MLWMLQ_DOWN_BPTerms.csv")
```


#### MF
```{r mf_DEG_MLWMLQ_down}
DEgenes.MLWMLQ_DOWN.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "MLWMLQ", sep = "_"),
                ontology = "MF",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_MLWMLQ_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.MLWMLQ_DOWN.mf_result_fisher <- runTest(DEgenes.MLWMLQ_DOWN.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.MLWMLQ_DOWN.mf_result_fisher

DEG.DOWN_MF.MLWMLQ.table <- GenTable(DEgenes.MLWMLQ_DOWN.GO.mf, 
                                     raw.p.value = DEG.MLWMLQ_DOWN.mf_result_fisher, 
                                     topNodes = 50, 
                                     numChar = 120)

write.csv(DEG.DOWN_MF.MLWMLQ.table, "Results/UpDownSplit/DOWN/byontology/MLWMLQ_DOWN_MFTerms.csv")
```



#### CC
```{r cc_DEG_MLWMLQ_down}
DEgenes.MLWMLQ_DOWN.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "MLWMLQ", sep = "_"),
                ontology = "CC",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_MLWMLQ_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.MLWMLQ_DOWN.cc_result_fisher <- runTest(DEgenes.MLWMLQ_DOWN.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.MLWMLQ_DOWN.cc_result_fisher

DEG.DOWN_CC.MLWMLQ.table <- GenTable(DEgenes.MLWMLQ_DOWN.GO.cc, 
                                     raw.p.value = DEG.MLWMLQ_DOWN.cc_result_fisher, 
                                     topNodes = 50,
                                     numChar = 120)

write.csv(DEG.DOWN_CC.MLWMLQ.table, "Results/UpDownSplit/DOWN/byontology/MLWMLQ_DOWN_CCTerms.csv")
```

```{r MLWMLQlargetable.DOWN}
#ALL
DEG.DOWN_BP.MLWMLQ.table <- cbind(DEG.DOWN_BP.MLWMLQ.table, ontology=rep("BP", length(DEG.DOWN_BP.MLWMLQ.table$GO.ID)))
DEG.DOWN_MF.MLWMLQ.table <- cbind(DEG.DOWN_MF.MLWMLQ.table, ontology=rep("MF", length(DEG.DOWN_MF.MLWMLQ.table$GO.ID)))
DEG.DOWN_CC.MLWMLQ.table <- cbind(DEG.DOWN_CC.MLWMLQ.table, ontology=rep("CC", length(DEG.DOWN_CC.MLWMLQ.table$GO.ID)))

DEG_GOallontology.MLWMLQ.DOWNregulated <-rbind(DEG.DOWN_BP.MLWMLQ.table, DEG.DOWN_MF.MLWMLQ.table, DEG.DOWN_CC.MLWMLQ.table)
write.csv(DEG_GOallontology.MLWMLQ.DOWNregulated, "Results/UpDownSplit/DOWN/ALL/DEGs_GOallontology_MLWMLQ_downregulated.csv")


#Less than .01
DEG.DOWN_BP.MLWMLQ.table.01 <- DEG.DOWN_BP.MLWMLQ.table[DEG.DOWN_BP.MLWMLQ.table$raw.p.value < 0.01,]
DEG.DOWN_MF.MLWMLQ.table.01 <- DEG.DOWN_MF.MLWMLQ.table[DEG.DOWN_MF.MLWMLQ.table$raw.p.value < 0.01,]
DEG.DOWN_CC.MLWMLQ.table.01 <- DEG.DOWN_CC.MLWMLQ.table[DEG.DOWN_CC.MLWMLQ.table$raw.p.value < 0.01,]

DEG_GOallontology.MLWMLQ.DOWNregulated.01 <-rbind(DEG.DOWN_BP.MLWMLQ.table.01, DEG.DOWN_MF.MLWMLQ.table.01, DEG.DOWN_CC.MLWMLQ.table.01)
write.csv(DEG_GOallontology.MLWMLQ.DOWNregulated.01, "Results/UpDownSplit/point1percent/DEGs_GOallontology_MLWMLQ_downregulated.01.csv")
```



```{r extractgenesforGOs.MLWMLQ}
#get results table
results.MLWMLQ <- list(BP.ALL = DEG_BP.MLWMLQ.table,
  BP.UP = DEG.UP_BP.MLWMLQ.table, 
  BP.DOWN = DEG.DOWN_BP.MLWMLQ.table,
  MF.ALL = DEG_MF.MLWMLQ.table,
  MF.UP = DEG.UP_MF.MLWMLQ.table, 
  MF.DOWN = DEG.DOWN_MF.MLWMLQ.table,
  CC.ALL = DEG_CC.MLWMLQ.table,
  CC.UP = DEG.UP_CC.MLWMLQ.table, 
  CC.DOWN = DEG.DOWN_CC.MLWMLQ.table
)
  
#get a list of the godata objects
GOdata_list.MLWMLQ <- list(
  BP.ALL = DEgenes.MLWMLQ.GO.bp,
  BP.UP = DEgenes.MLWMLQ_UP.GO.bp, 
  BP.DOWN = DEgenes.MLWMLQ_DOWN.GO.bp,
  MF.ALL = DEgenes.MLWMLQ.GO.mf,
  MF.UP = DEgenes.MLWMLQ_UP.GO.mf, 
  MF.DOWN = DEgenes.MLWMLQ_DOWN.GO.mf,
  CC.ALL = DEgenes.MLWMLQ.GO.cc,
  CC.UP = DEgenes.MLWMLQ_UP.GO.cc, 
  CC.DOWN = DEgenes.MLWMLQ_DOWN.GO.cc
)

#geneID Extraction

# Step 1: Filter significant GO terms (p < 0.01), handling scientific notation
sig.terms_list.MLWMLQ <- lapply(results.MLWMLQ, function(tbl) {
    # Ensure p-values are numeric (handles cases where they are character strings like "1.2e-5")
    pvals <- as.numeric(tbl$raw.p.value)
    
    # Return GO.IDs with p < 0.01
    tbl$GO.ID[which(pvals < 0.01)]
})

# Step 2: Extract significant genes per GO term from each GOdata object
significant_genes_per_GO.MLWMLQ <- mapply(function(GOdata, go_ids) {
    # Get genes associated with each GO term
    genes_per_GO <- genesInTerm(object = GOdata, whichGO = go_ids)
    # Get significant genes from the GOdata object
    sigGenesSet <- sigGenes(GOdata)
    # Intersect genes in each GO term with the significant genes
    lapply(genes_per_GO, function(genes) intersect(genes, sigGenesSet))
  },
GOdata_list.MLWMLQ,
sig.terms_list.MLWMLQ,
SIMPLIFY = FALSE)



#plus terms
# Make a named vector of GO terms (Term names), per group
go_term_names_list.MLWMLQ <- lapply(results.MLWMLQ, function(tbl) {
  setNames(tbl$Term, tbl$GO.ID)
})



####This is the function that will be used across PW comparisons
#write_sig_genes_per_GO_to_excel()


write_sig_genes_per_GO_to_excel(
  sig_gene_list = significant_genes_per_GO.MLWMLQ,
  go_term_name_list = go_term_names_list.MLWMLQ,
  output_file = "Results/GenesPerGO/MLWMLQ_significant_genes_per_GO.xlsx"
)



#Verify the overlap of genes between all, up, down
#BP
genesMLWMLQ_BP_ALL <- unique(unlist(significant_genes_per_GO.MLWMLQ$BP.ALL))
genesMLWMLQ_BP_UP <- unique(unlist(significant_genes_per_GO.MLWMLQ$BP.UP))
genesMLWMLQ_BP_DOWN <- unique(unlist(significant_genes_per_GO.MLWMLQ$BP.DOWN))

  #OVERLAP
  intersect(genesMLWMLQ_BP_ALL, genesMLWMLQ_BP_UP)
  table(genesMLWMLQ_BP_UP %in% genesMLWMLQ_BP_ALL)
  table(genesMLWMLQ_BP_UP %in% names(DEG.genelist.MLWMLQ))
  
  intersect(genesMLWMLQ_BP_ALL, genesMLWMLQ_BP_DOWN)
  table(genesMLWMLQ_BP_DOWN %in% genesMLWMLQ_BP_ALL)
  table(genesMLWMLQ_BP_DOWN %in% names(DEG.genelist.MLWMLQ))



#MF
genesMLWMLQ_MF_ALL <- unique(unlist(significant_genes_per_GO.MLWMLQ$MF.ALL))
genesMLWMLQ_MF_UP <- unique(unlist(significant_genes_per_GO.MLWMLQ$MF.UP)) #no ups
genesMLWMLQ_MF_DOWN <- unique(unlist(significant_genes_per_GO.MLWMLQ$MF.DOWN))

  #OVERLAP
  intersect(genesMLWMLQ_MF_ALL, genesMLWMLQ_MF_UP)
  table(genesMLWMLQ_MF_UP %in% genesMLWMLQ_MF_ALL)
  table(genesMLWMLQ_MF_UP %in% names(DEG.genelist.MLWMLQ))
  
  intersect(genesMLWMLQ_MF_ALL, genesMLWMLQ_MF_DOWN)
  table(genesMLWMLQ_MF_DOWN %in% genesMLWMLQ_MF_ALL)
  table(genesMLWMLQ_MF_DOWN %in% names(DEG.genelist.MLWMLQ))



#CC
genesMLWMLQ_CC_ALL <- unique(unlist(significant_genes_per_GO.MLWMLQ$CC.ALL))
genesMLWMLQ_CC_UP <- unique(unlist(significant_genes_per_GO.MLWMLQ$CC.UP))
genesMLWMLQ_CC_DOWN <- unique(unlist(significant_genes_per_GO.MLWMLQ$CC.DOWN))


  #OVERLAP
  intersect(genesMLWMLQ_CC_ALL, genesMLWMLQ_CC_UP)
  table(genesMLWMLQ_CC_UP %in% genesMLWMLQ_CC_ALL)
  table(genesMLWMLQ_CC_UP %in% names(DEG.genelist.MLWMLQ))
  
  intersect(genesMLWMLQ_CC_ALL, genesMLWMLQ_CC_DOWN)
  table(genesMLWMLQ_CC_DOWN %in% genesMLWMLQ_CC_ALL)
  table(genesMLWMLQ_CC_DOWN %in% names(DEG.genelist.MLWMLQ))


```






### QLWQLQ
#### Upregulated

#### BP
```{r BP_DEG_QLWQLQ_up}
GeneList_QLWQLQ_up <- as.integer(geneUniverse %in% DEgenes_UP.QLWQLQ)
names(GeneList_QLWQLQ_up) <- geneUniverse
LFCgenes <- function(x) x == 1

#verifying number of upreg genes with LFC < 1.5
table(GeneList_QLWQLQ_up)
#verifying the function
table(LFCgenes(GeneList_QLWQLQ_up))


DEgenes.QLWQLQ_UP.GO.bp <- new("topGOdata",
                description = paste("DEGUpGo", "BP", "QLWQLQ", sep = "_"),
                ontology = "BP",
                # all upregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_QLWQLQ_up, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Run enrichment test using Fisher's exact test
DEG.QLWQLQ_UP.bp_result_Fisher <- runTest(DEgenes.QLWQLQ_UP.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.QLWQLQ_UP.bp_result_Fisher

DEG.UP_BP.QLWQLQ.table <- GenTable(DEgenes.QLWQLQ_UP.GO.bp, 
                                   raw.p.value = DEG.QLWQLQ_UP.bp_result_Fisher, 
                                   topNodes = 50, 
                                   numChar = 120)

write.csv(DEG.UP_BP.QLWQLQ.table, "Results/UpDownSplit/UP/byontology/QLWQLQ_UP_BPTerms.csv")
```




#### MF
```{r mf_DEG_QLWQLQ_up}
DEgenes.QLWQLQ_UP.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "QLWQLQ", sep = "_"),
                ontology = "MF",
                # all upregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_QLWQLQ_up, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.QLWQLQ_UP.mf_result_fisher <- runTest(DEgenes.QLWQLQ_UP.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.QLWQLQ_UP.mf_result_fisher

DEG.UP_MF.QLWQLQ.table <- GenTable(DEgenes.QLWQLQ_UP.GO.mf, 
                                   raw.p.value = DEG.QLWQLQ_UP.mf_result_fisher, 
                                   topNodes = 50, 
                                   numChar = 120)

write.csv(DEG.UP_MF.QLWQLQ.table, "Results/UpDownSplit/UP/byontology/QLWQLQ_UP_MFTerms.csv")
```



#### CC
```{r cc_DEG_QLWQLQ_up}
DEgenes.QLWQLQ_UP.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "QLWQLQ", sep = "_"),
                ontology = "CC",
                # all upregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_QLWQLQ_up, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.QLWQLQ_UP.cc_result_fisher <- runTest(DEgenes.QLWQLQ_UP.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.QLWQLQ_UP.cc_result_fisher

DEG.UP_CC.QLWQLQ.table <- GenTable(DEgenes.QLWQLQ_UP.GO.cc, 
                                   raw.p.value = DEG.QLWQLQ_UP.cc_result_fisher, 
                                   topNodes = 50, 
                                   numChar = 120)

write.csv(DEG.UP_CC.QLWQLQ.table, "Results/UpDownSplit/UP/byontology/QLWQLQ_UP_CCTerms.csv")
```

```{r QLWQLQlargetable.UP}
#ALL
DEG.UP_BP.QLWQLQ.table <- cbind(DEG.UP_BP.QLWQLQ.table, ontology=rep("BP", length(DEG.UP_BP.QLWQLQ.table$GO.ID)))
DEG.UP_MF.QLWQLQ.table <- cbind(DEG.UP_MF.QLWQLQ.table, ontology=rep("MF", length(DEG.UP_MF.QLWQLQ.table$GO.ID)))
DEG.UP_CC.QLWQLQ.table <- cbind(DEG.UP_CC.QLWQLQ.table, ontology=rep("CC", length(DEG.UP_CC.QLWQLQ.table$GO.ID)))

DEG_GOallontology.QLWQLQ.UPregulated <-rbind(DEG.UP_BP.QLWQLQ.table, DEG.UP_MF.QLWQLQ.table, DEG.UP_CC.QLWQLQ.table)
write.csv(DEG_GOallontology.QLWQLQ.UPregulated, "Results/UpDownSplit/UP/ALL/DEGs_GOallontology_QLWQLQ_upregulated.csv")


#Less than .01
DEG.UP_BP.QLWQLQ.table.01 <- DEG.UP_BP.QLWQLQ.table[DEG.UP_BP.QLWQLQ.table$raw.p.value < 0.01,]
DEG.UP_MF.QLWQLQ.table.01 <- DEG.UP_MF.QLWQLQ.table[DEG.UP_MF.QLWQLQ.table$raw.p.value < 0.01,]
DEG.UP_CC.QLWQLQ.table.01 <- DEG.UP_CC.QLWQLQ.table[DEG.UP_CC.QLWQLQ.table$raw.p.value < 0.01,]

DEG_GOallontology.QLWQLQ.UPregulated.01 <-rbind(DEG.UP_BP.QLWQLQ.table.01, DEG.UP_MF.QLWQLQ.table.01, DEG.UP_CC.QLWQLQ.table.01)
write.csv(DEG_GOallontology.QLWQLQ.UPregulated.01, "Results/UpDownSplit/point1percent/DEGs_GOallontology_QLWQLQ_upregulated.01.csv")
```


#### Downregulated
#### BP
```{r BP_DEG_QLWQLQ_down}
GeneList_QLWQLQ_down <- as.integer(geneUniverse %in% DEgenes_DOWN.QLWQLQ)
names(GeneList_QLWQLQ_down) <- geneUniverse

#verifying number of upreg genes with LFC < -1.5
table(GeneList_QLWQLQ_down)
#verifying the function
table(LFCgenes(GeneList_QLWQLQ_down))

DEgenes.QLWQLQ_DOWN.GO.bp <- new("topGOdata",
                description = paste("DEGDownGo", "BP", "QLWQLQ", sep = "_"),
                ontology = "BP",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_QLWQLQ_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.QLWQLQ_DOWN.bp_result_fisher <- runTest(DEgenes.QLWQLQ_DOWN.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.QLWQLQ_DOWN.bp_result_fisher

DEG.DOWN_BP.QLWQLQ.table <- GenTable(DEgenes.QLWQLQ_DOWN.GO.bp, 
                                     raw.p.value = DEG.QLWQLQ_DOWN.bp_result_fisher, 
                                     topNodes = 50,
                                     numChar = 120)

write.csv(DEG.DOWN_BP.QLWQLQ.table, "Results/UpDownSplit/DOWN/byontology/QLWQLQ_DOWN_BPTerms.csv")
```


#### MF
```{r mf_DEG_QLWQLQ_down}
DEgenes.QLWQLQ_DOWN.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "QLWQLQ", sep = "_"),
                ontology = "MF",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_QLWQLQ_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.QLWQLQ_DOWN.mf_result_fisher <- runTest(DEgenes.QLWQLQ_DOWN.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.QLWQLQ_DOWN.mf_result_fisher

DEG.DOWN_MF.QLWQLQ.table <- GenTable(DEgenes.QLWQLQ_DOWN.GO.mf, 
                                     raw.p.value = DEG.QLWQLQ_DOWN.mf_result_fisher, 
                                     topNodes = 50, 
                                     numChar = 120)

write.csv(DEG.DOWN_MF.QLWQLQ.table, "Results/UpDownSplit/DOWN/byontology/QLWQLQ_DOWN_MFTerms.csv")
```



#### CC
```{r cc_DEG_QLWQLQ_down}
DEgenes.QLWQLQ_DOWN.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "QLWQLQ", sep = "_"),
                ontology = "CC",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_QLWQLQ_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.QLWQLQ_DOWN.cc_result_fisher <- runTest(DEgenes.QLWQLQ_DOWN.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.QLWQLQ_DOWN.cc_result_fisher

DEG.DOWN_CC.QLWQLQ.table <- GenTable(DEgenes.QLWQLQ_DOWN.GO.cc, 
                                     raw.p.value = DEG.QLWQLQ_DOWN.cc_result_fisher, 
                                     topNodes = 50,
                                     numChar = 120)

write.csv(DEG.DOWN_CC.QLWQLQ.table, "Results/UpDownSplit/DOWN/byontology/QLWQLQ_DOWN_CCTerms.csv")
```

```{r QLWQLQlargetable.DOWN}
#ALL
DEG.DOWN_BP.QLWQLQ.table <- cbind(DEG.DOWN_BP.QLWQLQ.table, ontology=rep("BP", length(DEG.DOWN_BP.QLWQLQ.table$GO.ID)))
DEG.DOWN_MF.QLWQLQ.table <- cbind(DEG.DOWN_MF.QLWQLQ.table, ontology=rep("MF", length(DEG.DOWN_MF.QLWQLQ.table$GO.ID)))
DEG.DOWN_CC.QLWQLQ.table <- cbind(DEG.DOWN_CC.QLWQLQ.table, ontology=rep("CC", length(DEG.DOWN_CC.QLWQLQ.table$GO.ID)))

DEG_GOallontology.QLWQLQ.DOWNregulated <-rbind(DEG.DOWN_BP.QLWQLQ.table, DEG.DOWN_MF.QLWQLQ.table, DEG.DOWN_CC.QLWQLQ.table)
write.csv(DEG_GOallontology.QLWQLQ.DOWNregulated, "Results/UpDownSplit/DOWN/ALL/DEGs_GOallontology_QLWQLQ_downregulated.csv")


#Less than .01
DEG.DOWN_BP.QLWQLQ.table.01 <- DEG.DOWN_BP.QLWQLQ.table[DEG.DOWN_BP.QLWQLQ.table$raw.p.value < 0.01,]
DEG.DOWN_MF.QLWQLQ.table.01 <- DEG.DOWN_MF.QLWQLQ.table[DEG.DOWN_MF.QLWQLQ.table$raw.p.value < 0.01,]
DEG.DOWN_CC.QLWQLQ.table.01 <- DEG.DOWN_CC.QLWQLQ.table[DEG.DOWN_CC.QLWQLQ.table$raw.p.value < 0.01,]

DEG_GOallontology.QLWQLQ.DOWNregulated.01 <-rbind(DEG.DOWN_BP.QLWQLQ.table.01, DEG.DOWN_MF.QLWQLQ.table.01, DEG.DOWN_CC.QLWQLQ.table.01)
write.csv(DEG_GOallontology.QLWQLQ.DOWNregulated.01, "Results/UpDownSplit/point1percent/DEGs_GOallontology_QLWQLQ_downregulated.01.csv")
```



```{r extractgenesforGOs.QLWQLQ}
#get results table
results.QLWQLQ <- list(BP.ALL = DEG_BP.QLWQLQ.table,
  BP.UP = DEG.UP_BP.QLWQLQ.table, 
  BP.DOWN = DEG.DOWN_BP.QLWQLQ.table,
  MF.ALL = DEG_MF.QLWQLQ.table,
  MF.UP = DEG.UP_MF.QLWQLQ.table, 
  MF.DOWN = DEG.DOWN_MF.QLWQLQ.table,
  CC.ALL = DEG_CC.QLWQLQ.table,
  CC.UP = DEG.UP_CC.QLWQLQ.table, 
  CC.DOWN = DEG.DOWN_CC.QLWQLQ.table
)
  
#get a list of the godata objects
GOdata_list.QLWQLQ <- list(
  BP.ALL = DEgenes.QLWQLQ.GO.bp,
  BP.UP = DEgenes.QLWQLQ_UP.GO.bp, 
  BP.DOWN = DEgenes.QLWQLQ_DOWN.GO.bp,
  MF.ALL = DEgenes.QLWQLQ.GO.mf,
  MF.UP = DEgenes.QLWQLQ_UP.GO.mf, 
  MF.DOWN = DEgenes.QLWQLQ_DOWN.GO.mf,
  CC.ALL = DEgenes.QLWQLQ.GO.cc,
  CC.UP = DEgenes.QLWQLQ_UP.GO.cc, 
  CC.DOWN = DEgenes.QLWQLQ_DOWN.GO.cc
)

#geneID Extraction

# Step 1: Filter significant GO terms (p < 0.01), handling scientific notation
sig.terms_list.QLWQLQ <- lapply(results.QLWQLQ, function(tbl) {
    # Ensure p-values are numeric (handles cases where they are character strings like "1.2e-5")
    pvals <- as.numeric(tbl$raw.p.value)
    
    # Return GO.IDs with p < 0.01
    tbl$GO.ID[which(pvals < 0.01)]
})

# Step 2: Extract significant genes per GO term from each GOdata object
significant_genes_per_GO.QLWQLQ <- mapply(function(GOdata, go_ids) {
    # Get genes associated with each GO term
    genes_per_GO <- genesInTerm(object = GOdata, whichGO = go_ids)
    # Get significant genes from the GOdata object
    sigGenesSet <- sigGenes(GOdata)
    # Intersect genes in each GO term with the significant genes
    lapply(genes_per_GO, function(genes) intersect(genes, sigGenesSet))
  },
GOdata_list.QLWQLQ,
sig.terms_list.QLWQLQ,
SIMPLIFY = FALSE)



#plus terms
# Make a named vector of GO terms (Term names), per group
go_term_names_list.QLWQLQ <- lapply(results.QLWQLQ, function(tbl) {
  setNames(tbl$Term, tbl$GO.ID)
})



####This is the function that will be used across PW comparisons
#write_sig_genes_per_GO_to_excel()


write_sig_genes_per_GO_to_excel(
  sig_gene_list = significant_genes_per_GO.QLWQLQ,
  go_term_name_list = go_term_names_list.QLWQLQ,
  output_file = "Results/GenesPerGO/QLWQLQ_significant_genes_per_GO.xlsx"
)



#Verify the overlap of genes between all, up, down
#BP
genesQLWQLQ_BP_ALL <- unique(unlist(significant_genes_per_GO.QLWQLQ$BP.ALL))
genesQLWQLQ_BP_UP <- unique(unlist(significant_genes_per_GO.QLWQLQ$BP.UP))
genesQLWQLQ_BP_DOWN <- unique(unlist(significant_genes_per_GO.QLWQLQ$BP.DOWN))

  #OVERLAP
  intersect(genesQLWQLQ_BP_ALL, genesQLWQLQ_BP_UP)
  table(genesQLWQLQ_BP_UP %in% genesQLWQLQ_BP_ALL)
  table(genesQLWQLQ_BP_UP %in% names(DEG.genelist.QLWQLQ))
  
  intersect(genesQLWQLQ_BP_ALL, genesQLWQLQ_BP_DOWN)
  table(genesQLWQLQ_BP_DOWN %in% genesQLWQLQ_BP_ALL)
  table(genesQLWQLQ_BP_DOWN %in% names(DEG.genelist.QLWQLQ))



#MF
genesQLWQLQ_MF_ALL <- unique(unlist(significant_genes_per_GO.QLWQLQ$MF.ALL))
genesQLWQLQ_MF_UP <- unique(unlist(significant_genes_per_GO.QLWQLQ$MF.UP))
genesQLWQLQ_MF_DOWN <- unique(unlist(significant_genes_per_GO.QLWQLQ$MF.DOWN))

  #OVERLAP
  intersect(genesQLWQLQ_MF_ALL, genesQLWQLQ_MF_UP)
  table(genesQLWQLQ_MF_UP %in% genesQLWQLQ_MF_ALL)
  table(genesQLWQLQ_MF_UP %in% names(DEG.genelist.QLWQLQ))
  
  intersect(genesQLWQLQ_MF_ALL, genesQLWQLQ_MF_DOWN)
  table(genesQLWQLQ_MF_DOWN %in% genesQLWQLQ_MF_ALL)
  table(genesQLWQLQ_MF_DOWN %in% names(DEG.genelist.QLWQLQ))



#CC
genesQLWQLQ_CC_ALL <- unique(unlist(significant_genes_per_GO.QLWQLQ$CC.ALL))
genesQLWQLQ_CC_UP <- unique(unlist(significant_genes_per_GO.QLWQLQ$CC.UP))
genesQLWQLQ_CC_DOWN <- unique(unlist(significant_genes_per_GO.QLWQLQ$CC.DOWN))


  #OVERLAP
  intersect(genesQLWQLQ_CC_ALL, genesQLWQLQ_CC_UP)
  table(genesQLWQLQ_CC_UP %in% genesQLWQLQ_CC_ALL)
  table(genesQLWQLQ_CC_UP %in% names(DEG.genelist.QLWQLQ))
  
  intersect(genesQLWQLQ_CC_ALL, genesQLWQLQ_CC_DOWN)
  table(genesQLWQLQ_CC_DOWN %in% genesQLWQLQ_CC_ALL)
  table(genesQLWQLQ_CC_DOWN %in% names(DEG.genelist.QLWQLQ))


```


### FLQMLQ
#### Upregulated

#### BP
```{r BP_DEG_FLQMLQ_up}
GeneList_FLQMLQ_up <- as.integer(geneUniverse %in% DEgenes_UP.FLQMLQ)
names(GeneList_FLQMLQ_up) <- geneUniverse
LFCgenes <- function(x) x == 1

#verifying number of upreg genes with LFC < 1.5
table(GeneList_FLQMLQ_up)
#verifying the function
table(LFCgenes(GeneList_FLQMLQ_up))


DEgenes.FLQMLQ_UP.GO.bp <- new("topGOdata",
                description = paste("DEGUpGo", "BP", "FLQMLQ", sep = "_"),
                ontology = "BP",
                # all upregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLQMLQ_up, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Run enrichment test using Fisher's exact test
DEG.FLQMLQ_UP.bp_result_Fisher <- runTest(DEgenes.FLQMLQ_UP.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.FLQMLQ_UP.bp_result_Fisher

DEG.UP_BP.FLQMLQ.table <- GenTable(DEgenes.FLQMLQ_UP.GO.bp, 
                                   raw.p.value = DEG.FLQMLQ_UP.bp_result_Fisher, 
                                   topNodes = 50, 
                                   numChar = 120)

write.csv(DEG.UP_BP.FLQMLQ.table, "Results/UpDownSplit/UP/byontology/FLQMLQ_UP_BPTerms.csv")
```




#### MF
```{r mf_DEG_FLQMLQ_up}
DEgenes.FLQMLQ_UP.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "FLQMLQ", sep = "_"),
                ontology = "MF",
                # all upregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLQMLQ_up, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLQMLQ_UP.mf_result_fisher <- runTest(DEgenes.FLQMLQ_UP.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.FLQMLQ_UP.mf_result_fisher

DEG.UP_MF.FLQMLQ.table <- GenTable(DEgenes.FLQMLQ_UP.GO.mf, 
                                   raw.p.value = DEG.FLQMLQ_UP.mf_result_fisher, 
                                   topNodes = 50, 
                                   numChar = 120)

write.csv(DEG.UP_MF.FLQMLQ.table, "Results/UpDownSplit/UP/byontology/FLQMLQ_UP_MFTerms.csv")
```



#### CC
```{r cc_DEG_FLQMLQ_up}
DEgenes.FLQMLQ_UP.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "FLQMLQ", sep = "_"),
                ontology = "CC",
                # all upregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLQMLQ_up, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLQMLQ_UP.cc_result_fisher <- runTest(DEgenes.FLQMLQ_UP.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.FLQMLQ_UP.cc_result_fisher

DEG.UP_CC.FLQMLQ.table <- GenTable(DEgenes.FLQMLQ_UP.GO.cc, 
                                   raw.p.value = DEG.FLQMLQ_UP.cc_result_fisher, 
                                   topNodes = 50, 
                                   numChar = 120)

write.csv(DEG.UP_CC.FLQMLQ.table, "Results/UpDownSplit/UP/byontology/FLQMLQ_UP_CCTerms.csv")
```

```{r FLQMLQlargetable.UP}
#ALL
DEG.UP_BP.FLQMLQ.table <- cbind(DEG.UP_BP.FLQMLQ.table, ontology=rep("BP", length(DEG.UP_BP.FLQMLQ.table$GO.ID)))
DEG.UP_MF.FLQMLQ.table <- cbind(DEG.UP_MF.FLQMLQ.table, ontology=rep("MF", length(DEG.UP_MF.FLQMLQ.table$GO.ID)))
DEG.UP_CC.FLQMLQ.table <- cbind(DEG.UP_CC.FLQMLQ.table, ontology=rep("CC", length(DEG.UP_CC.FLQMLQ.table$GO.ID)))

DEG_GOallontology.FLQMLQ.UPregulated <-rbind(DEG.UP_BP.FLQMLQ.table, DEG.UP_MF.FLQMLQ.table, DEG.UP_CC.FLQMLQ.table)
write.csv(DEG_GOallontology.FLQMLQ.UPregulated, "Results/UpDownSplit/UP/ALL/DEGs_GOallontology_FLQMLQ_upregulated.csv")


#Less than .01
DEG.UP_BP.FLQMLQ.table.01 <- DEG.UP_BP.FLQMLQ.table[DEG.UP_BP.FLQMLQ.table$raw.p.value < 0.01,]
DEG.UP_MF.FLQMLQ.table.01 <- DEG.UP_MF.FLQMLQ.table[DEG.UP_MF.FLQMLQ.table$raw.p.value < 0.01,]
DEG.UP_CC.FLQMLQ.table.01 <- DEG.UP_CC.FLQMLQ.table[DEG.UP_CC.FLQMLQ.table$raw.p.value < 0.01,]

DEG_GOallontology.FLQMLQ.UPregulated.01 <-rbind(DEG.UP_BP.FLQMLQ.table.01, DEG.UP_MF.FLQMLQ.table.01, DEG.UP_CC.FLQMLQ.table.01)
write.csv(DEG_GOallontology.FLQMLQ.UPregulated.01, "Results/UpDownSplit/point1percent/DEGs_GOallontology_FLQMLQ_upregulated.01.csv")
```


#### Downregulated
#### BP
```{r BP_DEG_FLQMLQ_down}
GeneList_FLQMLQ_down <- as.integer(geneUniverse %in% DEgenes_DOWN.FLQMLQ)
names(GeneList_FLQMLQ_down) <- geneUniverse

#verifying number of upreg genes with LFC < -1.5
table(GeneList_FLQMLQ_down)
#verifying the function
table(LFCgenes(GeneList_FLQMLQ_down))

DEgenes.FLQMLQ_DOWN.GO.bp <- new("topGOdata",
                description = paste("DEGDownGo", "BP", "FLQMLQ", sep = "_"),
                ontology = "BP",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLQMLQ_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLQMLQ_DOWN.bp_result_fisher <- runTest(DEgenes.FLQMLQ_DOWN.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.FLQMLQ_DOWN.bp_result_fisher

DEG.DOWN_BP.FLQMLQ.table <- GenTable(DEgenes.FLQMLQ_DOWN.GO.bp, 
                                     raw.p.value = DEG.FLQMLQ_DOWN.bp_result_fisher, 
                                     topNodes = 50,
                                     numChar = 120)

write.csv(DEG.DOWN_BP.FLQMLQ.table, "Results/UpDownSplit/DOWN/byontology/FLQMLQ_DOWN_BPTerms.csv")
```


#### MF
```{r mf_DEG_FLQMLQ_down}
DEgenes.FLQMLQ_DOWN.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "FLQMLQ", sep = "_"),
                ontology = "MF",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLQMLQ_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLQMLQ_DOWN.mf_result_fisher <- runTest(DEgenes.FLQMLQ_DOWN.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.FLQMLQ_DOWN.mf_result_fisher

DEG.DOWN_MF.FLQMLQ.table <- GenTable(DEgenes.FLQMLQ_DOWN.GO.mf, 
                                     raw.p.value = DEG.FLQMLQ_DOWN.mf_result_fisher, 
                                     topNodes = 50, 
                                     numChar = 120)

write.csv(DEG.DOWN_MF.FLQMLQ.table, "Results/UpDownSplit/DOWN/byontology/FLQMLQ_DOWN_MFTerms.csv")
```



#### CC
```{r cc_DEG_FLQMLQ_down}
DEgenes.FLQMLQ_DOWN.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "FLQMLQ", sep = "_"),
                ontology = "CC",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLQMLQ_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLQMLQ_DOWN.cc_result_fisher <- runTest(DEgenes.FLQMLQ_DOWN.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.FLQMLQ_DOWN.cc_result_fisher

DEG.DOWN_CC.FLQMLQ.table <- GenTable(DEgenes.FLQMLQ_DOWN.GO.cc, 
                                     raw.p.value = DEG.FLQMLQ_DOWN.cc_result_fisher, 
                                     topNodes = 50,
                                     numChar = 120)

write.csv(DEG.DOWN_CC.FLQMLQ.table, "Results/UpDownSplit/DOWN/byontology/FLQMLQ_DOWN_CCTerms.csv")
```

```{r FLQMLQlargetable.DOWN}
#ALL
DEG.DOWN_BP.FLQMLQ.table <- cbind(DEG.DOWN_BP.FLQMLQ.table, ontology=rep("BP", length(DEG.DOWN_BP.FLQMLQ.table$GO.ID)))
DEG.DOWN_MF.FLQMLQ.table <- cbind(DEG.DOWN_MF.FLQMLQ.table, ontology=rep("MF", length(DEG.DOWN_MF.FLQMLQ.table$GO.ID)))
DEG.DOWN_CC.FLQMLQ.table <- cbind(DEG.DOWN_CC.FLQMLQ.table, ontology=rep("CC", length(DEG.DOWN_CC.FLQMLQ.table$GO.ID)))

DEG_GOallontology.FLQMLQ.DOWNregulated <-rbind(DEG.DOWN_BP.FLQMLQ.table, DEG.DOWN_MF.FLQMLQ.table, DEG.DOWN_CC.FLQMLQ.table)
write.csv(DEG_GOallontology.FLQMLQ.DOWNregulated, "Results/UpDownSplit/DOWN/ALL/DEGs_GOallontology_FLQMLQ_downregulated.csv")


#Less than .01
DEG.DOWN_BP.FLQMLQ.table.01 <- DEG.DOWN_BP.FLQMLQ.table[DEG.DOWN_BP.FLQMLQ.table$raw.p.value < 0.01,]
DEG.DOWN_MF.FLQMLQ.table.01 <- DEG.DOWN_MF.FLQMLQ.table[DEG.DOWN_MF.FLQMLQ.table$raw.p.value < 0.01,]
DEG.DOWN_CC.FLQMLQ.table.01 <- DEG.DOWN_CC.FLQMLQ.table[DEG.DOWN_CC.FLQMLQ.table$raw.p.value < 0.01,]

DEG_GOallontology.FLQMLQ.DOWNregulated.01 <-rbind(DEG.DOWN_BP.FLQMLQ.table.01, DEG.DOWN_MF.FLQMLQ.table.01, DEG.DOWN_CC.FLQMLQ.table.01)
write.csv(DEG_GOallontology.FLQMLQ.DOWNregulated.01, "Results/UpDownSplit/point1percent/DEGs_GOallontology_FLQMLQ_downregulated.01.csv")
```



```{r extractgenesforGOs.FLQMLQ}
#get results table
results.FLQMLQ <- list(BP.ALL = DEG_BP.FLQMLQ.table,
  BP.UP = DEG.UP_BP.FLQMLQ.table, 
  BP.DOWN = DEG.DOWN_BP.FLQMLQ.table,
  MF.ALL = DEG_MF.FLQMLQ.table,
  MF.UP = DEG.UP_MF.FLQMLQ.table, 
  MF.DOWN = DEG.DOWN_MF.FLQMLQ.table,
  CC.ALL = DEG_CC.FLQMLQ.table,
  CC.UP = DEG.UP_CC.FLQMLQ.table, 
  CC.DOWN = DEG.DOWN_CC.FLQMLQ.table
)
  
#get a list of the godata objects
GOdata_list.FLQMLQ <- list(
  BP.ALL = DEgenes.FLQMLQ.GO.bp,
  BP.UP = DEgenes.FLQMLQ_UP.GO.bp, 
  BP.DOWN = DEgenes.FLQMLQ_DOWN.GO.bp,
  MF.ALL = DEgenes.FLQMLQ.GO.mf,
  MF.UP = DEgenes.FLQMLQ_UP.GO.mf, 
  MF.DOWN = DEgenes.FLQMLQ_DOWN.GO.mf,
  CC.ALL = DEgenes.FLQMLQ.GO.cc,
  CC.UP = DEgenes.FLQMLQ_UP.GO.cc, 
  CC.DOWN = DEgenes.FLQMLQ_DOWN.GO.cc
)

#geneID Extraction

# Step 1: Filter significant GO terms (p < 0.01), handling scientific notation
sig.terms_list.FLQMLQ <- lapply(results.FLQMLQ, function(tbl) {
    # Ensure p-values are numeric (handles cases where they are character strings like "1.2e-5")
    pvals <- as.numeric(tbl$raw.p.value)
    
    # Return GO.IDs with p < 0.01
    tbl$GO.ID[which(pvals < 0.01)]
})

# Step 2: Extract significant genes per GO term from each GOdata object
significant_genes_per_GO.FLQMLQ <- mapply(function(GOdata, go_ids) {
    # Get genes associated with each GO term
    genes_per_GO <- genesInTerm(object = GOdata, whichGO = go_ids)
    # Get significant genes from the GOdata object
    sigGenesSet <- sigGenes(GOdata)
    # Intersect genes in each GO term with the significant genes
    lapply(genes_per_GO, function(genes) intersect(genes, sigGenesSet))
  },
GOdata_list.FLQMLQ,
sig.terms_list.FLQMLQ,
SIMPLIFY = FALSE)



#plus terms
# Make a named vector of GO terms (Term names), per group
go_term_names_list.FLQMLQ <- lapply(results.FLQMLQ, function(tbl) {
  setNames(tbl$Term, tbl$GO.ID)
})



####This is the function that will be used across PW comparisons
#write_sig_genes_per_GO_to_excel()


write_sig_genes_per_GO_to_excel(
  sig_gene_list = significant_genes_per_GO.FLQMLQ,
  go_term_name_list = go_term_names_list.FLQMLQ,
  output_file = "Results/GenesPerGO/FLQMLQ_significant_genes_per_GO.xlsx"
)



#Verify the overlap of genes between all, up, down
#BP
genesFLQMLQ_BP_ALL <- unique(unlist(significant_genes_per_GO.FLQMLQ$BP.ALL))
genesFLQMLQ_BP_UP <- unique(unlist(significant_genes_per_GO.FLQMLQ$BP.UP))
genesFLQMLQ_BP_DOWN <- unique(unlist(significant_genes_per_GO.FLQMLQ$BP.DOWN))

  #OVERLAP
  intersect(genesFLQMLQ_BP_ALL, genesFLQMLQ_BP_UP)
  table(genesFLQMLQ_BP_UP %in% genesFLQMLQ_BP_ALL)
  table(genesFLQMLQ_BP_UP %in% names(DEG.genelist.FLQMLQ))
  
  intersect(genesFLQMLQ_BP_ALL, genesFLQMLQ_BP_DOWN)
  table(genesFLQMLQ_BP_DOWN %in% genesFLQMLQ_BP_ALL)
  table(genesFLQMLQ_BP_DOWN %in% names(DEG.genelist.FLQMLQ))



#MF
genesFLQMLQ_MF_ALL <- unique(unlist(significant_genes_per_GO.FLQMLQ$MF.ALL))
genesFLQMLQ_MF_UP <- unique(unlist(significant_genes_per_GO.FLQMLQ$MF.UP))
genesFLQMLQ_MF_DOWN <- unique(unlist(significant_genes_per_GO.FLQMLQ$MF.DOWN))

  #OVERLAP
  intersect(genesFLQMLQ_MF_ALL, genesFLQMLQ_MF_UP)
  table(genesFLQMLQ_MF_UP %in% genesFLQMLQ_MF_ALL)
  table(genesFLQMLQ_MF_UP %in% names(DEG.genelist.FLQMLQ))
  
  intersect(genesFLQMLQ_MF_ALL, genesFLQMLQ_MF_DOWN)
  table(genesFLQMLQ_MF_DOWN %in% genesFLQMLQ_MF_ALL)
  table(genesFLQMLQ_MF_DOWN %in% names(DEG.genelist.FLQMLQ))



#CC
genesFLQMLQ_CC_ALL <- unique(unlist(significant_genes_per_GO.FLQMLQ$CC.ALL))
genesFLQMLQ_CC_UP <- unique(unlist(significant_genes_per_GO.FLQMLQ$CC.UP))
genesFLQMLQ_CC_DOWN <- unique(unlist(significant_genes_per_GO.FLQMLQ$CC.DOWN))


  #OVERLAP
  intersect(genesFLQMLQ_CC_ALL, genesFLQMLQ_CC_UP)
  table(genesFLQMLQ_CC_UP %in% genesFLQMLQ_CC_ALL)
  table(genesFLQMLQ_CC_UP %in% names(DEG.genelist.FLQMLQ))
  
  intersect(genesFLQMLQ_CC_ALL, genesFLQMLQ_CC_DOWN)
  table(genesFLQMLQ_CC_DOWN %in% genesFLQMLQ_CC_ALL)
  table(genesFLQMLQ_CC_DOWN %in% names(DEG.genelist.FLQMLQ))


```






### FLQQLQ
#### Upregulated

#### BP
```{r BP_DEG_FLQQLQ_up}
GeneList_FLQQLQ_up <- as.integer(geneUniverse %in% DEgenes_UP.FLQQLQ)
names(GeneList_FLQQLQ_up) <- geneUniverse
LFCgenes <- function(x) x == 1

#verifying number of upreg genes with LFC < 1.5
table(GeneList_FLQQLQ_up)
#verifying the function
table(LFCgenes(GeneList_FLQQLQ_up))


DEgenes.FLQQLQ_UP.GO.bp <- new("topGOdata",
                description = paste("DEGUpGo", "BP", "FLQQLQ", sep = "_"),
                ontology = "BP",
                # all upregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLQQLQ_up, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Run enrichment test using Fisher's exact test
DEG.FLQQLQ_UP.bp_result_Fisher <- runTest(DEgenes.FLQQLQ_UP.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.FLQQLQ_UP.bp_result_Fisher

DEG.UP_BP.FLQQLQ.table <- GenTable(DEgenes.FLQQLQ_UP.GO.bp, 
                                   raw.p.value = DEG.FLQQLQ_UP.bp_result_Fisher, 
                                   topNodes = 50, 
                                   numChar = 120)

write.csv(DEG.UP_BP.FLQQLQ.table, "Results/UpDownSplit/UP/byontology/FLQQLQ_UP_BPTerms.csv")
```




#### MF
```{r mf_DEG_FLQQLQ_up}
DEgenes.FLQQLQ_UP.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "FLQQLQ", sep = "_"),
                ontology = "MF",
                # all upregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLQQLQ_up, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLQQLQ_UP.mf_result_fisher <- runTest(DEgenes.FLQQLQ_UP.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.FLQQLQ_UP.mf_result_fisher

DEG.UP_MF.FLQQLQ.table <- GenTable(DEgenes.FLQQLQ_UP.GO.mf, 
                                   raw.p.value = DEG.FLQQLQ_UP.mf_result_fisher, 
                                   topNodes = 50, 
                                   numChar = 120)

write.csv(DEG.UP_MF.FLQQLQ.table, "Results/UpDownSplit/UP/byontology/FLQQLQ_UP_MFTerms.csv")
```



#### CC
```{r cc_DEG_FLQQLQ_up}
DEgenes.FLQQLQ_UP.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "FLQQLQ", sep = "_"),
                ontology = "CC",
                # all upregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLQQLQ_up, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLQQLQ_UP.cc_result_fisher <- runTest(DEgenes.FLQQLQ_UP.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.FLQQLQ_UP.cc_result_fisher

DEG.UP_CC.FLQQLQ.table <- GenTable(DEgenes.FLQQLQ_UP.GO.cc, 
                                   raw.p.value = DEG.FLQQLQ_UP.cc_result_fisher, 
                                   topNodes = 50, 
                                   numChar = 120)

write.csv(DEG.UP_CC.FLQQLQ.table, "Results/UpDownSplit/UP/byontology/FLQQLQ_UP_CCTerms.csv")
```

```{r FLQQLQlargetable.UP}
#ALL
DEG.UP_BP.FLQQLQ.table <- cbind(DEG.UP_BP.FLQQLQ.table, ontology=rep("BP", length(DEG.UP_BP.FLQQLQ.table$GO.ID)))
DEG.UP_MF.FLQQLQ.table <- cbind(DEG.UP_MF.FLQQLQ.table, ontology=rep("MF", length(DEG.UP_MF.FLQQLQ.table$GO.ID)))
DEG.UP_CC.FLQQLQ.table <- cbind(DEG.UP_CC.FLQQLQ.table, ontology=rep("CC", length(DEG.UP_CC.FLQQLQ.table$GO.ID)))

DEG_GOallontology.FLQQLQ.UPregulated <-rbind(DEG.UP_BP.FLQQLQ.table, DEG.UP_MF.FLQQLQ.table, DEG.UP_CC.FLQQLQ.table)
write.csv(DEG_GOallontology.FLQQLQ.UPregulated, "Results/UpDownSplit/UP/ALL/DEGs_GOallontology_FLQQLQ_upregulated.csv")


#Less than .01
DEG.UP_BP.FLQQLQ.table.01 <- DEG.UP_BP.FLQQLQ.table[DEG.UP_BP.FLQQLQ.table$raw.p.value < 0.01,]
DEG.UP_MF.FLQQLQ.table.01 <- DEG.UP_MF.FLQQLQ.table[DEG.UP_MF.FLQQLQ.table$raw.p.value < 0.01,]
DEG.UP_CC.FLQQLQ.table.01 <- DEG.UP_CC.FLQQLQ.table[DEG.UP_CC.FLQQLQ.table$raw.p.value < 0.01,]

DEG_GOallontology.FLQQLQ.UPregulated.01 <-rbind(DEG.UP_BP.FLQQLQ.table.01, DEG.UP_MF.FLQQLQ.table.01, DEG.UP_CC.FLQQLQ.table.01)
write.csv(DEG_GOallontology.FLQQLQ.UPregulated.01, "Results/UpDownSplit/point1percent/DEGs_GOallontology_FLQQLQ_upregulated.01.csv")
```


#### Downregulated
#### BP
```{r BP_DEG_FLQQLQ_down}
GeneList_FLQQLQ_down <- as.integer(geneUniverse %in% DEgenes_DOWN.FLQQLQ)
names(GeneList_FLQQLQ_down) <- geneUniverse

#verifying number of upreg genes with LFC < -1.5
table(GeneList_FLQQLQ_down)
#verifying the function
table(LFCgenes(GeneList_FLQQLQ_down))

DEgenes.FLQQLQ_DOWN.GO.bp <- new("topGOdata",
                description = paste("DEGDownGo", "BP", "FLQQLQ", sep = "_"),
                ontology = "BP",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLQQLQ_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLQQLQ_DOWN.bp_result_fisher <- runTest(DEgenes.FLQQLQ_DOWN.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.FLQQLQ_DOWN.bp_result_fisher

DEG.DOWN_BP.FLQQLQ.table <- GenTable(DEgenes.FLQQLQ_DOWN.GO.bp, 
                                     raw.p.value = DEG.FLQQLQ_DOWN.bp_result_fisher, 
                                     topNodes = 50,
                                     numChar = 120)

write.csv(DEG.DOWN_BP.FLQQLQ.table, "Results/UpDownSplit/DOWN/byontology/FLQQLQ_DOWN_BPTerms.csv")
```


#### MF
```{r mf_DEG_FLQQLQ_down}
DEgenes.FLQQLQ_DOWN.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "FLQQLQ", sep = "_"),
                ontology = "MF",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLQQLQ_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLQQLQ_DOWN.mf_result_fisher <- runTest(DEgenes.FLQQLQ_DOWN.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.FLQQLQ_DOWN.mf_result_fisher

DEG.DOWN_MF.FLQQLQ.table <- GenTable(DEgenes.FLQQLQ_DOWN.GO.mf, 
                                     raw.p.value = DEG.FLQQLQ_DOWN.mf_result_fisher, 
                                     topNodes = 50, 
                                     numChar = 120)

write.csv(DEG.DOWN_MF.FLQQLQ.table, "Results/UpDownSplit/DOWN/byontology/FLQQLQ_DOWN_MFTerms.csv")
```



#### CC
```{r cc_DEG_FLQQLQ_down}
DEgenes.FLQQLQ_DOWN.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "FLQQLQ", sep = "_"),
                ontology = "CC",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLQQLQ_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLQQLQ_DOWN.cc_result_fisher <- runTest(DEgenes.FLQQLQ_DOWN.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.FLQQLQ_DOWN.cc_result_fisher

DEG.DOWN_CC.FLQQLQ.table <- GenTable(DEgenes.FLQQLQ_DOWN.GO.cc, 
                                     raw.p.value = DEG.FLQQLQ_DOWN.cc_result_fisher, 
                                     topNodes = 50,
                                     numChar = 120)

write.csv(DEG.DOWN_CC.FLQQLQ.table, "Results/UpDownSplit/DOWN/byontology/FLQQLQ_DOWN_CCTerms.csv")
```

```{r FLQQLQlargetable.DOWN}
#ALL
DEG.DOWN_BP.FLQQLQ.table <- cbind(DEG.DOWN_BP.FLQQLQ.table, ontology=rep("BP", length(DEG.DOWN_BP.FLQQLQ.table$GO.ID)))
DEG.DOWN_MF.FLQQLQ.table <- cbind(DEG.DOWN_MF.FLQQLQ.table, ontology=rep("MF", length(DEG.DOWN_MF.FLQQLQ.table$GO.ID)))
DEG.DOWN_CC.FLQQLQ.table <- cbind(DEG.DOWN_CC.FLQQLQ.table, ontology=rep("CC", length(DEG.DOWN_CC.FLQQLQ.table$GO.ID)))

DEG_GOallontology.FLQQLQ.DOWNregulated <-rbind(DEG.DOWN_BP.FLQQLQ.table, DEG.DOWN_MF.FLQQLQ.table, DEG.DOWN_CC.FLQQLQ.table)
write.csv(DEG_GOallontology.FLQQLQ.DOWNregulated, "Results/UpDownSplit/DOWN/ALL/DEGs_GOallontology_FLQQLQ_downregulated.csv")


#Less than .01
DEG.DOWN_BP.FLQQLQ.table.01 <- DEG.DOWN_BP.FLQQLQ.table[DEG.DOWN_BP.FLQQLQ.table$raw.p.value < 0.01,]
DEG.DOWN_MF.FLQQLQ.table.01 <- DEG.DOWN_MF.FLQQLQ.table[DEG.DOWN_MF.FLQQLQ.table$raw.p.value < 0.01,]
DEG.DOWN_CC.FLQQLQ.table.01 <- DEG.DOWN_CC.FLQQLQ.table[DEG.DOWN_CC.FLQQLQ.table$raw.p.value < 0.01,]

DEG_GOallontology.FLQQLQ.DOWNregulated.01 <-rbind(DEG.DOWN_BP.FLQQLQ.table.01, DEG.DOWN_MF.FLQQLQ.table.01, DEG.DOWN_CC.FLQQLQ.table.01)
write.csv(DEG_GOallontology.FLQQLQ.DOWNregulated.01, "Results/UpDownSplit/point1percent/DEGs_GOallontology_FLQQLQ_downregulated.01.csv")
```



```{r extractgenesforGOs.FLQQLQ}
#get results table
results.FLQQLQ <- list(BP.ALL = DEG_BP.FLQQLQ.table,
  BP.UP = DEG.UP_BP.FLQQLQ.table, 
  BP.DOWN = DEG.DOWN_BP.FLQQLQ.table,
  MF.ALL = DEG_MF.FLQQLQ.table,
  MF.UP = DEG.UP_MF.FLQQLQ.table, 
  MF.DOWN = DEG.DOWN_MF.FLQQLQ.table,
  CC.ALL = DEG_CC.FLQQLQ.table,
  CC.UP = DEG.UP_CC.FLQQLQ.table, 
  CC.DOWN = DEG.DOWN_CC.FLQQLQ.table
)
  
#get a list of the godata objects
GOdata_list.FLQQLQ <- list(
  BP.ALL = DEgenes.FLQQLQ.GO.bp,
  BP.UP = DEgenes.FLQQLQ_UP.GO.bp, 
  BP.DOWN = DEgenes.FLQQLQ_DOWN.GO.bp,
  MF.ALL = DEgenes.FLQQLQ.GO.mf,
  MF.UP = DEgenes.FLQQLQ_UP.GO.mf, 
  MF.DOWN = DEgenes.FLQQLQ_DOWN.GO.mf,
  CC.ALL = DEgenes.FLQQLQ.GO.cc,
  CC.UP = DEgenes.FLQQLQ_UP.GO.cc, 
  CC.DOWN = DEgenes.FLQQLQ_DOWN.GO.cc
)

#geneID Extraction

# Step 1: Filter significant GO terms (p < 0.01), handling scientific notation
sig.terms_list.FLQQLQ <- lapply(results.FLQQLQ, function(tbl) {
    # Ensure p-values are numeric (handles cases where they are character strings like "1.2e-5")
    pvals <- as.numeric(tbl$raw.p.value)
    
    # Return GO.IDs with p < 0.01
    tbl$GO.ID[which(pvals < 0.01)]
})

# Step 2: Extract significant genes per GO term from each GOdata object
significant_genes_per_GO.FLQQLQ <- mapply(function(GOdata, go_ids) {
    # Get genes associated with each GO term
    genes_per_GO <- genesInTerm(object = GOdata, whichGO = go_ids)
    # Get significant genes from the GOdata object
    sigGenesSet <- sigGenes(GOdata)
    # Intersect genes in each GO term with the significant genes
    lapply(genes_per_GO, function(genes) intersect(genes, sigGenesSet))
  },
GOdata_list.FLQQLQ,
sig.terms_list.FLQQLQ,
SIMPLIFY = FALSE)



#plus terms
# Make a named vector of GO terms (Term names), per group
go_term_names_list.FLQQLQ <- lapply(results.FLQQLQ, function(tbl) {
  setNames(tbl$Term, tbl$GO.ID)
})



####This is the function that will be used across PW comparisons
#write_sig_genes_per_GO_to_excel()


write_sig_genes_per_GO_to_excel(
  sig_gene_list = significant_genes_per_GO.FLQQLQ,
  go_term_name_list = go_term_names_list.FLQQLQ,
  output_file = "Results/GenesPerGO/FLQQLQ_significant_genes_per_GO.xlsx"
)



#Verify the overlap of genes between all, up, down
#BP
genesFLQQLQ_BP_ALL <- unique(unlist(significant_genes_per_GO.FLQQLQ$BP.ALL))
genesFLQQLQ_BP_UP <- unique(unlist(significant_genes_per_GO.FLQQLQ$BP.UP))
genesFLQQLQ_BP_DOWN <- unique(unlist(significant_genes_per_GO.FLQQLQ$BP.DOWN))

  #OVERLAP
  intersect(genesFLQQLQ_BP_ALL, genesFLQQLQ_BP_UP)
  table(genesFLQQLQ_BP_UP %in% genesFLQQLQ_BP_ALL)
  table(genesFLQQLQ_BP_UP %in% names(DEG.genelist.FLQQLQ))
  
  intersect(genesFLQQLQ_BP_ALL, genesFLQQLQ_BP_DOWN)
  table(genesFLQQLQ_BP_DOWN %in% genesFLQQLQ_BP_ALL)
  table(genesFLQQLQ_BP_DOWN %in% names(DEG.genelist.FLQQLQ))



#MF
genesFLQQLQ_MF_ALL <- unique(unlist(significant_genes_per_GO.FLQQLQ$MF.ALL))
genesFLQQLQ_MF_UP <- unique(unlist(significant_genes_per_GO.FLQQLQ$MF.UP))
genesFLQQLQ_MF_DOWN <- unique(unlist(significant_genes_per_GO.FLQQLQ$MF.DOWN))

  #OVERLAP
  intersect(genesFLQQLQ_MF_ALL, genesFLQQLQ_MF_UP)
  table(genesFLQQLQ_MF_UP %in% genesFLQQLQ_MF_ALL)
  table(genesFLQQLQ_MF_UP %in% names(DEG.genelist.FLQQLQ))
  
  intersect(genesFLQQLQ_MF_ALL, genesFLQQLQ_MF_DOWN)
  table(genesFLQQLQ_MF_DOWN %in% genesFLQQLQ_MF_ALL)
  table(genesFLQQLQ_MF_DOWN %in% names(DEG.genelist.FLQQLQ))



#CC
genesFLQQLQ_CC_ALL <- unique(unlist(significant_genes_per_GO.FLQQLQ$CC.ALL))
genesFLQQLQ_CC_UP <- unique(unlist(significant_genes_per_GO.FLQQLQ$CC.UP))
genesFLQQLQ_CC_DOWN <- unique(unlist(significant_genes_per_GO.FLQQLQ$CC.DOWN))


  #OVERLAP
  intersect(genesFLQQLQ_CC_ALL, genesFLQQLQ_CC_UP)
  table(genesFLQQLQ_CC_UP %in% genesFLQQLQ_CC_ALL)
  table(genesFLQQLQ_CC_UP %in% names(DEG.genelist.FLQQLQ))
  
  intersect(genesFLQQLQ_CC_ALL, genesFLQQLQ_CC_DOWN)
  table(genesFLQQLQ_CC_DOWN %in% genesFLQQLQ_CC_ALL)
  table(genesFLQQLQ_CC_DOWN %in% names(DEG.genelist.FLQQLQ))


```








### FLWMLW
#### Upregulated
No upregulated genes.



#### Downregulated
#### BP
```{r BP_DEG_FLWMLW_down}
GeneList_FLWMLW_down <- as.integer(geneUniverse %in% DEgenes_DOWN.FLWMLW)
names(GeneList_FLWMLW_down) <- geneUniverse

#verifying number of upreg genes with LFC < -1.5
table(GeneList_FLWMLW_down)
#verifying the function
table(LFCgenes(GeneList_FLWMLW_down))

DEgenes.FLWMLW_DOWN.GO.bp <- new("topGOdata",
                description = paste("DEGDownGo", "BP", "FLWMLW", sep = "_"),
                ontology = "BP",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLWMLW_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLWMLW_DOWN.bp_result_fisher <- runTest(DEgenes.FLWMLW_DOWN.GO.bp, statistic = "fisher", algorithm = "weight01")
DEG.FLWMLW_DOWN.bp_result_fisher

DEG.DOWN_BP.FLWMLW.table <- GenTable(DEgenes.FLWMLW_DOWN.GO.bp, 
                                     raw.p.value = DEG.FLWMLW_DOWN.bp_result_fisher, 
                                     topNodes = 50,
                                     numChar = 120)

write.csv(DEG.DOWN_BP.FLWMLW.table, "Results/UpDownSplit/DOWN/byontology/FLWMLW_DOWN_BPTerms.csv")
```


#### MF
```{r mf_DEG_FLWMLW_down}
DEgenes.FLWMLW_DOWN.GO.mf <- new("topGOdata",
                description = paste("DEG Go", "MF", "FLWMLW", sep = "_"),
                ontology = "MF",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLWMLW_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLWMLW_DOWN.mf_result_fisher <- runTest(DEgenes.FLWMLW_DOWN.GO.mf, statistic = "fisher", algorithm = "weight01")
DEG.FLWMLW_DOWN.mf_result_fisher

DEG.DOWN_MF.FLWMLW.table <- GenTable(DEgenes.FLWMLW_DOWN.GO.mf, 
                                     raw.p.value = DEG.FLWMLW_DOWN.mf_result_fisher, 
                                     topNodes = 50, 
                                     numChar = 120)

write.csv(DEG.DOWN_MF.FLWMLW.table, "Results/UpDownSplit/DOWN/byontology/FLWMLW_DOWN_MFTerms.csv")
```



#### CC
```{r cc_DEG_FLWMLW_down}
DEgenes.FLWMLW_DOWN.GO.cc <- new("topGOdata",
                description = paste("DEG Go", "CC", "FLWMLW", sep = "_"),
                ontology = "CC",
                # all downregulated genes in this comparison, genes of interest = 1
                allGenes = GeneList_FLWMLW_down, 
                geneSelectionFun = LFCgenes,
                gene2GO = gene2mapping,
                annot = annFUN.gene2GO)


#Fisher Testing
DEG.FLWMLW_DOWN.cc_result_fisher <- runTest(DEgenes.FLWMLW_DOWN.GO.cc, statistic = "fisher", algorithm = "weight01")
DEG.FLWMLW_DOWN.cc_result_fisher

DEG.DOWN_CC.FLWMLW.table <- GenTable(DEgenes.FLWMLW_DOWN.GO.cc, 
                                     raw.p.value = DEG.FLWMLW_DOWN.cc_result_fisher, 
                                     topNodes = 50,
                                     numChar = 120)

write.csv(DEG.DOWN_CC.FLWMLW.table, "Results/UpDownSplit/DOWN/byontology/FLWMLW_DOWN_CCTerms.csv")
```

```{r FLWMLWlargetable.DOWN}
#ALL
DEG.DOWN_BP.FLWMLW.table <- cbind(DEG.DOWN_BP.FLWMLW.table, ontology=rep("BP", length(DEG.DOWN_BP.FLWMLW.table$GO.ID)))
DEG.DOWN_MF.FLWMLW.table <- cbind(DEG.DOWN_MF.FLWMLW.table, ontology=rep("MF", length(DEG.DOWN_MF.FLWMLW.table$GO.ID)))
DEG.DOWN_CC.FLWMLW.table <- cbind(DEG.DOWN_CC.FLWMLW.table, ontology=rep("CC", length(DEG.DOWN_CC.FLWMLW.table$GO.ID)))

DEG_GOallontology.FLWMLW.DOWNregulated <-rbind(DEG.DOWN_BP.FLWMLW.table, DEG.DOWN_MF.FLWMLW.table, DEG.DOWN_CC.FLWMLW.table)
write.csv(DEG_GOallontology.FLWMLW.DOWNregulated, "Results/UpDownSplit/DOWN/ALL/DEGs_GOallontology_FLWMLW_downregulated.csv")


#Less than .01
DEG.DOWN_BP.FLWMLW.table.01 <- DEG.DOWN_BP.FLWMLW.table[DEG.DOWN_BP.FLWMLW.table$raw.p.value < 0.01,]
DEG.DOWN_MF.FLWMLW.table.01 <- DEG.DOWN_MF.FLWMLW.table[DEG.DOWN_MF.FLWMLW.table$raw.p.value < 0.01,]
DEG.DOWN_CC.FLWMLW.table.01 <- DEG.DOWN_CC.FLWMLW.table[DEG.DOWN_CC.FLWMLW.table$raw.p.value < 0.01,]

DEG_GOallontology.FLWMLW.DOWNregulated.01 <-rbind(DEG.DOWN_BP.FLWMLW.table.01, DEG.DOWN_MF.FLWMLW.table.01, DEG.DOWN_CC.FLWMLW.table.01)
write.csv(DEG_GOallontology.FLWMLW.DOWNregulated.01, "Results/UpDownSplit/point1percent/DEGs_GOallontology_FLWMLW_downregulated.01.csv")
```



```{r extractgenesforGOs.FLWMLW}
#get results table
results.FLWMLW <- list(BP.ALL = DEG_BP.FLWMLW.table,
  BP.DOWN = DEG.DOWN_BP.FLWMLW.table,
  MF.ALL = DEG_MF.FLWMLW.table,
  MF.DOWN = DEG.DOWN_MF.FLWMLW.table,
  CC.ALL = DEG_CC.FLWMLW.table,
  CC.DOWN = DEG.DOWN_CC.FLWMLW.table
)
  
#get a list of the godata objects
GOdata_list.FLWMLW <- list(
  BP.ALL = DEgenes.FLWMLW.GO.bp,
  BP.DOWN = DEgenes.FLWMLW_DOWN.GO.bp,
  MF.ALL = DEgenes.FLWMLW.GO.mf,
  MF.DOWN = DEgenes.FLWMLW_DOWN.GO.mf,
  CC.ALL = DEgenes.FLWMLW.GO.cc,
  CC.DOWN = DEgenes.FLWMLW_DOWN.GO.cc
)

#geneID Extraction

# Step 1: Filter significant GO terms (p < 0.01), handling scientific notation
sig.terms_list.FLWMLW <- lapply(results.FLWMLW, function(tbl) {
    # Ensure p-values are numeric (handles cases where they are character strings like "1.2e-5")
    pvals <- as.numeric(tbl$raw.p.value)
    
    # Return GO.IDs with p < 0.01
    tbl$GO.ID[which(pvals < 0.01)]
})

# Step 2: Extract significant genes per GO term from each GOdata object
significant_genes_per_GO.FLWMLW <- mapply(function(GOdata, go_ids) {
    # Get genes associated with each GO term
    genes_per_GO <- genesInTerm(object = GOdata, whichGO = go_ids)
    # Get significant genes from the GOdata object
    sigGenesSet <- sigGenes(GOdata)
    # Intersect genes in each GO term with the significant genes
    lapply(genes_per_GO, function(genes) intersect(genes, sigGenesSet))
  },
GOdata_list.FLWMLW,
sig.terms_list.FLWMLW,
SIMPLIFY = FALSE)



#plus terms
# Make a named vector of GO terms (Term names), per group
go_term_names_list.FLWMLW <- lapply(results.FLWMLW, function(tbl) {
  setNames(tbl$Term, tbl$GO.ID)
})



####This is the function that will be used across PW comparisons
#write_sig_genes_per_GO_to_excel()


write_sig_genes_per_GO_to_excel(
  sig_gene_list = significant_genes_per_GO.FLWMLW,
  go_term_name_list = go_term_names_list.FLWMLW,
  output_file = "Results/GenesPerGO/FLWMLW_significant_genes_per_GO.xlsx"
)



#Verify the overlap of genes between all, up, down
#BP
genesFLWMLW_BP_ALL <- unique(unlist(significant_genes_per_GO.FLWMLW$BP.ALL))
genesFLWMLW_BP_DOWN <- unique(unlist(significant_genes_per_GO.FLWMLW$BP.DOWN))

  #OVERLAP
  intersect(genesFLWMLW_BP_ALL, genesFLWMLW_BP_DOWN)
  table(genesFLWMLW_BP_DOWN %in% genesFLWMLW_BP_ALL)
  table(genesFLWMLW_BP_DOWN %in% names(DEG.genelist.FLWMLW))



#MF
genesFLWMLW_MF_ALL <- unique(unlist(significant_genes_per_GO.FLWMLW$MF.ALL))
genesFLWMLW_MF_DOWN <- unique(unlist(significant_genes_per_GO.FLWMLW$MF.DOWN))

  #OVERLAP
  intersect(genesFLWMLW_MF_ALL, genesFLWMLW_MF_DOWN)
  table(genesFLWMLW_MF_DOWN %in% genesFLWMLW_MF_ALL)
  table(genesFLWMLW_MF_DOWN %in% names(DEG.genelist.FLWMLW))



#CC
genesFLWMLW_CC_ALL <- unique(unlist(significant_genes_per_GO.FLWMLW$CC.ALL))
genesFLWMLW_CC_DOWN <- unique(unlist(significant_genes_per_GO.FLWMLW$CC.DOWN))


  #OVERLAP
  intersect(genesFLWMLW_CC_ALL, genesFLWMLW_CC_DOWN)
  table(genesFLWMLW_CC_DOWN %in% genesFLWMLW_CC_ALL)
  table(genesFLWMLW_CC_DOWN %in% names(DEG.genelist.FLWMLW))


```


## Get Gene Lists for GO IDs of interest
```{r saveforgenework}
save(significant_genes_per_GO.FLQQLQ, significant_genes_per_GO.FLQMLQ, significant_genes_per_GO.FLWFLQ, significant_genes_per_GO.FLWMLW, significant_genes_per_GO.MLWMLQ, significant_genes_per_GO.QLWQLQ, file = "Results/GenesPerGO/SignificantGenesPerGO.Rdata")

```


### Create Vector of GO IDs of interest
```{r GOIDs}
#Use when not knitting whole docu
load("Results/GenesPerGO/SignificantGenesPerGO.Rdata")

GOvector <- c("GO:0007631", "GO:0019563", "GO:0004371", "GO:0004970", "GO:0022008", "GO:0045202", "GO:0004890", "GO:0007218", "GO:0062129", "GO:0035725", "GO:0019367", "GO:0034625", "GO:0034626", "GO:0042761", "GO:0006680", "GO:0006629", "GO:0005615", "GO:0034464", "GO:0004181", "GO:0071683", "GO:0007601", "GO:0007602","GO:0050911", "GO:0019230", "GO:0050954", "GO:0004984", "GO:0016499", "GO:0015179", "GO:0005576", "GO:0004707"

)



GeneObj <- list(
  FLQQLQ=significant_genes_per_GO.FLQQLQ,
  FLQMLQ=significant_genes_per_GO.FLQMLQ,
  FLWFLQ=significant_genes_per_GO.FLWFLQ,
  FLWMLW=significant_genes_per_GO.FLWMLW,
  MLWMLQ=significant_genes_per_GO.MLWMLQ,
  QLWQLQ=significant_genes_per_GO.QLWQLQ
)
```


### Function to Isolate Genes for GOs of interest
```{r GOIDstoGENES, include=FALSE}
extract_genes_by_go <- function(go_terms, objects) {
  if (!is.list(objects)) stop("`objects` must be a list of GO objects")
  
  result <- lapply(go_terms, function(go) {
    # collect genes from all objects and all slots
    all_genes <- unlist(lapply(names(objects), function(obj_name) {
      obj <- objects[[obj_name]]
      
      # loop through all slots in this object
      genes_in_obj <- unlist(lapply(names(obj), function(slot_name) {
        slot_obj <- obj[[slot_name]]
        if (go %in% names(slot_obj)) {
          return(slot_obj[[go]])
        } else {
          message("GO term ", go, " not found in ", obj_name, " (", slot_name, ")")
          return(character(0))
        }
      }))
      
      return(genes_in_obj)
    }))
    
    return(unique(all_genes))
  })
  
  # Clean names: remove ":" from GO IDs
  names(result) <- gsub(":", "", go_terms)
  return(result)
}


genes_byGO <- extract_genes_by_go(go_terms = GOvector, objects = GeneObj)
```


```{r loadincountdata}
#load in the same count matrix used for WGCNA
rawCounts0 <- read.csv2("../../../RNASeqAnalysis/BimpMB_RNAanalysis/Filtering/GECountData_GenesOutliers_Filtered.csv", sep=";")
head(rawCounts0)
dim(rawCounts0)


#Read in with samples as column names and gene as column 1



#load in the sample mappings
sampleData0 <- read.delim("../../../RNASeqAnalysis/BimpMB_RNAanalysis/Filtering/sampleData_OutliersRemoved.csv", sep=";")
head(sampleData0)

# Also save a copy for later
#sampleData <- sampleData0
```

```{r dosomereformating}

rawCounts <- rawCounts0
rownames(rawCounts) <- rawCounts$X
rawCounts <- rawCounts[,-1]

sampleData <- sampleData0
rownames(sampleData) <- sampleData$sample_id
sampleData <- sampleData[,1:2]


counts_pm <- cpm(rawCounts, log=TRUE)
# Make counts_pm a data frame and keep gene names
counts_df <- counts_pm %>%
  as.data.frame() %>%
  rownames_to_column(var = "gene")

```


## Compare Counts (pm) of genes of interest
### GO:0007631 - FEEDING BEHAVIOR
```{r FBgenes}
FBGene <- genes_byGO$GO0007631

FBGeneCounts_lcpm <- counts_df %>%
  filter(gene %in% FBGene)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
FBGeneCounts_lcpm_long <- FBGeneCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
FBGeneCounts_lcpm_long <- FBGeneCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

FBGeneCounts_lcpm_long$tx_group <- factor(FBGeneCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
ggplot(FBGeneCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) + #, label = sample_id)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  #geom_text(position = position_jitter(seed = 1)) +
  #facet_wrap(~Gene, scales = "free_y") + # One plot per gene
  theme_minimal() +
    labs(
    title = "Gene: LOC100742013 (GO:0007631-FEEDING BEHAVIOR)",
    x = "Treatment Group",
    y = "Log-CPM Expression"
    ) +
theme(axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(hjust = 0.5))

```



### GO:00016499 - OREXIN RECEPTOR ACTIVITY (MF)
```{r ORAGenes, fig.height=6}
ORAGenes <- genes_byGO$GO0016499

ORAGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% ORAGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
ORAGenesCounts_lcpm_long <- ORAGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
ORAGenesCounts_lcpm_long <- ORAGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

ORAGenesCounts_lcpm_long$tx_group <- factor(ORAGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(ORAGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y", ncol=2) +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:00016499 - orexin receptor activity",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```



### GO:0007218 - NEUROPEPTIDE SIGNALLING
```{r NPGenes}
NPGene <- genes_byGO$GO0007218

NPGeneCounts_lcpm <- counts_df %>%
  filter(gene %in% NPGene)

#as.data.frame
#as.data.frame(NPGeneCounts_lcpm)
#colnames(NPGeneCounts_lcpm) <- NPGene[1]


# Pivot longer so each row is gene × sample
NPGeneCounts_lcpm_long <- NPGeneCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
NPGeneCounts_lcpm_long <- NPGeneCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

NPGeneCounts_lcpm_long$tx_group <- factor(NPGeneCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
ggplot(NPGeneCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) + #, label = sample_id)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  #geom_text(position = position_jitter(seed = 1)) +
  #facet_wrap(~Gene, scales = "free_y") + # One plot per gene
  theme_minimal() +
    labs(
    title = "Gene: LOC100745005 (GO:0007218 - Neuropeptide Signalling)",
    x = "Treatment Group",
    y = "Log-CPM Expression"
    ) +
theme(axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(hjust = 0.5))

```



### GO:0022008 - NEUROGENESIS
```{r NGGenes, fig.height=9}
NGGenes <- genes_byGO$GO0022008

NGGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% NGGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
NGGenesCounts_lcpm_long <- NGGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
NGGenesCounts_lcpm_long <- NGGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

NGGenesCounts_lcpm_long$tx_group <- factor(NGGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
ggplot(NGGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y", ncol=2) +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0022008 - Neurogenesis (BP)",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )
```





### GO:0045202 - SYNAPSE
```{r SynGenes, fig.height=9}
SynGenes <- genes_byGO$GO0045202

SynGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% SynGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
SynGenesCounts_lcpm_long <- SynGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
SynGenesCounts_lcpm_long <- SynGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

SynGenesCounts_lcpm_long$tx_group <- factor(SynGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
ggplot(SynGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y", ncol=2) +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0045202 - Synapse (CC)",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )
```



### GO:0004707 - MAP KINASE ACTIVITY
```{r MAPGenes, fig.height=9}
MAPGenes <- genes_byGO$GO0004707

MAPGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% MAPGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
MAPGenesCounts_lcpm_long <- MAPGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
MAPGenesCounts_lcpm_long <- MAPGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

MAPGenesCounts_lcpm_long$tx_group <- factor(MAPGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
ggplot(MAPGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y", ncol=2) +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0004707 - MAP kinase activity (MF)",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )
```




### GO:0035725 - SODIUM ION TRANSMEMBRANE TRANSPORT
```{r SITTGenes, fig.height=8}
SITTGenes <- genes_byGO$GO0035725

SITTGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% SITTGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
SITTGenesCounts_lcpm_long <- SITTGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
SITTGenesCounts_lcpm_long <- SITTGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

SITTGenesCounts_lcpm_long$tx_group <- factor(SITTGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(SITTGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y", ncol=2) +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0035725 - sodium ion transmembrane transport",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```



### GO:0004970 - GLUTAMATE-GATED RECEPTOR ACTIVITY
```{r GGRAGenes, fig.height=9}
GGRAGenes <- genes_byGO$GO0004970

GGRAGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% GGRAGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
GGRAGenesCounts_lcpm_long <- GGRAGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
GGRAGenesCounts_lcpm_long <- GGRAGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

GGRAGenesCounts_lcpm_long$tx_group <- factor(GGRAGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(GGRAGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y", ncol=2) +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0004970 - Glutamate-gated receptor activity",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```




### GO:0004890 - GABA-A RECEPTOR ACTIVITY
```{r GARAGenes, fig.height=3}
GARAGenes <- genes_byGO$GO0004890

GARAGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% GARAGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
GARAGenesCounts_lcpm_long <- GARAGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
GARAGenesCounts_lcpm_long <- GARAGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

GARAGenesCounts_lcpm_long$tx_group <- factor(GARAGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(GARAGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y") +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0004890 - GABA-A receptor activity",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```



### GO:0034464 - BBSome
```{r BBSomeGenes, fig.height=6}
BBSomeGenes <- genes_byGO$GO0034464

BBSomeGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% BBSomeGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
BBSomeGenesCounts_lcpm_long <- BBSomeGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
BBSomeGenesCounts_lcpm_long <- BBSomeGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

BBSomeGenesCounts_lcpm_long$tx_group <- factor(BBSomeGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(BBSomeGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y", ncol=2) +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0034464 - BBSome",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```



### GO:0071683 - SENSORY DENDRITE
```{r SDGenes, fig.height=6}
SDGenes <- genes_byGO$GO0071683

SDGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% SDGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
SDGenesCounts_lcpm_long <- SDGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
SDGenesCounts_lcpm_long <- SDGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

SDGenesCounts_lcpm_long$tx_group <- factor(SDGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(SDGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y", ncol=2) +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0071683 - sensory dendrite",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```




### GO:0019230 - PROPRIOCEPTION
```{r PRGenes, fig.height=6}
PRGenes <- genes_byGO$GO0019230

PRGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% PRGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
PRGenesCounts_lcpm_long <- PRGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
PRGenesCounts_lcpm_long <- PRGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

PRGenesCounts_lcpm_long$tx_group <- factor(PRGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(PRGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y", ncol=2) +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0019230 - proprioception",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```





### GO:0050954 - Sensory Perception of mechanical stimuli
```{r PMSGenes, fig.height=6}
PMSGenes <- genes_byGO$GO0050954

PMSGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% PMSGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
PMSGenesCounts_lcpm_long <- PMSGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "exPMSession")

# Join metadata
PMSGenesCounts_lcpm_long <- PMSGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

PMSGenesCounts_lcpm_long$tx_group <- factor(PMSGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(PMSGenesCounts_lcpm_long, aes(x = tx_group, y = exPMSession, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y", ncol=2) +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0050954 - Sensory perception of mechanical stimuli",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```




### GO:0007601; GO:0007602 - VISUAL PERCEPTION AND PHOTOTRANSDUCTION
```{r VPPGenes, fig.height=6}
VPPGenes <- c(genes_byGO$GO0007601, genes_byGO$GO0007602)

VPPGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% VPPGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
VPPGenesCounts_lcpm_long <- VPPGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "exVPPession")

# Join metadata
VPPGenesCounts_lcpm_long <- VPPGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

VPPGenesCounts_lcpm_long$tx_group <- factor(VPPGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(VPPGenesCounts_lcpm_long, aes(x = tx_group, y = exVPPession, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y", ncol=2) +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0007601 - Visual perception (both genes) & GO:0007602 - phototransduction (LOC100746131)",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```




### GO:0050911; GO:0004984 - detection of chemical signal for smell perception & olfactory receptor activity
```{r SPOGenes, fig.height=6}
SPOGenes <- c(genes_byGO$GO0050911, genes_byGO$GO0004984)

SPOGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% SPOGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
SPOGenesCounts_lcpm_long <- SPOGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "exSPOession")

# Join metadata
SPOGenesCounts_lcpm_long <- SPOGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

SPOGenesCounts_lcpm_long$tx_group <- factor(SPOGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(SPOGenesCounts_lcpm_long, aes(x = tx_group, y = exSPOession, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y", ncol=2) +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0050911; GO:0004984 - detection of chemical signal for smell perception & olfactory receptor activity",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```




### GO:0004181 - METALLOCARBOXYPEPTIDASE ACTIVITY
```{r MCPGenes, fig.height=6}
MCPGenes <- genes_byGO$GO0004181

MCPGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% MCPGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
MCPGenesCounts_lcpm_long <- MCPGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
MCPGenesCounts_lcpm_long <- MCPGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

MCPGenesCounts_lcpm_long$tx_group <- factor(MCPGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(MCPGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y") +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0004181 - Metallocarboxypeptidase Activity",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```









### GO:0006629 - LIPID METABOLIC PROCESS
```{r LMPGenes, fig.height=9}
LMPGenes <- genes_byGO$GO0006629

LMPGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% LMPGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
LMPGenesCounts_lcpm_long <- LMPGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
LMPGenesCounts_lcpm_long <- LMPGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

LMPGenesCounts_lcpm_long$tx_group <- factor(LMPGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(LMPGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y", ncol=2) +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0006629 - lipid metabolic process",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```



### GO:0006680 - GLUCOSYLCERAMIDE CATABOLIC PROCESS
```{r GluCPGenes, fig.height=5}
GluCPGenes <- genes_byGO$GO0006680

GluCPGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% GluCPGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
GluCPGenesCounts_lcpm_long <- GluCPGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
GluCPGenesCounts_lcpm_long <- GluCPGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

GluCPGenesCounts_lcpm_long$tx_group <- factor(GluCPGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(GluCPGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y") +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0006680 - Glucosylceramide Catabolic Process",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```





### GO:0019563 - GLYCEROL CATABOLIC PROCESS
```{r GCPGenes, fig.height=6}
GCPGenes <- genes_byGO$GO0019563

GCPGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% GCPGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
GCPGenesCounts_lcpm_long <- GCPGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
GCPGenesCounts_lcpm_long <- GCPGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

GCPGenesCounts_lcpm_long$tx_group <- factor(GCPGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(GCPGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y") +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0019563 - Glycerol Catabolic Process",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```
 





### GO:0004371 - GLYCERONE KINASE ACTIVITY
```{r GKAGenes, fig.height=6}
GKAGenes <- genes_byGO$GO0004371

GKAGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% GKAGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
GKAGenesCounts_lcpm_long <- GKAGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
GKAGenesCounts_lcpm_long <- GKAGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

GKAGenesCounts_lcpm_long$tx_group <- factor(GKAGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(GKAGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y") +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0004371 - Glycerone Kinase Activity",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```






### GO:0019367, GO:0034625, GO:0034626, GO:0042761 - FATTY ACID LONGATION AND BIOSYNTHESIS (single gene: LOC100749525)
There is a single significant gene across these 4 similar go terms. So, considering them all here.
```{r FAEBGenes, fig.height=3}
FAEBGenes <- genes_byGO$GO0019367

FAEBGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% FAEBGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
FAEBGenesCounts_lcpm_long <- FAEBGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
FAEBGenesCounts_lcpm_long <- FAEBGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

FAEBGenesCounts_lcpm_long$tx_group <- factor(FAEBGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(FAEBGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y") +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0019367, GO:0034625, GO:0034626, GO:0042761 - fatty acid elongation and biosynthesis",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```








### GO:0062129 - CHITIN-BASED EXTRACELLULAR MATRIX
```{r CBEMGenes, fig.height=6}
CBEMGenes <- genes_byGO$GO0062129

CBEMGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% CBEMGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
CBEMGenesCounts_lcpm_long <- CBEMGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
CBEMGenesCounts_lcpm_long <- CBEMGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

CBEMGenesCounts_lcpm_long$tx_group <- factor(CBEMGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(CBEMGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y", ncol=2) +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0062129 - chitin-based extracellular matrix
",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```





### GO:0015179 - L-AMINO ACID TRANSMEMBRANE TRANSPORTER ACTIVITY
```{r LAATTAGenes, fig.height=6}
LAATTAGenes <- genes_byGO$GO0015179

LAATTAGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% LAATTAGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
LAATTAGenesCounts_lcpm_long <- LAATTAGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
LAATTAGenesCounts_lcpm_long <- LAATTAGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

LAATTAGenesCounts_lcpm_long$tx_group <- factor(LAATTAGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(LAATTAGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y", ncol=2) +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0015179 - L-amino acid transmembrane transporter activity",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```



### GO:0005576 - EXTRACELLULAR REGION (cc)
```{r ECRGenes, fig.height=6}
ECRGenes <- genes_byGO$GO0005576

ECRGenesCounts_lcpm <- counts_df %>%
  filter(gene %in% ECRGenes)

#as.data.frame
#as.data.frame(FBGeneCounts_lcpm)
#colnames(FBGeneCounts_lcpm) <- FBGene[1]


# Pivot longer so each row is gene × sample
ECRGenesCounts_lcpm_long <- ECRGenesCounts_lcpm %>%
  pivot_longer(-gene, names_to = "sample_id", values_to = "expression")

# Join metadata
ECRGenesCounts_lcpm_long <- ECRGenesCounts_lcpm_long %>%
  left_join(sampleData, by = c("sample_id" = "sample_id"))

ECRGenesCounts_lcpm_long$tx_group <- factor(ECRGenesCounts_lcpm_long$tx_group, levels= c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))


#Make PLOT
# Multipanel plot for all genes in this GO term
ggplot(ECRGenesCounts_lcpm_long, aes(x = tx_group, y = expression, fill = tx_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~gene, scales = "free_y", ncol=2) +  # one panel per gene
  theme_minimal() +
  labs(
    title = "GO:0005576 - extracellular region",
    x = "Treatment Group",
    y = "Log-CPM Expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10) # facet labels
  )

```









# Trying to locate the GO assigned to Gene LOC105680522 and LOC117151569

## Looking for Gene ID: LOC105680522
Gene LOC105680522 is the last of the 6 DEGs across all across caste TX groups. I can't determine its GO assignment using the sig go terms, so I am trying to ID across all GOs. Reworking here as not to interfere with analysis. 
```{r findingLOC105680522}

LOC105680522 <- "LOC105680522"

FLWFLQ_GOs <- c(DEG_GOallontology.FLWFLQ.UPregulated$GO.ID, DEG_GOallontology.FLWFLQ.DOWNregulated$GO.ID)

ID_LOC105680522 <- unique(FLWFLQ_GOs)


#GOs_FLWFLQ <- genesInTerm(object = DEgenes.FLWFLQ.GO.bp, whichGO = ID_LOC105680522)
#GOforLOC105680522 <- LOC105680522[LOC105680522 %in% unlist(GOs_FLWFLQ)]
#GOforLOC105680522
# Combine your GO objects into a named list
GOobjects <- list(
  BP.up = DEgenes.FLWFLQ_UP.GO.bp,
  MF.up = DEgenes.FLWFLQ_UP.GO.mf,
  CC.up = DEgenes.FLWFLQ_UP.GO.cc,
  BP.down = DEgenes.FLWFLQ_DOWN.GO.bp,
  MF.down = DEgenes.FLWFLQ_DOWN.GO.mf,
  CC.down = DEgenes.FLWFLQ_DOWN.GO.cc
)




# function: find all GO terms across BP, MF, CC
find_gene_in_all_GO <- function(gene, GOlist) {
  result <- list()
  
  for (ont in names(GOlist)) {
    go_terms <- GOlist[[ont]]
    
    # ensure the inner structure is a list (each GO term = vector of genes)
    if (!is.list(go_terms)) next
    
    # find which GO terms contain the gene
    matched_terms <- names(Filter(function(gvec) gene %in% gvec, go_terms))
    
    result[[ont]] <- matched_terms
  }
  
  result
}


GOmatches_LOC105680522 <- find_gene_in_all_GO(
  LOC105680522,
  significant_genes_per_GO.FLWFLQ
)

GOmatches_LOC105680522



```

Okay, we aren't finding the Gene among our go terms. There are a few possibilities: 1) the function doesn't work, 2) the gene isn't in the gene universe, 3) the gene is among the NAs removed, and 4) the gene hasn't been annotated. I will assess these in that order.

### 1) Test gene locating function
Testing the function with a gene we know is there!
```{r testingfun}

gene_known <- "LOC100742982"

GOmatches_LOC100742982 <- find_gene_in_all_GO(
  gene_known,
  significant_genes_per_GO.FLWFLQ
)

GOmatches_LOC100742982

```


### 2) See if the gene is in the universe
Checking to make sure it is in the gene universe.
```{r geneuniversecheck}

LOC105680522 %in% geneUniverse

```

The gene is present in the gene universe.

### 3) Was the gene removed as an outlier pre-Fisher test
Checking to make sure it is not one of the NAs removed.
```{r outliercheck}

LOC105680522 %in% FLWFLQ_napos

```

It is not a gene that was removed as an outlier.


### 4) Has the gene been annotated
Checking to make sure the gene has been annotated (i.e., is in your gene2Mapping annotation file).
```{r annotationcheck}

LOC105680522 %in% gene2mapping

```

The gene is not in the mapping file. Therefore, I am concluding that **LOC105680522** hasn't been annotated.



## Looking for Gene ID: LOC117151569
Gene LOC117151569 is the last of the 6 DEGs across all across caste TX groups. I can't determine its GO assignment using the sig go terms, so I am trying to ID across all GOs. Reworking here as not to interfere with analysis. 
```{r findingLOC117151569}

LOC117151569 <- "LOC117151569"

FLWFLQ_GOs <- c(DEG_GOallontology.FLWFLQ.UPregulated$GO.ID, DEG_GOallontology.FLWFLQ.DOWNregulated$GO.ID)

ID_LOC117151569 <- unique(FLWFLQ_GOs)


#GOobject and function already made when testing the prior gene




GOmatches_LOC117151569 <- find_gene_in_all_GO(
  LOC117151569,
  significant_genes_per_GO.FLWFLQ
)

GOmatches_LOC117151569



```






Okay, we aren't finding the Gene among our go terms. There are a few possibilities: 1) the function doesn't work, 2) the gene isn't in the gene universe, 3) the gene is among the NAs removed, and 4) the gene hasn't been annotated. I will assess these in that order.

### 1) Test gene locating function
The same function as before, so we know it works.


### 2) See if the gene is in the universe
Checking to make sure it is in the gene universe.
```{r geneuniversecheck_LOC117151569}

LOC117151569 %in% geneUniverse

```

The gene is present in the gene universe.

### 3) Was the gene removed as an outlier pre-Fisher test
Checking to make sure it is not one of the NAs removed.
```{r outliercheck_LOC117151569}

LOC117151569 %in% FLWFLQ_napos

```

It is not a gene that was removed as an outlier.


### 4) Has the gene been annotated
Checking to make sure the gene has been annotated (i.e., is in your gene2Mapping annotation file).
```{r annotationcheck_LOC117151569}

LOC117151569 %in% gene2mapping

```

The gene is not in the mapping file. Therefore, I am concluding that **LOC117151569** hasn't been annotated.







