---
title: "B. impatiens Maternal Brain Differential Gene Expression Analysis"
author: "M.A. Hagadorn"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = TRUE,
	highlight = TRUE,
	tidy.opts = list(width.cutoff = 80), tidy = TRUE)
```


```{r loadpackages, warning=FALSE, echo=FALSE, message=FALSE, include=FALSE}
#Library Necessary Packages
library(DESeq2) 
library(ggplot2)
library(scales) # needed for oob parameter
library(viridis) #color scale on MA plot
library(gridExtra)
library(gtable)
library(grid)
library(knitr)
library(reshape2)
library(cowplot)
library(ggdendro)
library(kableExtra)
library(writexl)

options(stringsAsFactors = FALSE)
```


```{r packageinfo}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```

I used the following tutorials while preparing my data and working through the process:
https://genviz.org/module-04-expression/0004/02/01/DifferentialExpression/
https://www.bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html

#Load Data
An initial filtering step occurred prior to this analysis. This included removing 3,332 genes using the **filterByExpr()** function in the edgeR. The remaining expression data included 9,593 genes. Two samples were also clear outliers when examing the PCA plots of the filtered gene data. These two samples were removed. Thus, the subsequent DEG input is countdata for 9,593 genes across 34 samples. Note: the inputs for DEG and WGCNA are identical.
```{r loadcountdata}
#load in the same count matrix used for WGCNA
rawCounts0 <- read.csv2("../../../RNASeqAnalysis/BimpMB_RNAanalysis/Filtering/GECountData_GenesOutliers_Filtered.csv", sep=";")
head(rawCounts0)
dim(rawCounts0)


#Read in with samples as column names and gene as column 1



#load in the sample mappings
sampleData0 <- read.delim("../../../RNASeqAnalysis/BimpMB_RNAanalysis/Filtering/sampleData_OutliersRemoved.csv", sep=";")
head(sampleData0)

# Also save a copy for later
#sampleData <- sampleData0
```



# Format Data
Now we need to get the data in the right format for conducting the analyses. \ 

To do so, we need to store readcounts and intermediate calculations needed for the expression analysis in an object of class **DESeqDataSet**, which is specific to DEseq2. This object also stores the dispersion and log2 fold changes. Analysis Formula: "~" followed by "+" separating factors. The factor of interest should be listed last and the control group within those factors should be listed first when viewing levels of the variable.\ 
 
Notes from tutorial:\ 
"When specifying the formula it should take the form of a “~” followed by “+” separating factors. When using the default DEseq2 parameters the factor of interest (tissue type in this case) should be specified last and the control within that factor should be first when viewing the levels() for that variable."

## Convert count data to matrix of DEseq2 appropriate format
Make rows genes and columns samples.
```{r counttomat}
# Convert count data to a matrix of appropriate form that DEseq2 can read
colnames(rawCounts0)[1] <- c("GeneID")
geneID <- rawCounts0$GeneID


#colnames(rawCounts) == Gene.ID --> Gene.Name --> and samples as columns

##THE FORMAT DESEQ2 WANTS IS ROWS = GENES & COLUMNS = SAMPLES
sampleIndex <- grepl("K\\d+", colnames(rawCounts0)) #ID samples as those starting with a K

#makes a new matrix that subset the original for only samples (i.e., not the gene id or name)
rawCounts <- as.matrix(rawCounts0[,sampleIndex])
rawCounts <- apply(rawCounts, 2, as.numeric)

#rawCounts <- as.numeric(rawCounts)
dim(rawCounts)
head(rawCounts)

#save the GeneID as the row names
rownames(rawCounts) <- geneID
head(rawCounts)
```


## Convert Sample Variables to DEseq2 appropriate format
Make rows samples and columns specific variables, keeping only the ones we want to analyze. Assign as factors
```{r samplemapping}
# Convert sample variable mappings to an appropriate form that DESeq2 can read
#head(sampleData0)
print(sampleData0[1:5, 1:7])

#make new df to work with moving forward
sampleData <- sampleData0
#Assigning the sample names as rownames
rownames(sampleData) <- sampleData$sample_id

#columns/variables to keep--column name
keeperCols <- c("tx_group", "tx_category", "colony", "caste")

#####Update the TX_category name to include no dashes
# Use replace() to replace the names in the 'Names' column
sampleData$tx_category <- replace(sampleData$tx_category, sampleData$tx_category %in% "foundress-like", "foundresslike")
sampleData$tx_category <- replace(sampleData$tx_category, sampleData$tx_category %in% "maternal-like", "maternallike")
sampleData$tx_category <- replace(sampleData$tx_category, sampleData$tx_category %in% "queen-like", "queenlike")


#subset sample data (sample names as row names) by the variables of interest
sampleData <- sampleData[, keeperCols]
sampleData$tx_group <- factor(sampleData$tx_group)
sampleData$tx_category <- factor(sampleData$tx_category)
sampleData$colony <- factor(sampleData$colony)
sampleData$caste <- factor(sampleData$caste)
head(sampleData)

```



## Sort Order: make sure the columns of count data are in the same order as the rows of sample mapping
From the tutorial:\ 
"For DEseq2 to work properly the column names of the count matrix must be in the same order as the row names of the sample mapping data, to ensure this we re-order the column names of the count data and run a check to ensure this has occurred correctly."\ 

Followed by:\ 
"To take advantage of the default settings of DEseq2 the control of the variable of interest, in our case the tissue type, should be the first element in the levels of that variable. Because we have more than 2 conditions for this variable we will not be taking advantage of the default settings however it’s good to get into the practice of doing this so we do it here."
```{r ordermatching}
# Put the columns of the count data in the same order as rows names of the sample mapping, then make sure it worked
rawCounts <- rawCounts[,unique(rownames(sampleData))] 
all(colnames(rawCounts) == rownames(sampleData))

# Order the tx groups so that it is sensible and make sure the control sample is first: FL -> ML -> QL
sampleData$tx_group <- factor(sampleData$tx_group, levels=c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))
sampleData$tx_category <- factor(sampleData$tx_category, levels=c("foundresslike", "maternallike", "queenlike"))


# Create the DEseq2DataSet object
  #NOTE ADDING THE TRUNC() BECAUSE WONT ACCEPT DECIMALS
deseq2Data <- DESeqDataSetFromMatrix(countData=trunc(rawCounts), colData=sampleData, design= ~ tx_group)
#deseq2Data_m2 <- DESeqDataSetFromMatrix(countData=trunc(rawCounts), colData=sampleData, design= ~ caste + tx_category)
```


## Assess filter criteria
The first line of code should return zero genes (rows), i.e., no genes are unexpressed in any sample. The second should return all genes and samples. This is because we have already filtered the genes using edgeR and removed outliers via PCA results. I am just using this step to verify.
```{r preprocessing-1}
#From Karens approach just modified for this format
dim(deseq2Data[rowSums(counts(deseq2Data)) == 0,]) # How many genes are unexpressed in any samples?
#counts(deseq2Data) give the counts for each gene by sample
#rowSums(counts(deseq2Data)) == 0; a read out of the false true status for rowSums == 0
#table(rowSums(counts(deseq2Data)) == 0); give FALSE vs TRUE numbers

#greater than 10
dim(deseq2Data[rowSums(counts(deseq2Data)) > 10, ])

```




# Run DESeq--MODEL: Counts ~ tx_group w/ levels=c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ")
```{r runDEseq}
deseq2Data <- DESeq(deseq2Data)

#####Make to use with heatmap#####
# Transform count data using the variance stablilizing transform
#vst() = variance stablizing transformation--provides a faster estimation of dispersion
deseq2VST <- vst(deseq2Data)

# Convert the DESeq transformed object to a data frame
deseq2VST <- assay(deseq2VST)
deseq2VST <- as.data.frame(deseq2VST)
deseq2VST$Gene <- rownames(deseq2VST)
dim(deseq2VST)
print(deseq2VST[1:5,1:6])


# Save for use in subsequent parts
save(deseq2Data, deseq2VST, sampleData, file = "DEG_data") 


#assign LFC and pthreshold
LFC <- 1.5
adjP <- 0.05
```

## Across Castes

### FLW V FLQ

```{r extractDEseq_FLWFLQ}
# Extract differential expression results
# For "tx_group" perform primary vs normal comparison
res_FLWFLQ <- results(deseq2Data, contrast=c("tx_group", "FLW", "FLQ"), independentFiltering=FALSE, pAdjustMethod = "BH")

# View summary of results
summary(res_FLWFLQ, alpha=0.05)
summary(res_FLWFLQ, alpha=0.10)
summary(res_FLWFLQ, alpha=0.01)
```

MA plots: display the log ratio (M) vs an average (A) in order to visualize the differences between two groups. \ 

Expectations of an MA plot = gene expression to remain consistent between conditions (NULL), so the MA plot should be similar to the shape of the trumpet.
```{r MA_FLWFLQ, warning=FALSE, fig.height=5}
# Using DEseq2 built in method
plotMA(res_FLWFLQ, alpha=0.05)
```


```{r ggplotsMA_FLWFLQ, warning=FALSE}
# Coerce to a data frame
res_FLWFLQ <- as.data.frame(res_FLWFLQ)

# Examine this data frame
head(res_FLWFLQ)

# Set a boolean column for significance
res_FLWFLQ$significant <- ifelse(res_FLWFLQ$padj < adjP, "Significant", NA)

# Let's add some more detail
ggplot(res_FLWFLQ, aes(baseMean, log2FoldChange, colour=padj)) +
geom_point(size=1) +
scale_y_continuous(limits=c(-3, 3), oob=squish) + 
scale_x_log10() + 
geom_hline(yintercept = 0, colour="darkorchid4", size=1, linetype="longdash") +
labs(x="mean of normalized counts", y="log fold change") + 
scale_colour_viridis(direction=-1, trans='sqrt') + 
theme_bw() + 
  geom_density_2d(colour="black", size=1)
```

#### Remove outliers and Subset by TX

```{r outliers_FLWFLQ}
#VERIFY COUNT
#Assess number of genes with adjustedp=NA; N=14
count(is.na(res_FLWFLQ$padj))

#VERIFY SAME POSITIONS ACROSS ALL PAIRWISE
FLWFLQ_napos <- which(is.na(res_FLWFLQ$padj), arr.ind = TRUE)


#Create a new data frame to use going forward W/O those padj NAs
deseq2ResDF_FLWFLQ.noNA <- res_FLWFLQ[-FLWFLQ_napos,]


# Keep only the significantly differentiated genes where the fold-change was at least at threshold
sigGenes_FLWFLQ <- rownames(deseq2ResDF_FLWFLQ.noNA[deseq2ResDF_FLWFLQ.noNA$padj <= adjP & 
            abs(deseq2ResDF_FLWFLQ.noNA$log2FoldChange) > LFC,])

length(sigGenes_FLWFLQ)

deseq2VST_FLWFLQ <- deseq2VST[deseq2VST$Gene %in% sigGenes_FLWFLQ,]


# First compare wide vs long version
deseq2VST_FLWFLQ_wide <- deseq2VST_FLWFLQ
deseq2VST_FLWFLQ_long <- melt(deseq2VST_FLWFLQ, id.vars=c("Gene")) #needed for ggplot

print(deseq2VST_FLWFLQ_wide[1:5,1:7])
head(deseq2VST_FLWFLQ_long)


###ONLY FLW AND FLQ SAMPLES
FLWFLQ_samplessubset <- rownames(sampleData[which(sampleData$tx_group=="FLW" | sampleData$tx_group=="FLQ"),])
#make empty dataframe for saving in the for loop
deseq2VST_long_FLWFLQsubset <- data.frame()

for (id in FLWFLQ_samplessubset) {
  subset_id <- deseq2VST_FLWFLQ_long[deseq2VST_FLWFLQ_long$variable == id, ]  # Subset df by the current value of id
  # at to the the subset dataframe each iteration
  deseq2VST_long_FLWFLQsubset <- rbind(deseq2VST_long_FLWFLQsubset, subset_id)
}

#verifying status of data frame
# Check if all ids in the variable column of the subsetted list are from the allowed id per the TX group used for the pairwise


# Output the result
invalid_names <- deseq2VST_long_FLWFLQsubset$variable[!deseq2VST_long_FLWFLQsubset$variable %in% FLWFLQ_samplessubset]
if (length(invalid_names) > 0) {
  print(paste("Invalid IDs in the DF:", paste(invalid_names, collapse = ", ")))
} else {
  print("All IDs in the DF are valid.")
}

```

```{r makeheatmap_FLWFLQ, eval=FALSE}
# Make a heatmap
ggplot(deseq2VST_long_FLWFLQsubset, aes(x=variable, y=Gene, fill=value)) + 
  geom_raster() + 
  scale_fill_viridis(trans="sqrt") + 
  theme(axis.text.x=element_text(angle=65, hjust=1), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())
```

```{r clusteredheatmap_FLWFLQ, fig.height=10}
# Convert the significant genes back to a matrix for clustering
deseq2VSTMatrix_FLWFLQ <- dcast(deseq2VST_long_FLWFLQsubset, Gene ~ variable)
rownames(deseq2VSTMatrix_FLWFLQ) <- deseq2VSTMatrix_FLWFLQ$Gene
deseq2VSTMatrix_FLWFLQ$Gene <- NULL

# Compute a distance calculation on both dimensions of the matrix
#dist() computes the distance based on the rows of the df
#in the current format, gives distance based on genes
distanceGene_FLWFLQ <- dist(deseq2VSTMatrix_FLWFLQ)
#transpose to get distance based on samples
distanceSample_FLWFLQ <- dist(t(deseq2VSTMatrix_FLWFLQ))

# Cluster based on the distance calculations
#hclust: hierarchical cluster
clusterGene_FLWFLQ <- hclust(distanceGene_FLWFLQ, method="average")
clusterSample_FLWFLQ <- hclust(distanceSample_FLWFLQ, method="average")

#get specific colors by TX group:
clustorder_FLWFLQ <- clusterSample_FLWFLQ$order
color_vector_FLWFLQ <- vector("character", length(clustorder_FLWFLQ))
# Loop through each name in clustorder_FLWFLQ
for (i in 1:length(clustorder_FLWFLQ)) {
  name <- clustorder_FLWFLQ[i]
 
  # Check if the name exists in rownames of sampleData and if the tx_group is "FLW"
  if (name >= 6) {  #6 not 7 because this FLQ have one less sample
    color_vector_FLWFLQ[i] <- "blue"  # Assign blue if location is 6 or above
  } else {
    color_vector_FLWFLQ[i] <- "goldenrod"   # Assign blue if the name does not match "FLW"
  }
}


color_vector_FLWFLQ_xcol <- c("goldenrod","goldenrod","goldenrod","goldenrod","goldenrod","blue","blue","blue","blue","blue","blue")

color_vector_FLWFLQ == color_vector_FLWFLQ_xcol


# Construct a dendogram for samples
sampleModel_FLWFLQ <- as.dendrogram(clusterSample_FLWFLQ)
sampleDendrogramData_FLWFLQ <- segment(dendro_data(sampleModel_FLWFLQ, type = "rectangle"))
sampleDendrogram_FLWFLQ <- ggplot(sampleDendrogramData_FLWFLQ) + 
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) +
  theme_dendro()

# Re-factor samples for ggplot2
#use to match dendogram
deseq2VST_long_FLWFLQsubset$variable <- factor(deseq2VST_long_FLWFLQsubset$variable,levels=clusterSample_FLWFLQ$labels[clusterSample_FLWFLQ$order])

# Construct the heatmap. note that at this point we have only clustered the samples NOT the genes
heatmap_FLWFLQ <- ggplot(deseq2VST_long_FLWFLQsubset, aes(x=variable, y=Gene, fill=value)) + 
  geom_raster() + 
  scale_fill_viridis(trans="sqrt") +
  theme(axis.text.x=element_text(angle=65, hjust=1, color = color_vector_FLWFLQ_xcol),
        #axis.text.y=element_blank(), 
        axis.ticks.y=element_blank(),
        plot.margin = unit(c(-.75, 0, 0, 0), "cm"))

# Combine the dendrogram and the heatmap
#install.packages("gridExtra")

title <- ggdraw() + draw_label("Significant Genes for FL workers (blue) v FL queens (gold)", fontface='bold')
plot_grid(title, NULL, sampleDendrogram_FLWFLQ, NULL, heatmap_FLWFLQ, NULL, align= "v", ncol=2, nrow=3, rel_heights = c(.05,.1,1,1,1,1), rel_widths = c(2.7,.1,2.7,.1,2.5,.1))
```


```{r regdirection_FLWFLQ}
DEGs_FLWFLQ <- deseq2ResDF_FLWFLQ.noNA[rownames(deseq2ResDF_FLWFLQ.noNA) %in% sigGenes_FLWFLQ,]

#write the DEGs and data to CSV
write.csv(DEGs_FLWFLQ,"SigGeneLists/DEGs_FLWFLQ.csv", row.names = TRUE)
DEG_genelist_FLWFLQ <- rownames(DEGs_FLWFLQ)
#write gene list to txt
write.table(DEG_genelist_FLWFLQ,"SigGeneLists/DEG_genelist_FLWFLQ.txt", row.names = FALSE, col.names = FALSE, quote=FALSE)


DEGs.FCdir_FLWFLQ <- DEGs_FLWFLQ[,c("log2FoldChange", "padj")]
length(DEGs.FCdir_FLWFLQ$log2FoldChange)


#print in document
DEGs.FCdir_FLWFLQ 


###Write All Results to document

write.csv(res_FLWFLQ,"DEGFullResults/DEGananalysisResults_FLWFLQ.csv", row.names = TRUE)

```


### MLW V MLQ

From here on out, I will *echo=FALSE* for the parts of the code that generate the heat map. I am doing so to save space and making the document easier to follow.

```{r extractDEseq_MLWMLQ}
# Extract differential expression results
# For "tx_group" perform primary vs normal comparison
res_MLWMLQ <- results(deseq2Data, contrast=c("tx_group", "MLW", "MLQ"), independentFiltering=FALSE, pAdjustMethod = "BH")

# View summary of results
summary(res_MLWMLQ, alpha=0.05)
summary(res_MLWMLQ, alpha=0.10)
summary(res_MLWMLQ, alpha=0.01)
```

```{r MA_MLWMLQ, warning=FALSE, fig.height=5, echo=FALSE}
# Using DEseq2 built in method
plotMA(res_MLWMLQ, alpha=0.05)
```


```{r ggplotsMA_MLWMLQ, warning=FALSE}
# Coerce to a data frame
res_MLWMLQ <- as.data.frame(res_MLWMLQ)

# Examine this data frame
head(res_MLWMLQ)

# Set a boolean column for significance
res_MLWMLQ$significant <- ifelse(res_MLWMLQ$padj < adjP, "Significant", NA)

# Let's add some more detail
ggplot(res_MLWMLQ, aes(baseMean, log2FoldChange, colour=padj)) +
geom_point(size=1) +
scale_y_continuous(limits=c(-3, 3), oob=squish) +
scale_x_log10() +
geom_hline(yintercept = 0, colour="darkorchid4", size=1, linetype="longdash") +
labs(x="mean of normalized counts", y="log fold change") +
scale_colour_viridis(direction=-1, trans='sqrt') +
theme_bw() +
  geom_density_2d(colour="black", size=1)
```


#### Remove outliers and Subset by TX

```{r outliers_MLWMLQ}
#VERIFY COUNT
#Assess number of genes with adjustedp=NA; N=14
count(is.na(res_MLWMLQ$padj))

#VERIFY SAME POSITIONS ACROSS ALL PAIRWISE
MLWMLQ_napos <- which(is.na(res_MLWMLQ$padj), arr.ind = TRUE)


#Create a new data frame to use going forward W/O those padj NAs
deseq2ResDF_MLWMLQ.noNA <- res_MLWMLQ[-MLWMLQ_napos,]


# Keep only the significantly differentiated genes where the fold-change was at least 3
sigGenes_MLWMLQ <- rownames(deseq2ResDF_MLWMLQ.noNA[deseq2ResDF_MLWMLQ.noNA$padj <= adjP &
            abs(deseq2ResDF_MLWMLQ.noNA$log2FoldChange) > LFC,])

length(sigGenes_MLWMLQ)

deseq2VST_MLWMLQ <- deseq2VST[deseq2VST$Gene %in% sigGenes_MLWMLQ,]


# First compare wide vs long version
deseq2VST_MLWMLQ_wide <- deseq2VST_MLWMLQ
deseq2VST_MLWMLQ_long <- melt(deseq2VST_MLWMLQ, id.vars=c("Gene")) #needed for ggplot

print(deseq2VST_MLWMLQ_wide[1:5,1:7])
head(deseq2VST_MLWMLQ_long)


###ONLY MLW AND FLQ SAMPLES
MLWMLQ_samplessubset <- rownames(sampleData[which(sampleData$tx_group=="MLW" | sampleData$tx_group=="MLQ"),])
#make empty dataframe for saving in the for loop
deseq2VST_long_MLWMLQsubset <- data.frame()

for (id in MLWMLQ_samplessubset) {
  subset_id <- deseq2VST_MLWMLQ_long[deseq2VST_MLWMLQ_long$variable == id, ]  # Subset df by the current value of id
  # at to the the subset dataframe each iteration
  deseq2VST_long_MLWMLQsubset <- rbind(deseq2VST_long_MLWMLQsubset, subset_id)
}

#verifying status of data frame
# Check if all ids in the variable column of the subsetted list are from the allowed id per the TX group used for the pairwise


# Output the result
invalid_names <- deseq2VST_long_MLWMLQsubset$variable[!deseq2VST_long_MLWMLQsubset$variable %in% MLWMLQ_samplessubset]
if (length(invalid_names) > 0) {
  print(paste("Invalid IDs in the DF:", paste(invalid_names, collapse = ", ")))
} else {
  print("All IDs in the DF are valid.")
}

```

```{r makeheatmap_MLWMLQ, eval=FALSE, echo=FALSE}
# Make a heatmap
ggplot(deseq2VST_long_MLWMLQsubset, aes(x=variable, y=Gene, fill=value)) +
  geom_raster() +
  scale_fill_viridis(trans="sqrt") +
  theme(axis.text.x=element_text(angle=65, hjust=1),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r clusteredheatmap_MLWMLQ, echo=FALSE, fig.height=5.5}
# Convert the significant genes back to a matrix for clustering
deseq2VSTMatrix_MLWMLQ <- dcast(deseq2VST_long_MLWMLQsubset, Gene ~ variable)
rownames(deseq2VSTMatrix_MLWMLQ) <- deseq2VSTMatrix_MLWMLQ$Gene
deseq2VSTMatrix_MLWMLQ$Gene <- NULL

# Compute a distance calculation on both dimensions of the matrix
#dist() computes the distance based on the rows of the df
#in the current format, gives distance based on genes
distanceGene_MLWMLQ <- dist(deseq2VSTMatrix_MLWMLQ)
#transpose to get distance based on samples
distanceSample_MLWMLQ <- dist(t(deseq2VSTMatrix_MLWMLQ))

# Cluster based on the distance calculations
#hclust: hierarchical cluster
clusterGene_MLWMLQ <- hclust(distanceGene_MLWMLQ, method="average")
clusterSample_MLWMLQ <- hclust(distanceSample_MLWMLQ, method="average")

#get specific colors by TX group:
clustorder_MLWMLQ <- clusterSample_MLWMLQ$order
color_vector_MLWMLQ <- vector("character", length(clustorder_MLWMLQ))
# Loop through each name in clustorder_MLWMLQ
for (i in 1:length(clustorder_MLWMLQ)) {
  name <- clustorder_MLWMLQ[i]

  # Check if the name exists in rownames of sampleData and if the tx_group is "FLW"
  if (name >= 7) {
    color_vector_MLWMLQ[i] <- "blue"  # Assign green if the name matches "FLW" in tx_group
  } else {
    color_vector_MLWMLQ[i] <- "goldenrod"   # Assign blue if the name does not match "FLW"
  }
}



color_vector_MLWMLQ_xcol <- c("blue","goldenrod","goldenrod","goldenrod","goldenrod","goldenrod","goldenrod","blue","blue","blue","blue","blue")

color_vector_MLWMLQ == color_vector_MLWMLQ_xcol


# color_vector_FLWFLQ_xcol <- c("goldenrod","goldenrod","goldenrod","blue","blue","blue","blue","blue","blue","goldenrod","goldenrod","goldenrod")
#
# color_vector_FLWFLQ == color_vector_FLWFLQ_xcol


# Construct a dendogram for samples
sampleModel_MLWMLQ <- as.dendrogram(clusterSample_MLWMLQ)
sampleDendrogramData_MLWMLQ <- segment(dendro_data(sampleModel_MLWMLQ, type = "rectangle"))
sampleDendrogram_MLWMLQ <- ggplot(sampleDendrogramData_MLWMLQ) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) +
  theme_dendro()

# Re-factor samples for ggplot2
#use to match dendogram
deseq2VST_long_MLWMLQsubset$variable <- factor(deseq2VST_long_MLWMLQsubset$variable,levels=clusterSample_MLWMLQ$labels[clusterSample_MLWMLQ$order])

# Construct the heatmap. note that at this point we have only clustered the samples NOT the genes
heatmap_MLWMLQ <- ggplot(deseq2VST_long_MLWMLQsubset, aes(x=variable, y=Gene, fill=value)) +
  geom_raster() +
  scale_fill_viridis(trans="sqrt") +
  theme(axis.text.x=element_text(angle=65, hjust=1, color = color_vector_MLWMLQ_xcol),
        #axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        plot.margin = unit(c(-.75, 0, 0, 0), "cm"))


titleMLWMLQ <- ggdraw() + draw_label("Significant Genes for ML workers (blue) v ML queens (gold)", fontface='bold')
plot_grid(titleMLWMLQ, NULL, sampleDendrogram_MLWMLQ, NULL, heatmap_MLWMLQ, NULL, align= "v", ncol=2, nrow=3, rel_heights = c(.1,.3,1,1,1,1), rel_widths = c(2.7,.1,2.7,.1,2.5,.1))
```

```{r regdirection_MLWMLQ, echo=FALSE}
DEGs_MLWMLQ <- deseq2ResDF_MLWMLQ.noNA[rownames(deseq2ResDF_MLWMLQ.noNA) %in% sigGenes_MLWMLQ,]

#write the DEGs and data to CSV
write.csv(DEGs_MLWMLQ,"SigGeneLists/DEGs_MLWMLQ.csv", row.names = TRUE)
DEG_genelist_MLWMLQ <- rownames(DEGs_MLWMLQ)
#write gene list to txt
write.table(DEG_genelist_MLWMLQ,"SigGeneLists/DEG_genelist_MLWMLQ.txt", row.names = FALSE, col.names = FALSE, quote=FALSE)



DEGs.FCdir_MLWMLQ <- DEGs_MLWMLQ[,c("log2FoldChange", "padj")]
length(DEGs.FCdir_MLWMLQ$log2FoldChange)


DEGs.FCdir_MLWMLQ


write.csv(res_MLWMLQ,"DEGFullResults/DEGananalysisResults_MLWMLQ.csv", row.names = TRUE)


```


### QLW V QLQ

```{r extractDEseq_QLWQLQ}
# Extract differential expression results
# For "tx_group" perform primary vs normal comparison
res_QLWQLQ <- results(deseq2Data, contrast=c("tx_group", "QLW", "QLQ"), independentFiltering=FALSE, pAdjustMethod = "BH")

# View summary of results
summary(res_QLWQLQ, alpha=0.05)
summary(res_QLWQLQ, alpha=0.10)
summary(res_QLWQLQ, alpha=0.01)
```


```{r MA_QLWQLQ, warning=FALSE, fig.height=5}
# Using DEseq2 built in method
plotMA(res_QLWQLQ, alpha=0.05)
```


```{r ggplotsMA_QLWQLQ, warning=FALSE}
# Coerce to a data frame
res_QLWQLQ <- as.data.frame(res_QLWQLQ)

# Examine this data frame
head(res_QLWQLQ)

# Set a boolean column for significance
res_QLWQLQ$significant <- ifelse(res_QLWQLQ$padj < adjP, "Significant", NA)

# Let's add some more detail
ggplot(res_QLWQLQ, aes(baseMean, log2FoldChange, colour=padj)) +
geom_point(size=1) +
scale_y_continuous(limits=c(-3, 3), oob=squish) +
scale_x_log10() +
geom_hline(yintercept = 0, colour="darkorchid4", size=1, linetype="longdash") +
labs(x="mean of normalized counts", y="log fold change") +
scale_colour_viridis(direction=-1, trans='sqrt') +
theme_bw() +
  geom_density_2d(colour="black", size=1)
```


#### Remove outliers and Subset by TX

```{r outliers_QLWQLQ}
#VERIFY COUNT
#Assess number of genes with adjustedp=NA; N=14
count(is.na(res_QLWQLQ$padj))

#VERIFY SAME POSITIONS ACROSS ALL PAIRWISE
QLWQLQ_napos <- which(is.na(res_QLWQLQ$padj), arr.ind = TRUE)


#Create a new data frame to use going forward W/O those padj NAs
deseq2ResDF_QLWQLQ.noNA <- res_QLWQLQ[-QLWQLQ_napos,]


# Keep only the significantly differentiated genes where the fold-change was at least 3
sigGenes_QLWQLQ <- rownames(deseq2ResDF_QLWQLQ.noNA[deseq2ResDF_QLWQLQ.noNA$padj <= adjP &
            abs(deseq2ResDF_QLWQLQ.noNA$log2FoldChange) > LFC,])

length(sigGenes_QLWQLQ)

deseq2VST_QLWQLQ <- deseq2VST[deseq2VST$Gene %in% sigGenes_QLWQLQ,]


# First compare wide vs long version
deseq2VST_QLWQLQ_wide <- deseq2VST_QLWQLQ
deseq2VST_QLWQLQ_long <- melt(deseq2VST_QLWQLQ, id.vars=c("Gene")) #needed for ggplot

print(deseq2VST_QLWQLQ_wide[1:3,1:7])
head(deseq2VST_QLWQLQ_long)


###ONLY QLW AND FLQ SAMPLES
QLWQLQ_samplessubset <- rownames(sampleData[which(sampleData$tx_group=="QLW" | sampleData$tx_group=="QLQ"),])
#make empty dataframe for saving in the for loop
deseq2VST_long_QLWQLQsubset <- data.frame()

for (id in QLWQLQ_samplessubset) {
  subset_id <- deseq2VST_QLWQLQ_long[deseq2VST_QLWQLQ_long$variable == id, ]  # Subset df by the current value of id
  # at to the the subset dataframe each iteration
  deseq2VST_long_QLWQLQsubset <- rbind(deseq2VST_long_QLWQLQsubset, subset_id)
}

#verifying status of data frame
# Check if all ids in the variable column of the subsetted list are from the allowed id per the TX group used for the pairwise


# Output the result
invalid_names <- deseq2VST_long_QLWQLQsubset$variable[!deseq2VST_long_QLWQLQsubset$variable %in% QLWQLQ_samplessubset]
if (length(invalid_names) > 0) {
  print(paste("Invalid IDs in the DF:", paste(invalid_names, collapse = ", ")))
} else {
  print("All IDs in the DF are valid.")
}

```

```{r makeheatmap_QLWQLQ, eval=FALSE, echo=FALSE}
# Make a heatmap
ggplot(deseq2VST_long_QLWQLQsubset, aes(x=variable, y=Gene, fill=value)) +
  geom_raster() +
  scale_fill_viridis(trans="sqrt") +
  theme(axis.text.x=element_text(angle=65, hjust=1),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r clusteredheatmap_QLWQLQ, fig.height=5.5, echo=FALSE}
# Convert the significant genes back to a matrix for clustering
deseq2VSTMatrix_QLWQLQ <- dcast(deseq2VST_long_QLWQLQsubset, Gene ~ variable)
rownames(deseq2VSTMatrix_QLWQLQ) <- deseq2VSTMatrix_QLWQLQ$Gene
deseq2VSTMatrix_QLWQLQ$Gene <- NULL

# Compute a distance calculation on both dimensions of the matrix
#dist() computes the distance based on the rows of the df
#in the current format, gives distance based on genes
distanceGene_QLWQLQ <- dist(deseq2VSTMatrix_QLWQLQ)
#transpose to get distance based on samples
distanceSample_QLWQLQ <- dist(t(deseq2VSTMatrix_QLWQLQ))

# Cluster based on the distance calculations
#hclust: hierarchical cluster
clusterGene_QLWQLQ <- hclust(distanceGene_QLWQLQ, method="average")
clusterSample_QLWQLQ <- hclust(distanceSample_QLWQLQ, method="average")

#get specific colors by TX group:
clustorder_QLWQLQ <- clusterSample_QLWQLQ$order
color_vector_QLWQLQ <- vector("character", length(clustorder_QLWQLQ))
# Loop through each name in clustorder_QLWQLQ
for (i in 1:length(clustorder_QLWQLQ)) {
  name <- clustorder_QLWQLQ[i]

  # Check if the name exists in rownames of sampleData and if the tx_group is "FLW"
  if (name >= 6) {
    color_vector_QLWQLQ[i] <- "blue"  # Assign green if the name matches "FLW" in tx_group
  } else {
    color_vector_QLWQLQ[i] <- "goldenrod"   # Assign blue if the name does not match "FLW"
  }
}



color_vector_QLWQLQ_xcol <- c("goldenrod","goldenrod","goldenrod","goldenrod","goldenrod", "blue","blue","blue","blue","blue","blue")

color_vector_QLWQLQ == color_vector_QLWQLQ_xcol



# Construct a dendogram for samples
sampleModel_QLWQLQ <- as.dendrogram(clusterSample_QLWQLQ)
sampleDendrogramData_QLWQLQ <- segment(dendro_data(sampleModel_QLWQLQ, type = "rectangle"))
sampleDendrogram_QLWQLQ <- ggplot(sampleDendrogramData_QLWQLQ) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) +
  theme_dendro()

# Re-factor samples for ggplot2
#use to match dendogram
deseq2VST_long_QLWQLQsubset$variable <- factor(deseq2VST_long_QLWQLQsubset$variable,levels=clusterSample_QLWQLQ$labels[clusterSample_QLWQLQ$order])

# Construct the heatmap. note that at this point we have only clustered the samples NOT the genes
heatmap_QLWQLQ <- ggplot(deseq2VST_long_QLWQLQsubset, aes(x=variable, y=Gene, fill=value)) +
  geom_raster() +
  scale_fill_viridis(trans="sqrt") +
  theme(axis.text.x=element_text(angle=65, hjust=1, color = color_vector_QLWQLQ),
        #axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        plot.margin = unit(c(-.75, 0, 0, 0), "cm"))

# Combine the dendrogram and the heatmap
#install.packages("gridExtra")

titleQLWQLQ <- ggdraw() + draw_label("Significant Genes for QL workers (blue) v QL queens (gold)", fontface='bold')
plot_grid(titleQLWQLQ, NULL, sampleDendrogram_QLWQLQ, NULL, heatmap_QLWQLQ, NULL, align= "v", ncol=2, nrow=3, rel_heights = c(.1,.35,1,1,1,1), rel_widths = c(2.7,.1,2.7,.1,2.5,.1))
```


```{r regdirection_QLWQLQ, fig.height=7, echo=FALSE}
DEGs_QLWQLQ <- deseq2ResDF_QLWQLQ.noNA[rownames(deseq2ResDF_QLWQLQ.noNA) %in% sigGenes_QLWQLQ,]

#write the DEGs and data to CSV
write.csv(DEGs_QLWQLQ,"SigGeneLists/DEGs_QLWQLQ.csv", row.names = TRUE)
DEG_genelist_QLWQLQ <- rownames(DEGs_QLWQLQ)
#write gene list to txt
write.table(DEG_genelist_QLWQLQ,"SigGeneLists/DEG_genelist_QLWQLQ.txt", row.names = FALSE, col.names = FALSE, quote=FALSE)


DEGs.FCdir_QLWQLQ <- DEGs_QLWQLQ[,c("log2FoldChange", "padj")]
length(DEGs.FCdir_QLWQLQ$log2FoldChange)


DEGs.FCdir_QLWQLQ


#FOR GSEA
write.csv(res_QLWQLQ,"DEGFullResults/DEGananalysisResults_QLWQLQ.csv", row.names = TRUE)

```



## Within caste--WORKERS
### FLW V MLW

```{r extractDEseq_FLWMLW}
# Extract differential expression results
# For "tx_group" perform primary vs normal comparison
res_FLWMLW <- results(deseq2Data, contrast=c("tx_group", "FLW", "MLW"), independentFiltering=FALSE, pAdjustMethod = "BH")

# View summary of results
summary(res_FLWMLW, alpha=0.05)
summary(res_FLWMLW, alpha=0.10)
summary(res_FLWMLW, alpha=0.01)
```


```{r MA_FLWMLW, warning=FALSE}
# Using DEseq2 built in method
plotMA(res_FLWMLW, alpha=0.05)
```


```{r ggplotsMA_FLWMLW, warning=FALSE}
# Coerce to a data frame
res_FLWMLW <- as.data.frame(res_FLWMLW)

# Examine this data frame
head(res_FLWMLW)

# Set a boolean column for significance
res_FLWMLW$significant <- ifelse(res_FLWMLW$padj < adjP, "Significant", NA)

# Let's add some more detail
ggplot(res_FLWMLW, aes(baseMean, log2FoldChange, colour=padj)) +
geom_point(size=1) +
scale_y_continuous(limits=c(-3, 3), oob=squish) +
scale_x_log10() +
geom_hline(yintercept = 0, colour="darkorchid4", size=1, linetype="longdash") +
labs(x="mean of normalized counts", y="log fold change") +
scale_colour_viridis(direction=-1, trans='sqrt') +
theme_bw() +
  geom_density_2d(colour="black", size=1)
```


#### Remove outliers and Subset by TX

```{r outliers_FLWMLW}
#VERIFY COUNT
#Assess number of genes with adjustedp=NA; N=14
count(is.na(res_FLWMLW$padj))

#VERIFY SAME POSITIONS ACROSS ALL PAIRWISE
FLWMLW_napos <- which(is.na(res_FLWMLW$padj), arr.ind = TRUE)


#Create a new data frame to use going forward W/O those padj NAs
deseq2ResDF_FLWMLW.noNA <- res_FLWMLW[-FLWMLW_napos,]


# Keep only the significantly differentiated genes where the fold-change was at least 3
sigGenes_FLWMLW <- rownames(deseq2ResDF_FLWMLW.noNA[deseq2ResDF_FLWMLW.noNA$padj <= adjP &
            abs(deseq2ResDF_FLWMLW.noNA$log2FoldChange) > LFC,])

length(sigGenes_FLWMLW)

deseq2VST_FLWMLW <- deseq2VST[deseq2VST$Gene %in% sigGenes_FLWMLW,]


# First compare wide vs long version
deseq2VST_FLWMLW_wide <- deseq2VST_FLWMLW
deseq2VST_FLWMLW_long <- melt(deseq2VST_FLWMLW, id.vars=c("Gene")) #needed for ggplot

print(deseq2VST_FLWMLW_wide[1:4,1:7])
head(deseq2VST_FLWMLW_long)


###ONLY FLW AND FLQ SAMPLES
FLWMLW_samplessubset <- rownames(sampleData[which(sampleData$tx_group=="FLW" | sampleData$tx_group=="MLW"),])
#make empty dataframe for saving in the for loop
deseq2VST_long_FLWMLWsubset <- data.frame()

for (id in FLWMLW_samplessubset) {
  subset_id <- deseq2VST_FLWMLW_long[deseq2VST_FLWMLW_long$variable == id, ]  # Subset df by the current value of id
  # at to the the subset dataframe each iteration
  deseq2VST_long_FLWMLWsubset <- rbind(deseq2VST_long_FLWMLWsubset, subset_id)
}

#verifying status of data frame
# Check if all ids in the variable column of the subsetted list are from the allowed id per the TX group used for the pairwise


# Output the result
invalid_names <- deseq2VST_long_FLWMLWsubset$variable[!deseq2VST_long_FLWMLWsubset$variable %in% FLWMLW_samplessubset]
if (length(invalid_names) > 0) {
  print(paste("Invalid IDs in the DF:", paste(invalid_names, collapse = ", ")))
} else {
  print("All IDs in the DF are valid.")
}

```

```{r makeheatmap_FLWMLW, eval=FALSE, fig.height=3, echo=FALSE}
# Make a heatmap
ggplot(deseq2VST_long_FLWMLWsubset, aes(x=variable, y=Gene, fill=value)) +
  geom_raster() +
  scale_fill_viridis(trans="sqrt") +
  theme(axis.text.x=element_text(angle=65, hjust=1),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r clusteredheatmap_FLWMLW, fig.height=4, echo=FALSE}
# Convert the significant genes back to a atrix for clustering
deseq2VSTMatrix_FLWMLW <- dcast(deseq2VST_long_FLWMLWsubset, Gene ~ variable)
rownames(deseq2VSTMatrix_FLWMLW) <- deseq2VSTMatrix_FLWMLW$Gene
deseq2VSTMatrix_FLWMLW$Gene <- NULL

# Compute a distance calculation on both dimensions of the matrix
#dist() computes the distance based on the rows of the df
#in the current format, gives distance based on genes
distanceGene_FLWMLW <- dist(deseq2VSTMatrix_FLWMLW)
#transpose to get distance based on samples
distanceSample_FLWMLW <- dist(t(deseq2VSTMatrix_FLWMLW))

# Cluster based on the distance calculations
#hclust: hierarchical cluster
clusterGene_FLWMLW <- hclust(distanceGene_FLWMLW, method="average")
clusterSample_FLWMLW <- hclust(distanceSample_FLWMLW, method="average")

#get specific colors by TX group:
clustorder_FLWMLW <- clusterSample_FLWMLW$order
color_vector_FLWMLW <- vector("character", length(clustorder_FLWMLW))

# Loop through each name in clustorder_FLWMLW
for (i in 1:length(clustorder_FLWMLW)) {
  name <- clustorder_FLWMLW[i]

  # Check if the name exists in rownames of sampleData and if the tx_group is "FLW"
  if (name >= 7) {
    color_vector_FLWMLW[i] <- "yellowgreen"  # Assign green if the name matches "FLW" in tx_group
  } else {
    color_vector_FLWMLW[i] <- "coral"   # Assign blue if the name does not match "FLW"
  }
}



color_vector_FLWMLW_xcol <- c("coral","coral","coral","coral","yellowgreen","yellowgreen","yellowgreen","yellowgreen","coral","yellowgreen","coral","yellowgreen")

color_vector_FLWMLW == color_vector_FLWMLW_xcol


# Construct a dendogram for samples
sampleModel_FLWMLW <- as.dendrogram(clusterSample_FLWMLW)
sampleDendrogramData_FLWMLW <- segment(dendro_data(sampleModel_FLWMLW, type = "rectangle"))
sampleDendrogram_FLWMLW <- ggplot(sampleDendrogramData_FLWMLW) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) +
  theme_dendro()

# Re-factor samples for ggplot2
#use to match dendogram
deseq2VST_long_FLWMLWsubset$variable <- factor(deseq2VST_long_FLWMLWsubset$variable,levels=clusterSample_FLWMLW$labels[clusterSample_FLWMLW$order])

# Construct the heatmap. note that at this point we have only clustered the samples NOT the genes
heatmap_FLWMLW <- ggplot(deseq2VST_long_FLWMLWsubset, aes(x=variable, y=Gene, fill=value)) +
  geom_raster() +
  scale_fill_viridis(trans="sqrt") +
  theme(axis.text.x=element_text(angle=65, hjust=1, color = color_vector_FLWMLW),
        #axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        plot.margin = unit(c(-.75, 0, 0, 0), "cm"))

# Combine the dendrogram and the heatmap
#install.packages("gridExtra")


titleFLWMLW <- ggdraw() + draw_label("Significant Genes for FL workers (yellow-green) v ML workers (coral)", fontface='bold')
plot_grid(titleFLWMLW, NULL, sampleDendrogram_FLWMLW, NULL, heatmap_FLWMLW, NULL, align= "v", ncol=2, nrow=3, rel_heights = c(.1,.7,1,1,1,1), rel_widths = c(2.7,.1,2.7,.1,2.5,.1))
```


```{r regdirection_FLWMLW, fig.height=7, echo=FALSE}
DEGs_FLWMLW <- deseq2ResDF_FLWMLW.noNA[rownames(deseq2ResDF_FLWMLW.noNA) %in% sigGenes_FLWMLW,]

#write the DEGs and data to CSV
write.csv(DEGs_FLWMLW,"SigGeneLists/DEGs_FLWMLW.csv", row.names = TRUE)
DEG_genelist_FLWMLW <- rownames(DEGs_FLWMLW)
#write gene list to txt
write.table(DEG_genelist_FLWMLW,"SigGeneLists/DEG_genelist_FLWMLW.txt", row.names = FALSE, col.names = FALSE, quote=FALSE)


DEGs.FCdir_FLWMLW <- DEGs_FLWMLW[,c("log2FoldChange", "padj")]
length(DEGs.FCdir_FLWMLW$log2FoldChange)


DEGs.FCdir_FLWMLW


#FOR GSEA
write.csv(res_FLWMLW,"DEGFullResults/DEGananalysisResults_FLWMLW.csv", row.names = TRUE)
```


### FLW V QLW

```{r extractDEseq_FLWQLW}
# Extract differential expression results
# For "tx_group" perform primary vs normal comparison
res_FLWQLW <- results(deseq2Data, contrast=c("tx_group", "FLW", "QLW"), independentFiltering=FALSE, pAdjustMethod = "BH")

# View summary of results
summary(res_FLWQLW, alpha=0.05)
summary(res_FLWQLW, alpha=0.10)
summary(res_FLWQLW, alpha=0.01)
```


```{r MA_FLWQLW, warning=FALSE}
# Using DEseq2 built in method
plotMA(res_FLWQLW, alpha=0.05)
```


```{r ggplotsMA_FLWQLW, warning=FALSE}
# Coerce to a data frame
res_FLWQLW <- as.data.frame(res_FLWQLW)

# Examine this data frame
head(res_FLWQLW)

# Set a boolean column for significance
res_FLWQLW$significant <- ifelse(res_FLWQLW$padj < adjP, "Significant", NA)

#FOR GSEA Here because don't move on below
write.csv(res_FLWQLW,"DEGFullResults/DEGananalysisResults_FLWQLW.csv", row.names = TRUE)

# Let's add some more detail
ggplot(res_FLWQLW, aes(baseMean, log2FoldChange, colour=padj)) +
geom_point(size=1) +
scale_y_continuous(limits=c(-3, 3), oob=squish) +
scale_x_log10() +
geom_hline(yintercept = 0, colour="darkorchid4", size=1, linetype="longdash") +
labs(x="mean of normalized counts", y="log fold change") +
scale_colour_viridis(direction=-1, trans='sqrt') +
theme_bw() +
  geom_density_2d(colour="black", size=1)
```

No significant gene, so I will not move forward with producing a heatmap.
<!-- ####Remove outliers and Subset by TX -->
<!-- ```{r outliers_FLWQLW} -->
<!-- #VERIFY COUNT -->
<!-- #Assess number of genes with adjustedp=NA; N=14 -->
<!-- count(is.na(res_FLWQLW$padj)) -->

<!-- #VERIFY SAME POSITIONS ACROSS ALL PAIRWISE -->
<!-- _FLWQLW_napos <- which(is.na(res_FLWQLW$padj), arr.ind = TRUE) -->


<!-- #Create a new data frame to use going forward W/O those padj NAs -->
<!-- deseq2ResDF_FLWQLW.noNA <- res_FLWQLW[-_FLWQLW_napos,] -->


<!-- # Keep only the significantly differentiated genes where the fold-change was at least 3 -->
<!-- sigGenes_FLWQLW <- rownames(deseq2ResDF_FLWQLW.noNA[deseq2ResDF_FLWQLW.noNA$padj <= .01 &  -->
<!--             abs(deseq2ResDF_FLWQLW.noNA$log2FoldChange) > 3,]) -->

<!-- length(sigGenes_FLWQLW) -->

<!-- deseq2VST_FLWQLW <- deseq2VST[deseq2VST$Gene %in% sigGenes_FLWQLW,] -->


<!-- # First compare wide vs long version -->
<!-- deseq2VST_FLWQLW_wide <- deseq2VST_FLWQLW -->
<!-- deseq2VST_FLWQLW_long <- melt(deseq2VST_FLWQLW, id.vars=c("Gene")) #needed for ggplot -->

<!-- print(deseq2VST_FLWQLW_wide[1:5,1:7]) -->
<!-- head(deseq2VST_FLWQLW_long) -->


<!-- ###ONLY FLW AND FLQ SAMPLES -->
<!-- FLWQLW_samplessubset <- rownames(sampleData[which(sampleData$tx_group=="FLW" | sampleData$tx_group=="QLW"),]) -->
<!-- #make empty dataframe for saving in the for loop -->
<!-- deseq2VST_long_FLWQLWsubset <- data.frame() -->

<!-- for (id in FLWQLW_samplessubset) { -->
<!--   subset_id <- deseq2VST_FLWQLW_long[deseq2VST_FLWQLW_long$variable == id, ]  # Subset df by the current value of id -->
<!--   # at to the the subset dataframe each iteration -->
<!--   deseq2VST_long_FLWQLWsubset <- rbind(deseq2VST_long_FLWQLWsubset, subset_id) -->
<!-- } -->

<!-- #verifying status of data frame -->
<!-- # Check if all ids in the variable column of the subsetted list are from the allowed id per the TX group used for the pairwise -->


<!-- # Output the result -->
<!-- invalid_names <- deseq2VST_long_FLWQLWsubset$variable[!deseq2VST_long_FLWQLWsubset$variable %in% FLWQLW_samplessubset] -->
<!-- if (length(invalid_names) > 0) { -->
<!--   print(paste("Invalid IDs in the DF:", paste(invalid_names, collapse = ", "))) -->
<!-- } else { -->
<!--   print("All IDs in the DF are valid.") -->
<!-- } -->

<!-- ``` -->


<!-- ```{r makeheatmap_FLWQLW, eval=FALSE, fig.height=3} -->
<!-- # Make a heatmap -->
<!-- ggplot(deseq2VST_long_FLWQLWsubset, aes(x=variable, y=Gene, fill=value)) +  -->
<!--   geom_raster() +  -->
<!--   scale_fill_viridis(trans="sqrt") +  -->
<!--   theme(axis.text.x=element_text(angle=65, hjust=1),  -->
<!--         axis.text.y=element_blank(),  -->
<!--         axis.ticks.y=element_blank()) -->
<!-- ``` -->



<!-- ```{r clusteredheatmap_FLWQLW, fig.height=2.5} -->
<!-- # Convert the significant genes back to a atrix for clustering -->
<!-- deseq2VSTMatrix_FLWQLW <- dcast(deseq2VST_long_FLWQLWsubset, Gene ~ variable) -->
<!-- rownames(deseq2VSTMatrix_FLWQLW) <- deseq2VSTMatrix_FLWQLW$Gene -->
<!-- deseq2VSTMatrix_FLWQLW$Gene <- NULL -->

<!-- # Compute a distance calculation on both dimensions of the matrix -->
<!-- #dist() computes the distance based on the rows of the df -->
<!-- #in the current format, gives distance based on genes -->
<!-- distanceGene_FLWQLW <- dist(deseq2VSTMatrix_FLWQLW) -->
<!-- #transpose to get distance based on samples -->
<!-- distanceSample_FLWQLW <- dist(t(deseq2VSTMatrix_FLWQLW)) -->

<!-- # Cluster based on the distance calculations -->
<!-- #hclust: hierarchical cluster -->
<!-- clusterGene_FLWQLW <- hclust(distanceGene_FLWQLW, method="average") -->
<!-- clusterSample_FLWQLW <- hclust(distanceSample_FLWQLW, method="average") -->

<!-- #get specific colors by TX group: -->
<!-- clustorder_FLWQLW <- clusterSample_FLWQLW$labels -->
<!-- color_vector_FLWQLW <- vector("character", length(clustorder_FLWQLW)) -->
<!-- # Loop through each name in clustorder_FLWQLW -->
<!-- for (i in 1:length(clustorder_FLWQLW)) { -->
<!--   name <- clustorder_FLWQLW[i] -->

<!--   # Check if the name exists in rownames of sampleData and if the tx_group is "FLW" -->
<!--   if (name %in% rownames(sampleData) && sampleData[name, "tx_group"] == "FLW") { -->
<!--     color_vector_FLWQLW[i] <- "yellowgreen"  # Assign green if the name matches "FLW" in tx_group -->
<!--   } else { -->
<!--     color_vector_FLWQLW[i] <- "purple1"   # Assign blue if the name does not match "FLW" -->
<!--   } -->
<!-- } -->


<!-- # Construct a dendogram for samples -->
<!-- sampleModel_FLWQLW <- as.dendrogram(clusterSample_FLWQLW) -->
<!-- sampleDendrogramData_FLWQLW <- segment(dendro_data(sampleModel_FLWQLW, type = "rectangle")) -->
<!-- sampleDendrogram_FLWQLW <- ggplot(sampleDendrogramData_FLWQLW) +  -->
<!--   geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) +  -->
<!--   theme_dendro() -->

<!-- # Re-factor samples for ggplot2 -->
<!-- #use to match dendogram -->
<!-- deseq2VST_long_FLWQLWsubset$variable <- factor(deseq2VST_long_FLWQLWsubset$variable,levels=clusterSample_FLWQLW$labels[clusterSample_FLWQLW$order]) -->

<!-- # Construct the heatmap. note that at this point we have only clustered the samples NOT the genes -->
<!-- heatmap_FLWQLW <- ggplot(deseq2VST_long_FLWQLWsubset, aes(x=variable, y=Gene, fill=value)) +  -->
<!--   geom_raster() +  -->
<!--   scale_fill_viridis(trans="sqrt") +  -->
<!--   theme(axis.text.x=element_text(angle=65, hjust=1, color = color_vector_FLWQLW), -->
<!--         #axis.text.y=element_blank(),  -->
<!--         axis.ticks.y=element_blank(), -->
<!--         plot.margin = unit(c(-.75, 0, 0, 0), "cm")) -->

<!-- # Combine the dendrogram and the heatmap -->
<!-- #install.packages("gridExtra") -->
<!-- plot_grid(sampleDendrogram_FLWQLW, NULL, heatmap_FLWQLW, NULL, align= "v", ncol=2, nrow=2, rel_heights = c(.7,1,1,1), rel_widths = c(2.7,.1,2.5,.1)) -->
<!-- ``` -->



### MLW V QLW

```{r extractDEseq_MLWQLW}
# Extract differential expression results
# For "tx_group" perform primary vs normal comparison
res_MLWQLW <- results(deseq2Data, contrast=c("tx_group", "MLW", "QLW"), independentFiltering=FALSE, pAdjustMethod = "BH")

# View summary of results
summary(res_MLWQLW, alpha=0.05)
summary(res_MLWQLW, alpha=0.10)
summary(res_MLWQLW, alpha=0.01)
```


```{r MA_MLWQLW, warning=FALSE}
# Using DEseq2 built in method
plotMA(res_MLWQLW, alpha=0.05)
```


```{r ggplotsMA_MLWQLW, warning=FALSE}
# Coerce to a data frame
res_MLWQLW <- as.data.frame(res_MLWQLW)

# Examine this data frame
head(res_MLWQLW)

# Set a boolean column for significance
res_MLWQLW$significant <- ifelse(res_MLWQLW$padj < adjP, "Significant", NA)

#FOR GSEA Here because don't move on below
write.csv(res_MLWQLW,"DEGFullResults/DEGananalysisResults_MLWQLW.csv", row.names = TRUE)

# Let's add some more detail
ggplot(res_MLWQLW, aes(baseMean, log2FoldChange, colour=padj)) +
geom_point(size=1) +
scale_y_continuous(limits=c(-3, 3), oob=squish) +
scale_x_log10() +
geom_hline(yintercept = 0, colour="darkorchid4", size=1, linetype="longdash") +
labs(x="mean of normalized counts", y="log fold change") +
scale_colour_viridis(direction=-1, trans='sqrt') +
theme_bw() +
  geom_density_2d(colour="black", size=1)
```

No significant genes. Not continuing for this set.

<!-- ####Remove outliers and Subset by TX -->
<!-- ```{r outliers_MLWQLW} -->
<!-- #VERIFY COUNT -->
<!-- #Assess number of genes with adjustedp=NA; N=14 -->
<!-- count(is.na(res_MLWQLW$padj)) -->

<!-- #VERIFY SAME POSITIONS ACROSS ALL PAIRWISE -->
<!-- MLWQLW_napos <- which(is.na(res_MLWQLW$padj), arr.ind = TRUE) -->


<!-- #Create a new data frame to use going forward W/O those padj NAs -->
<!-- deseq2ResDF_MLWQLW.noNA <- res_MLWQLW[-MLWQLW_napos,] -->


<!-- # Keep only the significantly differentiated genes where the fold-change was at least 3 -->
<!-- sigGenes_MLWQLW <- rownames(deseq2ResDF_MLWQLW.noNA[deseq2ResDF_MLWQLW.noNA$padj <= .01 & -->
<!--             abs(deseq2ResDF_MLWQLW.noNA$log2FoldChange) > 3,]) -->
<!-- #Nothing sig at 0.05 padj or 2-fold -->
<!-- #confirmed when looking at the dataset -->


<!-- length(sigGenes_MLWQLW) -->

<!-- deseq2VST_MLWQLW <- deseq2VST[deseq2VST$Gene %in% sigGenes_MLWQLW,] -->


<!-- # First compare wide vs long version -->
<!-- deseq2VST_MLWQLW_wide <- deseq2VST_MLWQLW -->
<!-- deseq2VST_MLWQLW_long <- melt(deseq2VST_MLWQLW, id.vars=c("Gene")) #needed for ggplot -->

<!-- #print(deseq2VST_MLWQLW_wide[1:5,1:7]) -->
<!-- #head(deseq2VST_MLWQLW_long) -->


<!-- ###ONLY FLW AND FLQ SAMPLES -->
<!-- MLWQLW_samplessubset <- rownames(sampleData[which(sampleData$tx_group=="FLW" | sampleData$tx_group=="MLW"),]) -->
<!-- #make empty dataframe for saving in the for loop -->
<!-- deseq2VST_long_MLWQLWsubset <- data.frame() -->

<!-- for (id in MLWQLW_samplessubset) { -->
<!--   subset_id <- deseq2VST_MLWQLW_long[deseq2VST_MLWQLW_long$variable == id, ]  # Subset df by the current value of id -->
<!--   # at to the the subset dataframe each iteration -->
<!--   deseq2VST_long_MLWQLWsubset <- rbind(deseq2VST_long_MLWQLWsubset, subset_id) -->
<!-- } -->

<!-- #verifying status of data frame -->
<!-- # Check if all ids in the variable column of the subsetted list are from the allowed id per the TX group used for the pairwise -->


<!-- # Output the result -->
<!-- invalid_names <- deseq2VST_long_MLWQLWsubset$variable[!deseq2VST_long_MLWQLWsubset$variable %in% MLWQLW_samplessubset] -->
<!-- if (length(invalid_names) > 0) { -->
<!--   print(paste("Invalid IDs in the DF:", paste(invalid_names, collapse = ", "))) -->
<!-- } else { -->
<!--   print("All IDs in the DF are valid.") -->
<!-- } -->

<!-- ``` -->



<!-- ```{r makeheatmap_MLWQLW, eval=FALSE, fig.height=3} -->
<!-- # Make a heatmap -->
<!-- ggplot(deseq2VST_long_MLWQLWsubset, aes(x=variable, y=Gene, fill=value)) +  -->
<!--   geom_raster() +  -->
<!--   scale_fill_viridis(trans="sqrt") +  -->
<!--   theme(axis.text.x=element_text(angle=65, hjust=1),  -->
<!--         axis.text.y=element_blank(),  -->
<!--         axis.ticks.y=element_blank()) -->
<!-- ``` -->


<!-- ```{r clusteredheatmap_MLWQLW, fig.height=2.5} -->
<!-- # Convert the significant genes back to a atrix for clustering -->
<!-- deseq2VSTMatrix_MLWQLW <- dcast(deseq2VST_long_MLWQLWsubset, Gene ~ variable) -->
<!-- rownames(deseq2VSTMatrix_MLWQLW) <- deseq2VSTMatrix_MLWQLW$Gene -->
<!-- deseq2VSTMatrix_MLWQLW$Gene <- NULL -->

<!-- # Compute a distance calculation on both dimensions of the matrix -->
<!-- #dist() computes the distance based on the rows of the df -->
<!-- #in the current format, gives distance based on genes -->
<!-- distanceGene_MLWQLW <- dist(deseq2VSTMatrix_MLWQLW) -->
<!-- #transpose to get distance based on samples -->
<!-- distanceSample_MLWQLW <- dist(t(deseq2VSTMatrix_MLWQLW)) -->

<!-- # Cluster based on the distance calculations -->
<!-- #hclust: hierarchical cluster -->
<!-- clusterGene_MLWQLW <- hclust(distanceGene_MLWQLW, method="average") -->
<!-- clusterSample_MLWQLW <- hclust(distanceSample_MLWQLW, method="average") -->

<!-- #get specific colors by TX group: -->
<!-- clustorder_MLWQLW <- clusterSample_MLWQLW$labels -->
<!-- color_vector_MLWQLW <- vector("character", length(clustorder_MLWQLW)) -->
<!-- # Loop through each name in clustorder_MLWQLW -->
<!-- for (i in 1:length(clustorder_MLWQLW)) { -->
<!--   name <- clustorder_MLWQLW[i] -->

<!--   # Check if the name exists in rownames of sampleData and if the tx_group is "FLW" -->
<!--   if (name %in% rownames(sampleData) && sampleData[name, "tx_group"] == "FLW") { -->
<!--     color_vector_MLWQLW[i] <- "coral"  # Assign green if the name matches "FLW" in tx_group -->
<!--   } else { -->
<!--     color_vector_MLWQLW[i] <- "purple1"   # Assign blue if the name does not match "FLW" -->
<!--   } -->
<!-- } -->


<!-- # Construct a dendogram for samples -->
<!-- sampleModel_MLWQLW <- as.dendrogram(clusterSample_MLWQLW) -->
<!-- sampleDendrogramData_MLWQLW <- segment(dendro_data(sampleModel_MLWQLW, type = "rectangle")) -->
<!-- sampleDendrogram_MLWQLW <- ggplot(sampleDendrogramData_MLWQLW) +  -->
<!--   geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) +  -->
<!--   theme_dendro() -->

<!-- # Re-factor samples for ggplot2 -->
<!-- #use to match dendogram -->
<!-- deseq2VST_long_MLWQLWsubset$variable <- factor(deseq2VST_long_MLWQLWsubset$variable,levels=clusterSample_MLWQLW$labels[clusterSample_MLWQLW$order]) -->

<!-- # Construct the heatmap. note that at this point we have only clustered the samples NOT the genes -->
<!-- heatmap_MLWQLW <- ggplot(deseq2VST_long_MLWQLWsubset, aes(x=variable, y=Gene, fill=value)) +  -->
<!--   geom_raster() +  -->
<!--   scale_fill_viridis(trans="sqrt") +  -->
<!--   theme(axis.text.x=element_text(angle=65, hjust=1, color = color_vector_MLWQLW), -->
<!--         #axis.text.y=element_blank(),  -->
<!--         axis.ticks.y=element_blank(), -->
<!--         plot.margin = unit(c(-.75, 0, 0, 0), "cm")) -->

<!-- # Combine the dendrogram and the heatmap -->
<!-- #install.packages("gridExtra") -->
<!-- plot_grid(sampleDendrogram_MLWQLW, NULL, heatmap_MLWQLW, NULL, align= "v", ncol=2, nrow=2, rel_heights = c(.7,1,1,1), rel_widths = c(2.7,.1,2.5,.1)) -->
<!-- ``` -->



## Within caste--QUEENS
### FLQ V MLQ

```{r extractDEseq_FLQMLQ}
# Extract differential expression results
# For "tx_group" perform primary vs normal comparison
res_FLQMLQ <- results(deseq2Data, contrast=c("tx_group", "FLQ", "MLQ"), independentFiltering=FALSE, pAdjustMethod = "BH")

# View summary of results
summary(res_FLQMLQ, alpha=0.05)
summary(res_FLQMLQ, alpha=0.10)
summary(res_FLQMLQ, alpha=0.01)
```


```{r MA_FLQMLQ, warning=FALSE}
# Using DEseq2 built in method
plotMA(res_FLQMLQ, alpha=0.05)
```


```{r ggplotsMA_FLQMLQ, warning=FALSE}
# Coerce to a data frame
res_FLQMLQ <- as.data.frame(res_FLQMLQ)

# Examine this data frame
head(res_FLQMLQ)

# Set a boolean column for significance
res_FLQMLQ$significant <- ifelse(res_FLQMLQ$padj < adjP, "Significant", NA)

# Let's add some more detail
ggplot(res_FLQMLQ, aes(baseMean, log2FoldChange, colour=padj)) +
geom_point(size=1) +
scale_y_continuous(limits=c(-3, 3), oob=squish) +
scale_x_log10() +
geom_hline(yintercept = 0, colour="darkorchid4", size=1, linetype="longdash") +
labs(x="mean of normalized counts", y="log fold change") +
scale_colour_viridis(direction=-1, trans='sqrt') +
theme_bw() +
  geom_density_2d(colour="black", size=1)
```


#### Remove outliers and Subset by TX

```{r outliers_FLQMLQ}
#VERIFY COUNT
#Assess number of genes with adjustedp=NA; N=14
count(is.na(res_FLQMLQ$padj))

#VERIFY SAME POSITIONS ACROSS ALL PAIRWISE
FLQMLQ_napos <- which(is.na(res_FLQMLQ$padj), arr.ind = TRUE)


#Create a new data frame to use going forward W/O those padj NAs
deseq2ResDF_FLQMLQ.noNA <- res_FLQMLQ[-FLQMLQ_napos,]


# Keep only the significantly differentiated genes where the fold-change was at least 3
sigGenes_FLQMLQ <- rownames(deseq2ResDF_FLQMLQ.noNA[deseq2ResDF_FLQMLQ.noNA$padj <= adjP &
            abs(deseq2ResDF_FLQMLQ.noNA$log2FoldChange) > LFC,])

length(sigGenes_FLQMLQ)

deseq2VST_FLQMLQ <- deseq2VST[deseq2VST$Gene %in% sigGenes_FLQMLQ,]


# First compare wide vs long version
deseq2VST_FLQMLQ_wide <- deseq2VST_FLQMLQ
deseq2VST_FLQMLQ_long <- melt(deseq2VST_FLQMLQ, id.vars=c("Gene")) #needed for ggplot

print(deseq2VST_FLQMLQ_wide[1:5,1:7])
head(deseq2VST_FLQMLQ_long)


###ONLY FLQ AND FLQ SAMPLES
FLQMLQ_samplessubset <- rownames(sampleData[which(sampleData$tx_group=="FLQ" | sampleData$tx_group=="MLQ"),])
#make empty dataframe for saving in the for loop
deseq2VST_long_FLQMLQsubset <- data.frame()

for (id in FLQMLQ_samplessubset) {
  subset_id <- deseq2VST_FLQMLQ_long[deseq2VST_FLQMLQ_long$variable == id, ]  # Subset df by the current value of id
  # at to the the subset dataframe each iteration
  deseq2VST_long_FLQMLQsubset <- rbind(deseq2VST_long_FLQMLQsubset, subset_id)
}

#verifying status of data frame
# Check if all ids in the variable column of the subsetted list are from the allowed id per the TX group used for the pairwise


# Output the result
invalid_names <- deseq2VST_long_FLQMLQsubset$variable[!deseq2VST_long_FLQMLQsubset$variable %in% FLQMLQ_samplessubset]
if (length(invalid_names) > 0) {
  print(paste("Invalid IDs in the DF:", paste(invalid_names, collapse = ", ")))
} else {
  print("All IDs in the DF are valid.")
}

```

```{r makeheatmap_FLQMLQ, eval=FALSE, fig.height=3, echo=FALSE}
# Make a heatmap
ggplot(deseq2VST_long_FLQMLQsubset, aes(x=variable, y=Gene, fill=value)) +
  geom_raster() +
  scale_fill_viridis(trans="sqrt") +
  theme(axis.text.x=element_text(angle=65, hjust=1),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r clusteredheatmap_FLQMLQ, fig.height=10, echo=FALSE}
# Convert the significant genes back to a atrix for clustering
deseq2VSTMatrix_FLQMLQ <- dcast(deseq2VST_long_FLQMLQsubset, Gene ~ variable)
rownames(deseq2VSTMatrix_FLQMLQ) <- deseq2VSTMatrix_FLQMLQ$Gene
deseq2VSTMatrix_FLQMLQ$Gene <- NULL

# Compute a distance calculation on both dimensions of the matrix
#dist() computes the distance based on the rows of the df
#in the current format, gives distance based on genes
distanceGene_FLQMLQ <- dist(deseq2VSTMatrix_FLQMLQ)
#transpose to get distance based on samples
distanceSample_FLQMLQ <- dist(t(deseq2VSTMatrix_FLQMLQ))

# Cluster based on the distance calculations
#hclust: hierarchical cluster
clusterGene_FLQMLQ <- hclust(distanceGene_FLQMLQ, method="average")
clusterSample_FLQMLQ <- hclust(distanceSample_FLQMLQ, method="average")

#get specific colors by TX group:
clustorder_FLQMLQ <- clusterSample_FLQMLQ$order
color_vector_FLQMLQ <- vector("character", length(clustorder_FLQMLQ))
# Loop through each name in clustorder_FLQMLQ
for (i in 1:length(clustorder_FLQMLQ)) {
  name <- clustorder_FLQMLQ[i]

  # Check if the name exists in rownames of sampleData and if the tx_group is "FLW"
  #GREATER THAN 7 BECAUSE FLQ COMES BEFORE MLQ IN SEQ INPUT
  if (name >= 7) {
    color_vector_FLQMLQ[i] <- "yellowgreen"  # Assign green if the name matches "FLW" in tx_group
  } else {
    color_vector_FLQMLQ[i] <- "coral"   # Assign blue if the name does not match "FLW"
  }
}



color_vector_FLQMLQ_xcol <- c("yellowgreen","coral","coral","coral","coral","yellowgreen","coral","coral","yellowgreen","yellowgreen","yellowgreen")

color_vector_FLQMLQ == color_vector_FLQMLQ_xcol


# Construct a dendogram for samples
sampleModel_FLQMLQ <- as.dendrogram(clusterSample_FLQMLQ)
sampleDendrogramData_FLQMLQ <- segment(dendro_data(sampleModel_FLQMLQ, type = "rectangle"))
sampleDendrogram_FLQMLQ <- ggplot(sampleDendrogramData_FLQMLQ) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) +
  theme_dendro()

# Re-factor samples for ggplot2
#use to match dendogram
deseq2VST_long_FLQMLQsubset$variable <- factor(deseq2VST_long_FLQMLQsubset$variable,levels=clusterSample_FLQMLQ$labels[clusterSample_FLQMLQ$order])

# Construct the heatmap. note that at this point we have only clustered the samples NOT the genes
heatmap_FLQMLQ <- ggplot(deseq2VST_long_FLQMLQsubset, aes(x=variable, y=Gene, fill=value)) +
  geom_raster() +
  scale_fill_viridis(trans="sqrt") +
  theme(axis.text.x=element_text(angle=65, hjust=1, color = color_vector_FLQMLQ),
        #axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        plot.margin = unit(c(-.75, 0, 0, 0), "cm"))

# Combine the dendrogram and the heatma
titleFLQMLQ <- ggdraw() + draw_label("Significant Genes for FL queen (yellow-green) v ML queen (coral)", fontface='bold')
plot_grid(titleFLQMLQ, NULL, sampleDendrogram_FLQMLQ, NULL, heatmap_FLQMLQ, NULL, align= "v", ncol=2, nrow=3, rel_heights = c(.05,.2,1,1,1,1), rel_widths = c(2.7,.1,2.7,.1,2.5,.1))
```


```{r regdirection_FLQMLQ, fig.height=7, echo=FALSE}
DEGs_FLQMLQ <- deseq2ResDF_FLQMLQ.noNA[rownames(deseq2ResDF_FLQMLQ.noNA) %in% sigGenes_FLQMLQ,]

#write the DEGs and data to CSV
write.csv(DEGs_FLQMLQ,"SigGeneLists/DEGs_FLQMLQ.csv", row.names = TRUE)
DEG_genelist_FLQMLQ <- rownames(DEGs_FLQMLQ)
#write gene list to txt
write.table(DEG_genelist_FLQMLQ,"SigGeneLists/DEG_genelist_FLQMLQ.txt", row.names = FALSE, col.names = FALSE, quote=FALSE)


DEGs.FCdir_FLQMLQ <- DEGs_FLQMLQ[,c("log2FoldChange", "padj")]
length(DEGs.FCdir_FLQMLQ$log2FoldChange)


DEGs.FCdir_FLQMLQ

#FOR GSEA
write.csv(res_FLQMLQ,"DEGFullResults/DEGananalysisResults_FLQMLQ.csv", row.names = TRUE)
```



### FLQ V QLQ

```{r extractDEseq_FLQQLQ}
# Extract differential expression results
# For "tx_group" perform primary vs normal comparison
res_FLQQLQ <- results(deseq2Data, contrast=c("tx_group", "FLQ", "QLQ"), independentFiltering=FALSE, pAdjustMethod = "BH")

# View summary of results
summary(res_FLQQLQ, alpha=0.05)
summary(res_FLQQLQ, alpha=0.10)
summary(res_FLQQLQ, alpha=0.01)
```


```{r MA_FLQQLQ, warning=FALSE}
# Using DEseq2 built in method
plotMA(res_FLQQLQ, alpha=0.05)
```


```{r ggplotsMA_FLQQLQ, warning=FALSE}
# Coerce to a data frame
res_FLQQLQ <- as.data.frame(res_FLQQLQ)

# Examine this data frame
head(res_FLQQLQ)

# Set a boolean column for significance
res_FLQQLQ$significant <- ifelse(res_FLQQLQ$padj < adjP, "Significant", NA)

# Let's add some more detail
ggplot(res_FLQQLQ, aes(baseMean, log2FoldChange, colour=padj)) +
geom_point(size=1) +
scale_y_continuous(limits=c(-3, 3), oob=squish) +
scale_x_log10() +
geom_hline(yintercept = 0, colour="darkorchid4", size=1, linetype="longdash") +
labs(x="mean of normalized counts", y="log fold change") +
scale_colour_viridis(direction=-1, trans='sqrt') +
theme_bw() +
  geom_density_2d(colour="black", size=1)
```


#### Remove outliers and Subset by TX

```{r outliers_FLQQLQ}
#VERIFY COUNT
#Assess number of genes with adjustedp=NA; N=14
count(is.na(res_FLQQLQ$padj))

#VERIFY SAME POSITIONS ACROSS ALL PAIRWISE
FLQQLQ_napos <- which(is.na(res_FLQQLQ$padj), arr.ind = TRUE)


#Create a new data frame to use going forward W/O those padj NAs
deseq2ResDF_FLQQLQ.noNA <- res_FLQQLQ[-FLQQLQ_napos,]


# Keep only the significantly differentiated genes where the fold-change was at least 3
sigGenes_FLQQLQ <- rownames(deseq2ResDF_FLQQLQ.noNA[deseq2ResDF_FLQQLQ.noNA$padj <= adjP &
            abs(deseq2ResDF_FLQQLQ.noNA$log2FoldChange) > LFC,])

length(sigGenes_FLQQLQ)

deseq2VST_FLQQLQ <- deseq2VST[deseq2VST$Gene %in% sigGenes_FLQQLQ,]


# First compare wide vs long version
deseq2VST_FLQQLQ_wide <- deseq2VST_FLQQLQ
deseq2VST_FLQQLQ_long <- melt(deseq2VST_FLQQLQ, id.vars=c("Gene")) #needed for ggplot

print(deseq2VST_FLQQLQ_wide[1:3,1:7])
head(deseq2VST_FLQQLQ_long)


###ONLY FLQ AND FLQ SAMPLES
FLQQLQ_samplessubset <- rownames(sampleData[which(sampleData$tx_group=="FLQ" | sampleData$tx_group=="QLQ"),])
#make empty dataframe for saving in the for loop
deseq2VST_long_FLQQLQsubset <- data.frame()

for (id in FLQQLQ_samplessubset) {
  subset_id <- deseq2VST_FLQQLQ_long[deseq2VST_FLQQLQ_long$variable == id, ]  # Subset df by the current value of id
  # at to the the subset dataframe each iteration
  deseq2VST_long_FLQQLQsubset <- rbind(deseq2VST_long_FLQQLQsubset, subset_id)
}

#verifying status of data frame
# Check if all ids in the variable column of the subsetted list are from the allowed id per the TX group used for the pairwise


# Output the result
invalid_names <- deseq2VST_long_FLQQLQsubset$variable[!deseq2VST_long_FLQQLQsubset$variable %in% FLQQLQ_samplessubset]
if (length(invalid_names) > 0) {
  print(paste("Invalid IDs in the DF:", paste(invalid_names, collapse = ", ")))
} else {
  print("All IDs in the DF are valid.")
}

```

```{r makeheatmap_FLQQLQ, eval=FALSE, fig.height=3, echo=FALSE}
# Make a heatmap
ggplot(deseq2VST_long_FLQQLQsubset, aes(x=variable, y=Gene, fill=value)) +
  geom_raster() +
  scale_fill_viridis(trans="sqrt") +
  theme(axis.text.x=element_text(angle=65, hjust=1),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r clusteredheatmap_FLQQLQ, fig.height=6, echo=FALSE}
# Convert the significant genes back to a atrix for clustering
deseq2VSTMatrix_FLQQLQ <- dcast(deseq2VST_long_FLQQLQsubset, Gene ~ variable)
rownames(deseq2VSTMatrix_FLQQLQ) <- deseq2VSTMatrix_FLQQLQ$Gene
deseq2VSTMatrix_FLQQLQ$Gene <- NULL

# Compute a distance calculation on both dimensions of the matrix
#dist() computes the distance based on the rows of the df
#in the current format, gives distance based on genes
distanceGene_FLQQLQ <- dist(deseq2VSTMatrix_FLQQLQ)
#transpose to get distance based on samples
distanceSample_FLQQLQ <- dist(t(deseq2VSTMatrix_FLQQLQ))

# Cluster based on the distance calculations
#hclust: hierarchical cluster
clusterGene_FLQQLQ <- hclust(distanceGene_FLQQLQ, method="average")
clusterSample_FLQQLQ <- hclust(distanceSample_FLQQLQ, method="average")

#get specific colors by TX group:
clustorder_FLQQLQ <- clusterSample_FLQQLQ$order
color_vector_FLQQLQ <- vector("character", length(clustorder_FLQQLQ))
# Loop through each name in clustorder_FLQQLQ
for (i in 1:length(clustorder_FLQQLQ)) {
  name <- clustorder_FLQQLQ[i]

  # Check if the name exists in rownames of sampleData and if the tx_group is "FLW"
  #GREATER THAN 7 BECAUSE FLQ COMES BEFORE MLQ IN SEQ INPUT
  if (name >= 6) {
    color_vector_FLQQLQ[i] <- "yellowgreen"  # Assign green if the name matches "FLW" in tx_group
  } else {
    color_vector_FLQQLQ[i] <- "purple1"   # Assign blue if the name does not match "FLW"
  }
}



color_vector_FLQQLQ_xcol <- c("yellowgreen","yellowgreen","yellowgreen","yellowgreen","yellowgreen","purple1","purple1","purple1","purple1","purple1")

color_vector_FLQQLQ == color_vector_FLQQLQ_xcol

# Construct a dendogram for samples
sampleModel_FLQQLQ <- as.dendrogram(clusterSample_FLQQLQ)
sampleDendrogramData_FLQQLQ <- segment(dendro_data(sampleModel_FLQQLQ, type = "rectangle"))
sampleDendrogram_FLQQLQ <- ggplot(sampleDendrogramData_FLQQLQ) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) +
  theme_dendro()

# Re-factor samples for ggplot2
#use to match dendogram
deseq2VST_long_FLQQLQsubset$variable <- factor(deseq2VST_long_FLQQLQsubset$variable,levels=clusterSample_FLQQLQ$labels[clusterSample_FLQQLQ$order])

# Construct the heatmap. note that at this point we have only clustered the samples NOT the genes
heatmap_FLQQLQ <- ggplot(deseq2VST_long_FLQQLQsubset, aes(x=variable, y=Gene, fill=value)) +
  geom_raster() +
  scale_fill_viridis(trans="sqrt") +
  theme(axis.text.x=element_text(angle=65, hjust=1, color = color_vector_FLQQLQ_xcol),
        #axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        plot.margin = unit(c(-.75, 0, 0, 0), "cm"))

# Combine the dendrogram and the heatmap
#install.packages("gridExtra")

titleFLQQLQ <- ggdraw() + draw_label("Significant Genes for FL queen (yellow-green) v QL queen (purple)", fontface='bold')
plot_grid(titleFLQQLQ, NULL, sampleDendrogram_FLQQLQ, NULL, heatmap_FLQQLQ, NULL, align= "v", ncol=2, nrow=3, rel_heights = c(.1,.4,1,1,1,1), rel_widths = c(2.7,.1,2.7,.1,2.5,.1))

```


```{r regdirection_FLQQLQ, fig.height=7, echo=FALSE}
DEGs_FLQQLQ <- deseq2ResDF_FLQQLQ.noNA[rownames(deseq2ResDF_FLQQLQ.noNA) %in% sigGenes_FLQQLQ,]


#write the DEGs and data to CSV
write.csv(DEGs_FLQQLQ,"SigGeneLists/DEGs_FLQQLQ.csv", row.names = TRUE)
DEG_genelist_FLQQLQ <- rownames(DEGs_FLQQLQ)
#write gene list to txt
write.table(DEG_genelist_FLQQLQ,"SigGeneLists/DEG_genelist_FLQQLQ.txt", row.names = FALSE, col.names = FALSE, quote=FALSE)


DEGs.FCdir_FLQQLQ <- DEGs_FLQQLQ[,c("log2FoldChange", "padj")]
length(DEGs.FCdir_FLQQLQ$log2FoldChange)


DEGs.FCdir_FLQQLQ


#FOR GSEA
write.csv(res_FLQQLQ,"DEGFullResults/DEGananalysisResults_FLQQLQ.csv", row.names = TRUE)
```


### MLQ V QLQ

```{r extractDEseq_MLQQLQ}
# Extract differential expression results
# For "tx_group" perform primary vs normal comparison
res_MLQQLQ <- results(deseq2Data, contrast=c("tx_group", "MLQ", "QLQ"), independentFiltering=FALSE, pAdjustMethod = "BH")

# View summary of results
summary(res_MLQQLQ, alpha=0.05)
summary(res_MLQQLQ, alpha=0.10)
summary(res_MLQQLQ, alpha=0.01)
```


```{r MA_MLQQLQ, warning=FALSE, fig.height=5.5}
# Using DEseq2 built in method
plotMA(res_MLQQLQ, alpha=0.05)
```


```{r ggplotsMA_MLQQLQ, warning=FALSE}
# Coerce to a data frame
res_MLQQLQ <- as.data.frame(res_MLQQLQ)

# Examine this data frame
head(res_MLQQLQ)

# Set a boolean column for significance
res_MLQQLQ$significant <- ifelse(res_MLQQLQ$padj < adjP, "Significant", NA)

#FOR GSEA HEre because they don't have genes that are sign at the thresholds
write.csv(res_MLQQLQ,"DEGFullResults/DEGananalysisResults_MLQQLQ.csv", row.names = TRUE)



# Let's add some more detail
ggplot(res_MLQQLQ, aes(baseMean, log2FoldChange, colour=padj)) +
geom_point(size=1) +
scale_y_continuous(limits=c(-3, 3), oob=squish) +
scale_x_log10() +
geom_hline(yintercept = 0, colour="darkorchid4", size=1, linetype="longdash") +
labs(x="mean of normalized counts", y="log fold change") +
scale_colour_viridis(direction=-1, trans='sqrt') +
theme_bw() +
  geom_density_2d(colour="black", size=1)
```


####Remove outliers and Subset by TX

```{r outliers_MLQQLQ}
#VERIFY COUNT
#Assess number of genes with adjustedp=NA; N=14
count(is.na(res_MLQQLQ$padj))

#VERIFY SAME POSITIONS ACROSS ALL PAIRWISE
MLQQLQ_napos <- which(is.na(res_MLQQLQ$padj), arr.ind = TRUE)


#Create a new data frame to use going forward W/O those padj NAs
deseq2ResDF_MLQQLQ.noNA <- res_MLQQLQ[-MLQQLQ_napos,]


# Keep only the significantly differentiated genes where the fold-change was at least 3
sigGenes_MLQQLQ <- rownames(deseq2ResDF_MLQQLQ.noNA[deseq2ResDF_MLQQLQ.noNA$padj <= adjP &
            abs(deseq2ResDF_MLQQLQ.noNA$log2FoldChange) > LFC,])

length(sigGenes_MLQQLQ)

deseq2VST_MLQQLQ <- deseq2VST[deseq2VST$Gene %in% sigGenes_MLQQLQ,]


# First compare wide vs long version
deseq2VST_MLQQLQ_wide <- deseq2VST_MLQQLQ
deseq2VST_MLQQLQ_long <- melt(deseq2VST_MLQQLQ, id.vars=c("Gene")) #needed for ggplot

#print(deseq2VST_MLQQLQ_wide[1:5,1:7])
#head(deseq2VST_MLQQLQ_long)


###ONLY FLQ AND FLQ SAMPLES
MLQQLQ_samplessubset <- rownames(sampleData[which(sampleData$tx_group=="MLQ" | sampleData$tx_group=="QLQ"),])
#make empty dataframe for saving in the for loop
deseq2VST_long_MLQQLQsubset <- data.frame()

for (id in MLQQLQ_samplessubset) {
  subset_id <- deseq2VST_MLQQLQ_long[deseq2VST_MLQQLQ_long$variable == id, ]  # Subset df by the current value of id
  # at to the the subset dataframe each iteration
  deseq2VST_long_MLQQLQsubset <- rbind(deseq2VST_long_MLQQLQsubset, subset_id)
}

#verifying status of data frame
# Check if all ids in the variable column of the subsetted list are from the allowed id per the TX group used for the pairwise


# Output the result
invalid_names <- deseq2VST_long_MLQQLQsubset$variable[!deseq2VST_long_MLQQLQsubset$variable %in% MLQQLQ_samplessubset]
if (length(invalid_names) > 0) {
  print(paste("Invalid IDs in the DF:", paste(invalid_names, collapse = ", ")))
} else {
  print("All IDs in the DF are valid.")
}

```

The significant gene does not reach the adjusted p-value of 0.01 and 3-fold change criteria. Not producing heatmap. For reference, the fold change was ~ -0.5 and adj p value was just below 0.05.

<!-- ```{r makeheatmap_MLQQLQ, eval=FALSE, fig.height=3} -->
<!-- # Make a heatmap -->
<!-- ggplot(deseq2VST_long_MLQQLQsubset, aes(x=variable, y=Gene, fill=value)) +  -->
<!--   geom_raster() +  -->
<!--   scale_fill_viridis(trans="sqrt") +  -->
<!--   theme(axis.text.x=element_text(angle=65, hjust=1),  -->
<!--         axis.text.y=element_blank(),  -->
<!--         axis.ticks.y=element_blank()) -->
<!-- ``` -->

<!-- ```{r clusteredheatmap_MLQQLQ, fig.height=3.5} -->
<!-- # Convert the significant genes back to a atrix for clustering -->
<!-- deseq2VSTMatrix_MLQQLQ <- dcast(deseq2VST_long_MLQQLQsubset, Gene ~ variable) -->
<!-- rownames(deseq2VSTMatrix_MLQQLQ) <- deseq2VSTMatrix_MLQQLQ$Gene -->
<!-- deseq2VSTMatrix_MLQQLQ$Gene <- NULL -->

<!-- # Compute a distance calculation on both dimensions of the matrix -->
<!-- #dist() computes the distance based on the rows of the df -->
<!-- #in the current format, gives distance based on genes -->
<!-- distanceGene_MLQQLQ <- dist(deseq2VSTMatrix_MLQQLQ) -->
<!-- #transpose to get distance based on samples -->
<!-- distanceSample_MLQQLQ <- dist(t(deseq2VSTMatrix_MLQQLQ)) -->

<!-- # Cluster based on the distance calculations -->
<!-- #hclust: hierarchical cluster -->
<!-- clusterGene_MLQQLQ <- hclust(distanceGene_MLQQLQ, method="average") -->
<!-- clusterSample_MLQQLQ <- hclust(distanceSample_MLQQLQ, method="average") -->

<!-- #get specific colors by TX group: -->
<!-- clustorder_MLQQLQ <- clusterSample_MLQQLQ$order -->
<!-- color_vector_MLQQLQ <- vector("character", length(clustorder_MLQQLQ)) -->
<!-- # Loop through each name in clustorder_MLQQLQ -->
<!-- for (i in 1:length(clustorder_MLQQLQ)) { -->
<!--   name <- clustorder_MLQQLQ[i] -->

<!--   # Check if the name exists in rownames of sampleData and if the tx_group is "FLW" -->
<!--   #GREATER THAN 7 BECAUSE FLQ COMES BEFORE MLQ IN SEQ INPUT -->
<!--   if (name >= 7) { -->
<!--     color_vector_MLQQLQ[i] <- "yellowgreen"  # Assign green if the name matches "FLW" in tx_group -->
<!--   } else { -->
<!--     color_vector_MLQQLQ[i] <- "purple1"   # Assign blue if the name does not match "FLW" -->
<!--   } -->
<!-- } -->



<!-- color_vector_MLQQLQ_xcol <- c("yellowgreen","yellowgreen","yellowgreen","yellowgreen","purple1","purple1","yellowgreen","purple1","yellowgreen","purple1","purple1","purple1") -->

<!-- color_vector_MLQQLQ == color_vector_MLQQLQ_xcol -->

<!-- # Construct a dendogram for samples -->
<!-- sampleModel_MLQQLQ <- as.dendrogram(clusterSample_MLQQLQ) -->
<!-- sampleDendrogramData_MLQQLQ <- segment(dendro_data(sampleModel_MLQQLQ, type = "rectangle")) -->
<!-- sampleDendrogram_MLQQLQ <- ggplot(sampleDendrogramData_MLQQLQ) +  -->
<!--   geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) +  -->
<!--   theme_dendro() -->

<!-- # Re-factor samples for ggplot2 -->
<!-- #use to match dendogram -->
<!-- deseq2VST_long_MLQQLQsubset$variable <- factor(deseq2VST_long_MLQQLQsubset$variable,levels=clusterSample_MLQQLQ$labels[clusterSample_MLQQLQ$order]) -->

<!-- # Construct the heatmap. note that at this point we have only clustered the samples NOT the genes -->
<!-- heatmap_MLQQLQ <- ggplot(deseq2VST_long_MLQQLQsubset, aes(x=variable, y=Gene, fill=value)) +  -->
<!--   geom_raster() +  -->
<!--   scale_fill_viridis(trans="sqrt") +  -->
<!--   theme(axis.text.x=element_text(angle=65, hjust=1, color = color_vector_MLQQLQ_xcol), -->
<!--         #axis.text.y=element_blank(),  -->
<!--         axis.ticks.y=element_blank(), -->
<!--         plot.margin = unit(c(-.75, 0, 0, 0), "cm")) -->

<!-- # Combine the dendrogram and the heatmap -->
<!-- #install.packages("gridExtra") -->
<!-- plot_grid(sampleDendrogram_MLQQLQ, NULL, heatmap_MLQQLQ, NULL, align= "v", ncol=2, nrow=2, rel_heights = c(.3,1,1,1), rel_widths = c(2.7,.1,2.5,.1)) -->
<!-- ``` -->



# Get a Unique list of Differentially Expressed Genes
```{r UniqueDEGlist}
SigDEGs <- c(DEG_genelist_FLQMLQ, DEG_genelist_FLQQLQ, DEG_genelist_FLWFLQ,  DEG_genelist_FLWMLW, DEG_genelist_MLWMLQ, DEG_genelist_QLWQLQ)

SigDEGsUnique <- unique(SigDEGs)
write.table(SigDEGsUnique,"DEGGeneList/SigDEGsAcrossAllPWC.txt", row.names = FALSE, col.names = FALSE, quote=FALSE)

# DEGs <- c(DEG_genelist_FLQMLQ, DEG_genelist_FLQQLQ, DEG_genelist_FLWFLQ,  DEG_genelist_FLWMLW, DEG_genelist_MLWMLQ, DEG_genelist_QLWQLQ)
# 
```




# Save Data For Import Later
```{r SAVE}
save(DEG_genelist_FLQMLQ, DEG_genelist_FLQQLQ, DEG_genelist_FLWFLQ,  DEG_genelist_FLWMLW, DEG_genelist_MLWMLQ, DEG_genelist_QLWQLQ, file="BIMB_MB_DEGdataGeneLists.RData")

save(deseq2ResDF_FLWFLQ.noNA, deseq2ResDF_MLWMLQ.noNA, deseq2ResDF_QLWQLQ.noNA, deseq2ResDF_FLWMLW.noNA, deseq2ResDF_FLQMLQ.noNA, deseq2ResDF_FLQQLQ.noNA, deseq2ResDF_MLQQLQ.noNA, file="BIMB_MB_DESeqDFsFull.RData")

save(DEGs.FCdir_FLQMLQ, DEGs.FCdir_FLQQLQ, DEGs.FCdir_FLWFLQ, DEGs.FCdir_FLWMLW, DEGs.FCdir_MLWMLQ, DEGs.FCdir_QLWQLQ, file="BIMB_MB_DEGs_pLFC.RData")

save(SigDEGs, SigDEGsUnique, file="BIMB_MB_SigDEGs.RData")

```