---
title: "B. impatiens Maternal Brain Transcriptomics: gene filtering pre-analyses"
author: "M.A. Hagadorn"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = TRUE,
	highlight = TRUE,
	tidy.opts = list(width.cutoff = 80), tidy = TRUE)
```



```{r loadpackages, warning=FALSE, echo=FALSE, message=FALSE, include=FALSE}
#Library Necessary Packages
library(edgeR) 
library(tidyverse)
library(ggfortify)
library(dplyr)
library(tidyr)
library(ggplot2)
options(stringsAsFactors = FALSE)
```
# Rationale and Load Data
After speaking with KMK on 1/17/2025 I am adding a filtering step prior to all analyses. We are doing so using edgeR built in function *filterByExpr*. I used the following resources and conversations to do so: 1) https://bioconductor.org/packages/release/workflows/vignettes/RnaSeqGeneEdgeRQL/inst/doc/edgeRQL.html, 2) https://support.bioconductor.org/p/107032/ , and 3) https://support.bioconductor.org/p/9157228/.

The first step is to import the data and get a TX matrix set up for filtering.
```{r loadinthedata}
#load in the same count matrix used for WGCNA
GEcounttable <- read.csv2("../../../RNASeqAnalysis/Bimp_MB_GeneCountData.csv", sep=";")
head(GEcounttable)
dim(GEcounttable)

#Save OG copy for later
GEcounttable0 <- GEcounttable

#Read in with samples as column names and gene as column 1


#load in the sample mappings
sampleData <- read.delim("../../../RNASeqSamples_plusexperimentalandmoleculardata.csv", sep=",")
head(sampleData)
# Also save a copy for later
sampleData0 <- sampleData

#subset the data for only some of the columns
sampleData <- sampleData[,c(1:2,3:6,8,19)]
head(sampleData)
```


```{r makedesignmatrix}
#assign as factors
#below seems to assign FLW as the intercept with 1s for all. This is not right
#corrected by adding ~0+ to the model
sampleData$tx_group <- factor(sampleData$tx_group, levels=c("FLW", "FLQ", "MLW", "MLQ", "QLW", "QLQ"))

#create the design matrix based on that model
design <- model.matrix(~ 0+sampleData$tx_group)

#assign the col colnames
colnames(design) <- levels(sampleData$tx_group)

#what the matrix looks like
design
```

# Filtering
When appling the *filterByExpr* function, the default parameters are based on count-per-million (CPM). The rule of thumb recommended by the author is "a count of at least 10 in at least some of the libraries before it is considered to be expressed in the study".  
```{r filterbyexpression}
#using the design matrix, we filter by expression

#prefilter info
dim(GEcounttable)

#reformat for genes to be rownames
rownames(GEcounttable) <- GEcounttable$X
GEcounttable <- GEcounttable[,-1] #remove gene names column

#filterbyexpresion
keepergenes <- filterByExpr(GEcounttable, design)
table(keepergenes)

#new filtered table
GEcounttable_Filtered <- GEcounttable[keepergenes,]
```


Advise from author Gordon Smyth recommends using default settings in all circumstances: https://support.bioconductor.org/p/9157228/. The one case where it isn't Smyth argues that that was inappropriate.
```{r saveFilteredGenes}
write.table(GEcounttable_Filtered, file = "../../Bimp_MB_GeneCountData_FilteredGenes.txt", sep = "")
write.csv2(GEcounttable_Filtered, file = "../../Bimp_MB_GeneCountData_FilteredGenes.csv")
```


# PCA: explore the dimensionality of the filtered data
Used the following resource as a guide: https://tavareshugo.github.io/data-carpentry-rnaseq/03_rnaseq_pca.html
```{r PCASetup}
# Create a matrix from our table of counts
#convert to matrix and then transpose
pca_matrix <- t(as.matrix(GEcounttable_Filtered))
#Look at the first five genes and samples
pca_matrix[1:5, 1:5]

pca_matrix0 <- pca_matrix

# Convert matrix to tibble for future use
as_tibble(pca_matrix, rownames = "sample")


# Perform the PCA
pca <- prcomp(pca_matrix)
pc_eigenvalues <- pca$sdev^2


#to view using ggplot2 create a tibble
# create a "tibble" manually with 
# a variable indicating the PC number
# and a variable with the variances
pc_eigenvalues <- tibble(PC = factor(1:length(pc_eigenvalues)), 
                         variance = pc_eigenvalues) %>% 
  # add a new column with the percent variance
  mutate(pct = variance/sum(variance)*100) %>% 
  # add another column with the cumulative variance explained
  mutate(pct_cum = cumsum(pct))

# print the result
pc_eigenvalues
```

To assess the fraction of total variance explained by each PC, we are making a Scree Plot.

```{r PCAScree}
pc_eigenvalues %>% 
  ggplot(aes(x = PC)) +
  geom_col(aes(y = pct)) +
  geom_line(aes(y = pct_cum, group = 1)) + 
  geom_point(aes(y = pct_cum)) +
  labs(x = "Principal component", y = "Fraction variance explained")
```


To assess the fraction of total variance explained by each PC, we are making a Scree Plot.

```{r PCAscoresplot}
pc_scores <- pca$x

pc_scores <- pc_scores %>% 
  # convert to a tibble retaining the sample names as a new column
  as_tibble(rownames = "sample")

# print the result
pc_scores




#PLOT
PC_plot1 <- 
  ggplot(aes(x = PC1, y = PC2), data=pc_scores) +
  geom_point()

PC_plot1

PC_plot1 + geom_label(aes(label = sample))


#Cleaner version
pc_scores[1,1:3]

PC_plot1 <- 
  ggplot(aes(x = PC1, y = PC2), data=pc_scores) +
  geom_point()+
  annotate("text", x = -330377.2 + 20000, y = 10603.47 + 5000, label = "K23C051", color = "red", size = 4) 

PC_plot1
```

Sample **K23C051** falls well away from all other samples. This is the same gene considered an outlier in the WGCNA analysis. Per my notes, there are potentially a few issues with this sample. First, this is the only sample that we have from colony K23C, meaning it could be differences at the colony-level. Second, while preparing this sample for dissection, notes indicate that *"tough to get queens head out of LN tube"*. Thus, it may be a RNA degradation issue, though the RINe and qubit values were among the highest.

Since this looks to  be the driver of PC1, I am going to remove this sample and then rerun the PCA to see what we can find.


## Removed: *K23C051*
```{r PCAwithoutCsample}
#Remove K23C051
samples <- rownames(pca_matrix0)
pca_matrix_no51 <- pca_matrix0[!samples %in% c("K23C051"),]
dim(pca_matrix_no51)



# Perform the PCA
pca_no51 <- prcomp(pca_matrix_no51)
pc_eigenvalues_no51 <- pca_no51$sdev^2


#to view using ggplot2 create a tibble
# create a "tibble" manually with 
# a variable indicating the PC number
# and a variable with the variances
pc_eigenvalues_no51 <- tibble(PC = factor(1:length(pc_eigenvalues_no51)), 
                         variance = pc_eigenvalues_no51) %>% 
  # add a new column with the percent variance
  mutate(pct = variance/sum(variance)*100) %>% 
  # add another column with the cumulative variance explained
  mutate(pct_cum = cumsum(pct))

# print the result
pc_eigenvalues_no51



####SCREE
pc_eigenvalues_no51 %>% 
  ggplot(aes(x = PC)) +
  geom_col(aes(y = pct)) +
  geom_line(aes(y = pct_cum, group = 1)) + 
  geom_point(aes(y = pct_cum)) +
  labs(x = "Principal component", y = "Fraction variance explained")


##PCA plot##
pc_scores_no51 <- pca_no51$x


pc_scores_no51 <- pc_scores_no51 %>% 
  # convert to a tibble retaining the sample names as a new column
  as_tibble(rownames = "sample")

# print the result
pc_scores_no51

#PLOT
PC_plot2 <- 
  ggplot(aes(x = PC1, y = PC2), data=pc_scores_no51) +
  geom_point() +
  annotate("text", x = 122659 - 8000, y = -68010 + 5000, label = "K23D079", color = "red")

PC_plot2

#PC_plot2 + geom_label(aes(label = sample))
```




*K23D079* is labeled because this sample was also considered an outlier in the WGCNA analysis. I've kept it here for now. The information I have on this samples is: *"took a while to get head out of tube"*. Thus, again, it could be an RNA degradation issue, though the RINe and qubit values were among the highest of all samples.

I proceeded to looking at PCAs with various factors highlighted.
\newpage

```{r plotspice, fig.width= 8, fig.height=4.75, echo=FALSE}
##lets add some details
sampleData_no51 <- sampleData[-9,]


##tx_category x caste
autoplot(pca_no51, data = sampleData_no51, colour = "tx_category", shape = "caste", size=3.5) +
   scale_shape_manual(values=c(8,19)) +
   scale_colour_manual(values=c("forestgreen","coral","purple1"))+
   scale_fill_manual(values=c("forestgreen","coral","purple1")) +
   annotate("text", x = 0.59, y = -.34, label = "K23D079", color = "black") +
   ggtitle("Category x Caste") +
    theme(plot.title = element_text(hjust = 0.5))

##caste by colony
autoplot(pca_no51, data = sampleData_no51, colour = "colony", shape = "caste", size=3.5) +
   scale_shape_manual(values=c(8,19)) +
   annotate("text", x = 0.59, y = -.34, label = "K23D079", color = "black") +
   ggtitle("Colony x Caste") +
    theme(plot.title = element_text(hjust = 0.5))

##caste x age
autoplot(pca_no51, data = sampleData_no51, colour = "age_at_end", shape = "caste", size=3.5) +
   scale_shape_manual(values=c(8,19)) +
   scale_color_continuous(type = "viridis") +
   annotate("text", x = 0.59, y = -.34, label = "K23D079", color = "black") +
   ggtitle("Age x Caste") +
    theme(plot.title = element_text(hjust = 0.5))

##tx_category x age
autoplot(pca_no51, data = sampleData_no51, colour = "age_at_end", shape = "tx_category", size=3.5) +
   scale_shape_manual(values=c(15,18, 17)) +
   scale_color_continuous(type = "viridis") +
   annotate("text", x = 0.59, y = -.34, label = "K23D079", color = "black") +
   ggtitle("Age x Category") +
    theme(plot.title = element_text(hjust = 0.5))


```


# PC Loading
Looking at PC loading.
```{r loadings, echo=FALSE, fig.width= 8, fig.height=8}
pc_loadings <- pca_no51$rotation


pc_loadings <- pc_loadings %>% 
  as_tibble(rownames = "gene")


top_genes <- pc_loadings %>% 
  # select only the PCs we are interested in
  select(gene, PC1, PC2) %>%
  # convert to a "long" format
  pivot_longer(matches("PC"), names_to = "PC", values_to = "loading") %>% 
  # for each PC
  group_by(PC) %>% 
  # arrange by descending order of loading
  arrange(desc(abs(loading))) %>% 
  # take the 10 top rows
  slice(1:10) %>% 
  # pull the gene column as a vector
  pull(gene) %>% 
  # ensure only unique genes are retained
  unique()

top_genes


top_loadings <- pc_loadings %>% 
  filter(gene %in% top_genes)


ggplot(data = top_loadings) +
  geom_segment(aes(x = 0, y = 0, xend = PC1, yend = PC2), 
               arrow = arrow(length = unit(0.1, "in")),
               color = "brown") +
  geom_text(aes(x = PC1, y = PC2, label = gene), size = 3) +
  scale_x_continuous(expand = c(0.05, 0.02))


```




## Removed: *K23C051* AND *"K23D079"*
```{r PCAwithoutCsample2}
#Remove K23C051
samples <- rownames(pca_matrix0)
pca_matrix_no5179 <- pca_matrix0[!samples %in% c("K23C051","K23D079"),]
dim(pca_matrix_no5179)



# Perform the PCA
pca_no5179 <- prcomp(pca_matrix_no5179)
pc_eigenvalues_no5179 <- pca_no5179$sdev^2


#to view using ggplot2 create a tibble
# create a "tibble" manually with 
# a variable indicating the PC number
# and a variable with the variances
pc_eigenvalues_no5179 <- tibble(PC = factor(1:length(pc_eigenvalues_no5179)), 
                         variance = pc_eigenvalues_no5179) %>% 
  # add a new column with the percent variance
  mutate(pct = variance/sum(variance)*100) %>% 
  # add another column with the cumulative variance explained
  mutate(pct_cum = cumsum(pct))

# print the result
pc_eigenvalues_no5179



####SCREE
pc_eigenvalues_no5179 %>% 
  ggplot(aes(x = PC)) +
  geom_col(aes(y = pct)) +
  geom_line(aes(y = pct_cum, group = 1)) + 
  geom_point(aes(y = pct_cum)) +
  labs(x = "Principal component", y = "Fraction variance explained")


##PCA plot##
pc_scores_no5179 <- pca_no5179$x


pc_scores_no5179 <- pc_scores_no5179 %>% 
  # convert to a tibble retaining the sample names as a new column
  as_tibble(rownames = "sample")

# print the result
pc_scores_no5179

#PLOT
PC_plot3 <- 
  ggplot(aes(x = PC1, y = PC2), data=pc_scores_no5179) +
  geom_point() 

PC_plot3

#PC_plot2 + geom_label(aes(label = sample))
```


```{r plotspice2, fig.width= 8, fig.height=4.75, echo=FALSE}
##lets add some details
sampleData_no5179 <- sampleData[-c(9,14),]


##tx_category x caste
autoplot(pca_no5179, data = sampleData_no5179, colour = "tx_category", shape = "caste", size=3.5) +
   scale_shape_manual(values=c(8,19)) +
   scale_colour_manual(values=c("forestgreen","coral","purple1"))+
   scale_fill_manual(values=c("forestgreen","coral","purple1")) +
   ggtitle("Category x Caste") +
    theme(plot.title = element_text(hjust = 0.5))

##caste by colony
autoplot(pca_no5179, data = sampleData_no5179, colour = "colony", shape = "caste", size=3.5) +
   scale_shape_manual(values=c(8,19)) +
   ggtitle("Colony x Caste") +
    theme(plot.title = element_text(hjust = 0.5))

##caste x age
autoplot(pca_no5179, data = sampleData_no5179, colour = "age_at_end", shape = "caste", size=3.5) +
   scale_shape_manual(values=c(8,19)) +
   scale_color_continuous(type = "viridis") +
   ggtitle("Age x Caste") +
    theme(plot.title = element_text(hjust = 0.5))


##tx_category x age
autoplot(pca_no5179, data = sampleData_no5179, colour = "age_at_end", shape = "tx_category", size=3.5) +
   scale_shape_manual(values=c(15,18, 17)) +
   scale_color_continuous(type = "viridis") +
   ggtitle("Age x Category") +
    theme(plot.title = element_text(hjust = 0.5))


```


# PC Loading
Looking at PC loading.
```{r loadings2, echo=FALSE, fig.width= 8, fig.height=8}
pc_loadings_no5179 <- pca_no5179$rotation


pc_loadings_no5179 <- pc_loadings_no5179 %>% 
  as_tibble(rownames = "gene")


top_genes_no5179 <- pc_loadings_no5179 %>% 
  # select only the PCs we are interested in
  select(gene, PC1, PC2) %>%
  # convert to a "long" format
  pivot_longer(matches("PC"), names_to = "PC", values_to = "loading") %>% 
  # for each PC
  group_by(PC) %>% 
  # arrange by descending order of loading
  arrange(desc(abs(loading))) %>% 
  # take the 10 top rows
  slice(1:10) %>% 
  # pull the gene column as a vector
  pull(gene) %>% 
  # ensure only unique genes are retained
  unique()

top_genes_no5179


top_loadings_no5179 <- pc_loadings_no5179 %>% 
  filter(gene %in% top_genes_no5179)


ggplot(data = top_loadings_no5179) +
  geom_segment(aes(x = 0, y = 0, xend = PC1, yend = PC2), 
               arrow = arrow(length = unit(0.1, "in")),
               color = "brown") +
  geom_text(aes(x = PC1, y = PC2, label = gene), size = 3) +
  scale_x_continuous(expand = c(0.05, 0.02))


```

```{r saveFilteredGenes_extra}

GEcounttable_Filteredgenes_outliersremoved <- GEcounttable_Filtered %>% select(-"K23C051", -"K23D079")

sampleData_outliersRemoved <- sampleData0[-c(9,14),]



write.table(sampleData_outliersRemoved, file = "sampleData_OutliersRemoved.txt", sep = "")
write.csv2(sampleData_outliersRemoved, file = "sampleData_OutliersRemoved.csv", row.names = FALSE)
```

